### 2020-01-13 2.ResearchReport/2.Presentation/Presentation_Hanne.Rmd
clean-up research report folder and edit readme

---
title: "Developing ShinyMICE"
subtitle: "A Note on Convergence Diagnostics for Multiple Imputation Algorithms"
author: "Hanne Oberman, December 16, 2019"
date: "Supervisors: Stef van Buuren and Gerko Vink | Mentor: Daniela Cianci"
output: 
    ioslides_presentation:
        logo: figures/shinymice.png
        transition: 0
        smaller: TRUE
bibliography: ShinyMICE.bib
nocite: | 
  @buur18, @R, @mice, @shiny, @veht19
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library("mice")
library("magrittr")
library("DT")
library("plotly")
library("dplyr")
library("ggplot2")
library("ggpubr")

load("C:/Users/User/Desktop/shinyMice/Simulation/Results/final.Rdata")
```

<style>
div.footnotes {
  position: absolute;
  bottom: 0;
  margin-bottom: 10px;
  width: 80%;
  font-size: 0.6em;
}
</style>

<script src="https://ajax.googleapis.com/ajax/libs/jquery/3.1.1/jquery.min.js"></script>

<script>
  $(document).ready(function() {
    $('slide:not(.backdrop):not(.title-slide)').append('<div class=\"footnotes\">');

    $('footnote').each(function(index) {
      var text  = $(this).html();
      var fnNum = (index+1).toString().sup();
      $(this).html(text + fnNum);

      var footnote   = fnNum + ': ' + $(this).attr('content') + '<br/>';
      var oldContent = $(this).parents('slide').children('div.footnotes').html();
      var newContent = oldContent + footnote;
      $(this).parents('slide').children('div.footnotes').html(newContent);
    });
  });
</script>
<!-- ## Disclaimer -->

<!-- This presentation only serves the illustrative purpose of showing that I am able to create a presentation in R Markdown. -->

<!-- ## This presentation | and the entire research archive is available on:  -->

<!-- <center> -->
<!-- [github.com/gerkovink/ShinyMICE/](https://github.com/gerkovink/ShinyMICE/) -->
<!-- </center> -->
<!-- <br><br><br> -->
<!-- You can find all related materials, links and references there.  -->

<!-- ## Content of the Presentation -->

<!-- <div class="columns-2"> -->
<!-- <img src="figures/shinymice.png" alt="HTML5 Icon" width = 110%> -->

<!-- <br><br><br> -->

<!-- - Introduction -->
<!-- - Methodology -->
<!-- - Simulation Results -->
<!-- - Discussion -->
<!-- </div> -->

<!-- # Introduction -->

<!-- ## Multiple imputation -->
<!-- - Missing data is ubiquitous [@alli02].  -->
<!-- - *Ad hoc* solutions may yield invalid inferences [@buur18]. -->
<!-- - @rubin87 proposed the framework of MI. -->

## Multiple imputation <footnote content="Figure adapted from van Buuren (2018)"></footnote>

<br><br>

<center>
<img src="figures/steps.png" alt="HTML5 Icon" width = 100%>
</center>

## Algorithmic convergence <footnote content="Figure adapted from van Buuren (2018)"></footnote>
<center>
<img src="figures/non-conv.png" alt="HTML5 Icon" width = 77%>
<br>

<!-- Iteration number -->
</center>

<!-- # Methodology -->

<!-- ## Mock code -->
<!-- ```{r eval=FALSE} -->
<!-- simulation <- function() { -->
<!--   mvtnorm(X,Z1,Z2) %>%  -->
<!--   mutate(Y~X+Z1+Z2) %>%  -->
<!--   for (max_iterations in 1:100) { -->
<!--     ampute() %>%  -->
<!--     impute() %T%  -->
<!--     convergence_diagnostics %>%  -->
<!--     lm(Y~X+Z1+Z2) %>%  -->
<!--     pool %>%  -->
<!--     simulation_diagnostics %>%  -->
<!--     c(., convergence_diagnostics) -->
<!--   } -->
<!-- } -->

<!-- replicate(simulation, n = 1000) -->
<!-- ``` -->

<!-- # Simulation Results -->

## Results: simulation diagnostics <footnote content="Results obtained using R (2019)"></footnote>
<center>
<img src="figures/results_sim.png" alt="HTML5 Icon" width = 100%>
<br>

</center>

## Results: convergence diagnostics <footnote content="Results obtained using R (2019)"></footnote>
<center>
<img src="figures/results_conv.png" alt="HTML5 Icon" width = 100%>
<br>

</center>

## Results <footnote content="Results obtained using R (2019)"></footnote>
<center>
<img src="figures/results_sim_conv.png" alt="HTML5 Icon" width = 100%>
<br>

</center>

<!-- ## Results | Simulation diagnostics versus convergence diagnostics -->
<!-- ```{r echo=FALSE} -->
<!-- dat %>% select("It.", "bias", "CIW", "cov", "R_mean_X", "AC_mean_X") %>% round(3) %>% datatable(options = list(pageLength = 5), rownames = FALSE) -->
<!-- ``` -->

<!-- ## Inspect the autocorrelation dip -->
<!-- ```{r echo=FALSE} -->
<!-- dat %>% round(3) %$%  plot_ly(x = It., y = AC_mean_X,  -->
<!--         text = paste("R hat: ", R_mean_X), -->
<!--         type = "scatter", mode = "lines") %>%  -->
<!--   layout(#title = "Auto-correlation across iterations", -->
<!--          xaxis = list(title = "Nr. of iterations"), -->
<!--          yaxis = list(title = "Auto-correlation (at lag 1)")) -->
<!-- ``` -->

<!-- # Discussion -->

## Discussion <footnote content="Figure adapted from mice (2014) and Shiny (2017)"></footnote>
<center>
<img src="figures/shinymice.png" alt="HTML5 Icon" width = 71%>
</center>

<center>
<img src="figures/empty.png" alt="HTML5 Icon" width = 1%>
</center>

## Impression of ShinyMICE <footnote content="Figure adapted from mice (2014) and Shiny (2017)"></footnote>
<center>
<br>

<img src="figures/shiny.png" alt="HTML5 Icon" width = 90%>
</center>

## Inspect $\widehat{R}$
```{r echo=FALSE}
dat %>% round(3) %$%  plot_ly(x = It., y = R_mean_X, 
        text = paste("AC: ", AC_mean_X),
        type = "scatter", mode = "lines") %>% 
  layout(#title = "R across iterations",
         xaxis = list(title = "Nr. of iterations"),
         yaxis = list(title = expression(paste(widehat(R)))))

```

## Inspect autocorrelation
```{r echo=FALSE}
dat %>% round(3) %$%  plot_ly(x = It., y = AC_mean_X, 
        text = paste("R hat: ", R_mean_X),
        type = "scatter", mode = "lines") %>% 
  layout(#title = "Auto-correlation across iterations",
         xaxis = list(title = "Nr. of iterations"),
         yaxis = list(title = "Auto-correlation (at lag 1)"))
```

## Results | Simulation diagnostics versus convergence diagnostics
```{r echo=FALSE}
dat %>% select("It.", "bias", "CIW", "cov", "R_mean_X", "AC_mean_X") %>% round(3) %>% datatable(options = list(pageLength = 5), rownames = FALSE)
```

## Equations $\widehat{R}$ <footnote content="Adapted from Vehtari et al. (2019)"></footnote>
\[\begin{align}
B&=\frac{T}{M-1} \sum_{m=1}^{M}\left(\bar{\theta}^{(\cdot m)}-\bar{\theta}^{(\cdot \cdot)}\right)^{2}, \quad \text{where} \quad \bar{\theta}^{(\cdot m)}=\frac{1}{T} \sum_{t=1}^{T} \theta^{(t m)}, \\
&\quad \text{ and } \quad \bar{\theta}^{(\cdot \cdot)}=\frac{1}{M} \sum_{m=1}^{M} \bar{\theta}^{(\cdot m)} \\
W&=\frac{1}{M} \sum_{m=1}^{M} s_{j}^{2},  \quad \text{where} \quad s_{m}^{2}=\frac{1}{T-1} \sum_{t=1}^{T}\left(\theta^{(t m)}-\bar{\theta}^{(\cdot m)}\right)^{2}  \\
\widehat{R}&=\sqrt{\frac{\widehat{\operatorname{var}}^{+}(\theta | y)}{W}},
\quad \text{where} \quad \widehat{\operatorname{var}}^{+}(\theta | y)=\frac{N-1}{N} W+\frac{1}{N} B.
\end{align}\]

<center>
<img src="figures/empty.png" alt="HTML5 Icon" width = 1%>
</center>

## References