### 2020-02-18 3.Thesis/inspect_incomplete_data.Rmd
added functionalities to apply lm to all variables and missingness indicators

---
title: "Investigate relations in incomplete datasets to inform imputation model"
output:
  html_document:
    df_print: paged
---

The goal is to provide plots and tests to inspect: 1) the relation between observed values of one variable and the missingness indicator of another; and 2) the relation between the observed values of two variables before imputation.

# Part 1: Use 1 variable from 1 dataset

```{r echo=TRUE, message=FALSE, warning=FALSE}
# set-up environment
library(mice) # Data imputation
library(dplyr) # Data manipulation
library(magrittr)
library(broom)
library(ggplot2)
library(purrr)
set.seed(123)
```

We will use the boys data as an example, specifically the variables age and gen. We will later apply it to all variables of this dataset, and then make it compatible for different datasets as well.

```{r}
# # pre-processing
# Rgen <- is.na(boys$gen)
# levels(Rgen) <- c("Missing", "Observed")
dat <- boys # %>% as_tibble # %>% cbind(Rgen)

# plot relation between observed values in age and missingness in gen
dat %>%
  ggplot(aes(x = age, fill = is.na(gen))) +
  geom_histogram(aes(y = 100 * ..density..),
                 alpha = 0.5,
                 position = 'identity') +
  xlab("Age (years)") +
  ylab("Percentage observed") +
  scale_fill_manual(
    name = "Gen",
    breaks = c("FALSE", "TRUE"),
    labels = c("Observed", "Missing"),
    values = c("#006CC2", "#B61A51")
  )

# evaluate statistically
out_Rgen <- dat %>% lm(age ~ is.na(gen), data = .)
out_int <- dat %>% lm(age ~ 1, data = .)
anova(out_int, out_Rgen)
```

**To do:** Create a similar plot for categorical outcome variable.

Now look at relations between the observed values of these variables. **Question:** Should I also visualize this relationship?

```{r}
out_gen <- dat %>% lm(age ~ gen, data = .)
out_int_gen <- dat %>% filter(!is.na(gen)) %>% lm(age ~ 1, data = .)
anova(out_int_gen, out_gen)
```

So both tests indicate that we should include age in the imputation model for gen.

# Part 2: Use any variable from 1 dataset

To apply these evaluations to all variables, we need a function. 
```{r}
# explore map functionalities
test_miss <- function(.x){
  return(is.na(.x))
}
map(c(1, NA, 4, 7), test_miss)

# inline function
map_df(c(1, NA, 4, 7), function(.x) {
  return(data.frame(entry = .x, 
                    is_NA = test_miss(.x)))
})

# shorter inline function
map_df(c(1, NA, 4, 7), ~{
  return(data.frame(entry = .x, 
                    is_NA = test_miss(.x)))
})

# test map functionalities
dat %>%  map_df(is.na)
dat %>% map_df(~(data.frame(n_miss = sum(is.na(.x)),
                                  class = class(.x))),
                     .id = "variable")

# apply map to our own problem
interc_only <- dat %>% lm(age ~ 1, data = .)
#age <- dat[,1]
# output_for_table <- dat %>% 
#   map(~{lm(dat$age~is.na(.x)) %>% 
#       anova(interc_only, .)}, .id = "variable")
dat %>% 
  map(~{lm(age~is.na(.x), data = dat) %>% anova(interc_only, .)})

# this works, but only for age as outcome. now apply to all variables.

```

**Question:** how do we evaluate these? Just look at p-values? Then, do we need some alpha inflation correction?

Also create the density plots for all missingness indicators.

```{r}
# plot this for each missingness indicator
vars <- names(dat)
vars %>%
  map(~{ggplot(data=dat, aes(x = age, fill = is.na(.data[[.x]]))) +
  geom_histogram(aes(y = 100 * ..density..),
                 alpha = 0.5,
                 position = 'identity') +
  xlab("Age (years)") +
  ylab("Percentage observed") +
  scale_fill_manual(
    name = .x,
    breaks = c("FALSE", "TRUE"),
    labels = c("observed", "missing"),
    values = c("#006CC2", "#B61A51")
  )})

# now for all vars as outcome
```
# Part 2B: Apply to all variables as outcome

This works for age now, with all missingness indicators. But not yet for other variables as outcome.

```{r}
# apply map to our own problem
interc_only <- dat %>% lm(age ~ 1, data = .)
vars <- names(dat)
#dat %>% map(~class(.x)=="numeric")
#dat[ ,map_lgl(dat, is.numeric)]
#dat %>% map(~is.numeric(.x)) select_if(x, is.numeric)
cont_vars <- dat %>% select_if(is.numeric) %>% names()
#walk(cont_vars,~ walk(vars, ~ print(paste(.x, .y, sep = "-")),.y=.x))
#test <- map(cont_vars, function(x) map(vars, function(y) c(x,y)))
#test <- map(cont_vars, function(x) map(vars, function(y) c(dat[1,x],is.na(dat[1,y]))))
test <- map(cont_vars, function(x) map(vars, function(y) lm(dat[,x]~is.na(dat[,y]))))

# this works for the lm, but not yet for the intercep-only model
#all_interc_only <- cont_vars %>% map(~lm(dat[,.x] ~ 1))
# test2 <- map(cont_vars, function(x) map(vars, function(y) lm(dat[,x]~is.na(dat[,y])) %>% anova(interc_only[[x]], .)))
# map(test, function(x) map(interc_only, function(y) anova(y,x)))
test <- map(cont_vars, function(x) map(vars, function(y) {anova(lm(dat[,x]~1), lm(dat[,x]~is.na(dat[,y])))}))

```

# Part 3: Use any variable from any dataset