### 2020-02-19 3.Thesis/plot_imputation_densities.Rmd
created density plots of imputed values per case, but only for one variable and without map/rep to add iterations

---
title: "Plot density of imputations"
output: html_notebook
---

```{r echo=TRUE, message=FALSE, warning=FALSE}
# set-up environment
library(mice) # Data imputation
library(dplyr) # Data manipulation
# library(magrittr)
# library(broom)
library(ggplot2)
# library(purrr)
set.seed(123)
dat <- nhanes
mids <- list()

# impute missingness with one iteration
M <- 20
mids <- mice(dat, maxit = 1, m = M, print = FALSE)

####
# this should be repeated T times
####
# add an iteration
mids <- mice.mids(mids, maxit = 1, printFlag = FALSE)
# # traceplots
# plot(mids)

# plot density of imputed values
imputed_values <- mids[["imp"]]$bmi %>% mutate(case = row_number()) %>%  tidyr::gather("imp", "value", 1:M) %>% .[order(.$case, .$imp), ] #%>% group_by(case) %>% mutate(case_mean = mean(value)) %>% ungroup()
case_mean <- imputed_values %>% group_by(case) %>% summarise(mu = mean(value)) 
# # for one case
# imputed_values %>% filter(case==3) %>%  ggplot() +
#   geom_histogram(aes(x = value)) + 
#   geom_vline(xintercept = mean(imputed_values$value))

# # store names of cases
# missing_cases <- 1:max(imputed_values$case)#unique(imputed_values$case)
# # plot histograms
# missing_cases %>%
#   purrr::map(function(z) {imputed_values %>% filter(case==z) %>% ggplot() +
#   geom_histogram(aes(x = value)) 
#   })

# plot hist per case
imputed_values %>% ggplot(aes(x = value)) + 
  geom_histogram() +
  geom_vline(data = case_mean, mapping = aes(xintercept = mu)) +
  facet_wrap(~case)

```