### 2020-03-05 3.Thesis/quotes_to_include.Rmd
added file with references for manuscript, adjusted densities/mean/sd of imputations per iteration, per case

---
title: "Literature on non-convergence"
author: "Hanne Oberman"
date: "4-3-2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

"As MICE is an iterative procedure, it is important that convergence is achieved. This may be checked by computing, at
each cycle, the means of imputed values and/or the values of regression coefficients, and seeing if they are stable. For the
example in Section 9, these values appeared stable from the very first cycle. We have never found 10 cycles inadequate,
but larger numbers of cycles might in principle be required when incomplete variables are very strongly associated" (@whit11, p. 394).

# For ShinyMICE
"Determining How Many Imputations Are Needed
Working back from the computation of statistics and confidence intervals, we need
to ask how many imputations are required to generate useful statistical inferences. Surprisingly, the answer is often no more than five and sometimes as little as two or three.
The relative efficiency (RE) of the m imputations given a fraction of missing data
(Rubin, 1987, p. 114) is computed in standard error units as
...
Even when γ is .9, five imputations give a relative efficiency of .92. Consequently,
one rarely has to create more than five imputations, and 10 imputations are more than
suitable in almost any realistic application" (https://journals.sagepub.com/doi/pdf/10.1177/1094428103255532, p. 291).