### 2020-04-29 3.Thesis/1.SimulationStudy/Writeup/manuscript.Rmd
Merge old and new methods sections

---
title: "Missing the Point: Non-Convergence in Iterative Imputation Algorithms"
runninghead: Oberman
author:
  name: H. I. Oberman
  num: 1
address:
  num: 1
  org: Department of Methodology and Statistics, Utrecht University, Utrecht, The Netherlands
corrauth: "Hanne Oberman, Sjoerd Groenman building, Utrecht Science Park, Utrecht, The Netherlands."
email: h.i.oberman@uu.nl
abstract: "Iterative imputation is a popular tool to accommodate missing data. While it is widely accepted that valid inferences can be obtained with this technique, these inferences all rely on algorithmic convergence. There is no consensus on how to evaluate the convergence properties of the method. This paper provides insight into identifying non-convergence of iterative impuation algorithms."
keywords: missing data, iterative imputation, non-convergence, mice
classoption:
 - Royal
 - times
bibliography: thesis.bib
bibliographystyle: sageh
output:
 rticles::sage_article:
  keep_tex: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 4.9, echo = FALSE, fig.align = "center") 
# knitr::opts_chunk$set(fig.width = 4.8, fig.height = 3.6, echo = FALSE, fig.align = "center")

# packages
library(ggplot2)
library(patchwork)

# results and functions
load("../Results/complete.Rdata")
source("../2.CreateFigures.R")

# default graphing behavior
theme_update(
 plot.title = element_text(hjust = 0.5),
 plot.subtitle = element_text(hjust = 0.5),
 panel.border = element_blank(),
 panel.grid.major = element_blank(),
 panel.grid.minor = element_blank(),
 panel.background = element_blank(),
 axis.line = element_line(colour = "black"),
 legend.key = element_blank(),
 legend.position = "bottom",
 legend.margin = margin(0,0,0,0),
 # adjust text size
 text = element_text(size = 8),
 title = element_text(size = 8)
 # axis.title = element_text(size = 8),
 # axis.text = element_text(size = 7),
 # legend.text = element_text(size = 7)
)
```

# Introduction

Anyone who analyzes person-data may run into a missing data problem. Missing data is not only ubiquitous, but treating it can also be tedious. If a dataset contains just one incomplete observation, statistical inferences are undefined and will not produce any results. To circumvent this, many statistical packages employ list-wise deletion by default (i.e., ignoring incomplete observations). Unfortunately, this *ad hoc* solution may yield wildly invalid results [@buur18]. An alternative is to *impute* (i.e., fill in) the missing values in the incomplete observations. Subsequently, statistical inferences can be performed on the completed dataset. By repeating this process several times, a distribution of plausible results may be obtained, which reflects the uncertainty in the data due to missingness. This technique is known as 'multiple imputation' [MI; @rubin76]. MI has proven to be a powerful technique to yield unbiased and confidence valid estimates of the true---but missing---data inference under many circumstances [@buur18].

Figure \ref{fig:diagram} provides an overview of the steps involved with MI---from incomplete data, to $M$ multiply imputed datasets, to $M$ estimated quantities of interest $\hat{Q}$s, to a single pooled estimate $\bar{Q}$. Missing data in dataset $y$ is imputed $M$ times. The imputed data $y_{imp}$ is combined with the observed data $y_{obs}$ to create $M$ completed datasets. On each completed dataset, the analysis of scientific interest is performed. The quantity of scientific interest (e.g., a regression coefficient) is denoted with $Q$. Since $Q$ is estimated on each completed dataset, $M$ separate $\hat{Q}$-values are obtained. These $M$ values are combined into a single pooled estimate $\bar{Q}$.

<!-- Q is quantity of scientific interest: "A scientific estimand $Q$ is a quantity of scientific interest that we can calculate if we would observe the entire population" [@buur18, par 2.3.1] -->

```{r diagram, out.width="\\linewidth", include=TRUE, fig.cap="Scheme of the main steps in multiple imputation."}
knitr::include_graphics("./images/diagram.pdf") #adapted from [@buur18] par. 1.4.1.
```

A popular method to obtain imputations is to use the 'Multiple Imputation by Chained Equations' algorithm, shorthand 'MICE'[@mice]. MICE is an iterative algorithmic procedure to draw imputations from the posterior predictive distribution of the missing values. This introduces a potential threat to the validity of the imputations: What if the algorithm has not converged? Are the implications then to be trusted? And can we rely on the inference obtained on the completed data? These are all open questions, because the convergence properties of iterative imputation algorithms have not been systematically studied [@buur18]. Moreover, there is no scientific consensus on how to evaluate convergence of MI algorithms [@taka17]. Some default MICE techniques (e.g., 'predictive mean modeling') might not yield converged states at all [@murr18]. Therefore, algorithmic convergence should be monitored carefully.

Currently, the recommended practice for evaluating convergence in iterative imputation algorithms is to visually inspect imputations for signs of non-convergence. **Add that this refers to chain means and variances already? It's insufficient that these are univariate, so multivariate state space of the algorithm may not be converged when univariates are ok. And @buur18's suggestion to track user-defined statistics (e.g., regression coefficient) may be somewhat advanced for empirical researchers. Moreover, this is not model-independent, i.e., only apply to one substantive model, while one of the advantages of MI is that missing data problem and scientific problem are split. Then we can say 'we propose a novel...' and integrate this: Note that usually we only evaluate the convergence of univariate scalar summaries (e.g., chain means or variances). With these we cannot diagnose convergence of multivariable statistics (i.e., relations between scalar summaries). Van Buuren [-@buur18, $\S$ 4.5.2] proposed to implement multivariable evaluation of the MICE algorithm through eigenvalue decomposition building on the work of @mack03. Eigenvalues of a covariance matrix are a measure of the data's covariance.**.

This method is insufficient on two counts: 1) it may be challenging to the untrained eye, and 2) only severely pathological cases of non-convergence may be diagnosed [@buur18, $\S$ 6.5.2]. Therefore, a quantitative, diagnostic evaluation of convergence would be preferred. Yet, monitoring convergence of iterative imputation algorithms diagnostically is challenging. Iterative imputation algorithms such as MICE are Markov chain Monte Carlo (MCMC) methods. In MCMC methods, convergence is not from a scalar to a point but from one distribution to another. The values generated by the algorithm (e.g., imputed values) will vary even after convergence. Therefore, the aim of convergence diagnostics for MCMC methods is not to establish the point at which convergence is reached, but to monitor signs of non-convergence [@hoff09]. Several of such diagnostics exist for MCMC methods, but it is not known whether these are appropriate for iterative imputation algorithms. 
<!-- "a weak diagnostics is better than no diagnostic at all" [@cowl96]. -->

In this paper, we investigate how non-convergence in iterative imputation algorithms may be diagnosed, how well these methods perform, and at which point convergence may safely be assumed. For reasons of brevity, we only focus on the iterative imputation algorithm implemented in the popular `mice` package [@mice] in `R` [@R]. The convergence properties of the MICE algorithm are investigated through model-based simulation. The results of this simulation study are guidelines for assessing convergence of MI algorithms, which will aid applied researchers in drawing valid inference from incomplete datasets.

**Include sub-questions!** How can non-convergence be identified diagnostically? Are common MCMC non-convergence diagnostics appropriate for MICE? And if so, which threshold should be used to diagnose non-convergence? How many iterations are sufficient/needed to be able to diagnose non-convergence? Are the default number of iterations sufficient (i.e., 5 in mice, 10 in SPSS and Stata, 30 in mi)? **How severe is it when the algorithm has not converged? And what are guidelines for practice? Can the parameter of interest, estimand $Q$, be correct when the algorithm is not (yet) converged, and vice versa?** *(Maybe add these too? What are the effects of continued iterations on estimates, predictions and inferences? Do the answers differ with varying missingness proportions? That is, we vary the nr of iterations and the missingness proportion because we assume that the algorithm has not reached convergence at $t=1$, and performs worse with moderate to high missingness (not so much with little or a LOT of missingness).)*

<!-- The defaults of common MI software: SPSS = 10 [(link)](https://www.ibm.com/support/knowledgecenter/SSLVMB_24.0.0/spss/mva/syn_multiple_imputation_impute.html), Mplus = 100?? [(link)](https://pdfs.semanticscholar.org/e20e/29e008592cbfbaa567931f74cdfdb5451405.pdf?_ga=2.55354671.54033656.1584698748-527613517.1584698748), Stata = 10 [(link)](https://www.stata.com/manuals13/mi.pdf, p. 139), Amelia = NA, because re-sampling, not convergence [(link)](https://cran.r-project.org/web/packages/Amelia/Amelia.pdf), en MI = 30 [(link)](https://cran.r-project.org/web/packages/mi/mi.pdf). -->

<!-- The goal of this research project is to develop novel methodology and guidelines for evaluating MI methods. These tools will subsequently be implemented in an interactive evaluation framework for multiple imputation, which will aid applied researchers in drawing valid inference from incomplete datasets. -->

## Some notation

Let $y$ denote an $n \times k$ matrix containing the data values on $k$ variables for all $n$ units in a sample. The data value of unit $i$ ($i = 1, 2, \dots, n$) on variable $j$ ($j = 1, 2, \dots, k$) may be either observed or missing. **The number of units $i$ with at least one missing data value can be divided by the total number of units $n$ to obtain the 'missingness proportion'  $p_{mis}$.** The collection of observed data values in $y$ is denoted by $y_{obs}$; the missing part of $y$ is referred to as $y_{mis}$. For each datapoint in $y_{mis}$, we sample $M \times T$ times plausible values, where $M$ is the number of imputations ($m = 1, 2, \dots, M$) and $T$ is the number of iterations ($t = 1, 2, \dots, T$). The collection of samples between the initial value (at $t=1$) and the final imputed value (at $t=T$) will be referred to as an 'imputation chain'. **Add: $\theta$s are scalar summaries of interest in the iterative algorithm (e.g., chain means; the average of the imputed values in each imputation chain). **

# Identifying non-convergence

Mixing and stationarity have histroically been inspected visually, by evaluating traceplots of scalar summaries of interest ($\theta$s; e.g., chain means and chain variances). As @buur18 describes, users can also specify a model-specific scalar summary (e.g., a regression coefficient). A user-specified scalar summary, however, is not universal to all complete data problems. Therefore, as inspired by [@mack03], we propose a $\theta$ that summarizes the multivariate state space of the algorithm. Namely, the first eigenvalue of the variance-covariance matrix of the $M$ completed datasets. **The first eigenvalue has the appealing property that is not dependent on the substantive model of interest.**

**Since convergence of iterative imputation algorithms is in distribution, there is not a unique point at which the algorithm reaches a converged state. We can only evaluate scalar summaries of the multivariate state space of the algorithm, $\theta$s. The recommended scalar summaries to evaluate are chain means and chain variances. Additionally, researchers may "monitor some statistic of scientific interest" [@buur18, $\S$ 6.5.2]. Because of the issues mentioned above, we propose a novel $\theta$ to monitor: the first eigenvalue of the variance-covariance matrix of the completed data. Let $\lambda_1 \geq \lambda_2 \geq ... \geq \lambda_j$ be the eigenvalues of $\Sigma$ in each of the $M$ completed datasets ${y_{obs}, y_{imp}}$. $\lambda_1$ is measure that summarises the covariances in the completed datasets.**

There are two requirements for convergence of iterative algorithms: mixing and stationarity [@gelm13]. Without mixing, imputation chains do not intermingle nicely, **indicating that ...**. Without stationarity, there is trending within imputation chains, which implies that further iterations would yiled a different set of imputations.

To illustrate what non-convergence looks like in MI, we reproduce the example from van Buuren [-@buur18, $\S$ 6.5.2]. Figure \ref{fig:non-conv} shows the traceplot for one of the variables in the example. we see the average of the imputed values for a variable $j$ in $y_{imp}$. The average imputed values in the left hand plot are about a magnitude 2 lower than the right hand plot. So non-convergence has a impact on the average imputed value for $j$ in $y_{imp, m}$. This difference (presumably) translates to bias in the pooled estimate $\bar{\mu}_j$ as well. And maybe also to higher order statistics, e.g. variances, covariances, etc. Therefore, it's important to converge. **Explain what we see, namely example by @buur18 reproduced, showing the traceplots of chain means for some variable. The first plot is typical convergence of MICE, the second is pathological non-convergence because of a mis-specified imputation model. Each line is an imputation. In the first plot, the chains intermingle nicely and there is little to no trending. In the second plot, there is a lot of trending and some chains do not intermingle. Importantly, the chain means at the last iteration (the imputed value per $m$) are very different between the two plots. The algorithm with the mis-specified model yields imputed values that are on average a magnitude two larger than those of the typically converged algorithm. This shows the importance of reaching converged states in iterative imputation algorithms.**

## Methods under evaluation

<!-- Requirements for diagnostics to identify non-convergence in iterative imputation are: 1) work for a relatively low number of iterations (defaults in popular iterative imputation software between 5 and 30); 2) may be applied to different $\theta$s, including model-independent $\theta$s (e.g., chain mean, or eigenvalue). -->

Non-stationarity within chains may be diagnosed with e.g., autocorrelation [$AC$; @scha97; @gelm13], numeric standard error ['MC error'; @gewe92], or Raftery and Lewis's [-@raft91] procedure to determine the effect of trending on the precision of estimates. A widely used diagnostic to monitor mixing between chains is the potential scale reduction factor $\widehat{R}$ ['Gelman-Rubin statistic'; @gelm92]. With a recently proposed adaptation, $\widehat{R}$ might also serve to diagnose non-stationarity, but this has not yet been thoroughly investigated [@veht19]. Therefore, use we $\widehat{R}$ and $AC$ to evaluate mixing and stationarity separately, as recommended by e.g., @cowl96.
<!-- @cowl96, p. 898. -->

```{r non-conv, out.width='.49\\linewidth', fig.width=3.5, fig.height=3,fig.show='hold',fig.cap="Typical convergence versus pathological non-convergence."}
load("./images/nonc_plot.Rdata")
load("./images/conv_plot.Rdata")
par(mar = c(2, 2, 0.1, 0.1))
plot(conv)
plot(nonc)

```

### Potential scale reduction factor

An updated version of $\widehat{R}$ has been proposed by @veht19 (p. 5). **This version may be suitable for iterative imputation.** Let $M$ be the total number of chains, $T$ the number of iterations per chain, and $\theta$ the scalar summary of interest (e.g., chain mean or chain variance). For each chain ($m = 1, 2, \dots, M$), we estimate the variance of $\theta$, and average these to obtain within-chain variance $W$.

\begin{align*}
W&=\frac{1}{M} \sum_{m=1}^{M} s_{j}^{2}, \text { where } s_{m}^{2}=\frac{1}{T-1} \sum_{t=1}^{T}\left(\theta^{(t m)}-\bar{\theta}^{(\cdot m)}\right)^{2}. 
\end{align*}
<!-- %\text{ \cite[p.~5]{veht19}} -->

We then estimate between-chain variance $B$ as the variance of the collection of average $\theta$ per chain.

\begin{align*}
B&=\frac{T}{M-1} \sum_{m=1}^{M}\left(\bar{\theta}^{(\cdot m)}-\bar{\theta}^{(\cdot \cdot)}\right)^{2}, \text { where } \bar{\theta}^{(\cdot m)}=\frac{1}{T} \sum_{t=1}^{T} \theta^{(t m)} \text{, } \bar{\theta}^{(\cdot \cdot)}=\frac{1}{M} \sum_{m=1}^{M} \bar{\theta}^{(\cdot m)}. 
\end{align*}
<!-- %\text{ \cite[p.~5]{veht19}} -->

From the between- and within-chain variances we compute a weighted average, $\widehat{\operatorname{var}}^{+}$, which over-estimates the total variance of $\theta$ **remove or explain why, or leave in**. $\widehat{R}$ is then obtained as a ratio between the over-estimated total variance and the within-chain variance:

\begin{equation*}
\widehat{R}=\sqrt{\frac{\widehat{\operatorname{var}}^{+}(\theta | y)}{W}},
\text{ where } \widehat{\operatorname{var}}^{+}(\theta | y)=\frac{N-1}{N} W+\frac{1}{N} B.
\end{equation*}

We can interpret $\widehat{R}$ as potential scale reduction factor since it indicates by how much the variance of $\theta$ could be shrunken down if an infinite number of iterations per chain would be run [@gelm92]. This interpretation assumes that chains are 'over-dispersed' at $t=1$, and reach convergence as $T \to \infty$. Over-dispersion implies that the initial values of the chains are 'far away' from the target distribution and each other. When all chains sample independent of their initial values, the mixing component of convergence is satisfied, and $\widehat{R}$-values will be close to one. High $\widehat{R}$-values thus indicate non-convergence. The conventionally acceptable threshold for convergence was $\widehat{R} < 1.2$ [@gelm92]. More recently, @veht19 proposed a more stringent threshold of $\widehat{R} < 1.01$.

### Autocorrelation

Following the same notation, we define autocorrelation as the correlation between two subsequent $\theta$-values within the same chain [@lync07, p. 147]. In this study, we only consider $AC$ at lag 1, i.e., the correlation between the $t^{th}$ and $t+1^{th}$ iteration of the same chain.

\begin{equation*}
AC = \left( \frac{T}{T-1} \right) \frac{\sum_{t=1}^{T-1}(\theta_t - \bar{\theta}^{(\cdot m)})(\theta_{t+1} - \bar{\theta}^{(\cdot m)})}{\sum_{t=1}^{T}(\theta_t - \bar{\theta}^{(\cdot m)})^2}.
\end{equation*}

We can interpret $AC$-values as a measure of stationarity. If $AC$-values are close to zero, there is no dependence between subsequent samples within imputation chains. Negative $AC$-values indicate divergence within imputation chains. **Subsequent sampled values within each imputation chain are less alike.** Positive $AC$-values indicate recurrence. If $\theta$-values of subsequent iterations are similar, trending may occur. Negative $AC$-values show no threat to the stationarity component of convergence. On the contrary even---negative $AC$-values indicate that $\theta$-values of subsequent iterations diverge from one another, which may increase the variance of $\theta$ and speed up convergence. As convergence diagnostic, the interest is therefore in positive $AC$-values. **Maybe remove:** Moreover, the magnitude of $AC$-values may be evaluated statistically, but that is outside of our scope.

## In practice

**Before using them in the simulation, the two methods show at least be able to distinguish between the two algorithms plotted in Figure 2. From visual inspection we know that the typical convergence has some signs of non-mixing around iteration 2, and little trending. In the non-convergence situation, there is a lot of trending upto iteration 6, after which the chains reach a somewhat more stationary state. Mixing in this situation gets worse from the first iteration onward, and gradually gets better around iteration 7. To assess whether $\widehat{R}$ and $AC$ may be appropriate non-convergence identifiers for iterative imputation algorithms, the following condition must hold: the methods indicate worse performance for the mis-specified model with pathological non-convergence (i.e., higher $\widehat{R}$- and $AC$-values than the typical performance). And additionally the methods should reflect the increasing convergence in the typical convergence situation as the number of iterations goes up.**

**In Figure \ref{fig:diagnostics}A, the chain means from Figure \ref{fig:non-conv} are plotted again, now together---as are the non-convergence diagnostics. Panel B shows $\widehat{R}$ as computed by implementing @veht19 's recommendations. As required, $\widehat{R}$ indicates less signs of non-convergence as the number of iterations goes up **in the typical convergence situation**. The superior performance of the typical convergence over the pathological non-convergence is less prominent, even flipped for $t<4$. Panel C displays the $AC$ as computed with the R function `stats::acf()`. When we look at this panel, we conclude something weird. The $AC$-values indicate equal performance (up-to $t=5$) for the typical convergence and the pathological non-convergence, while there is obvious trending in the latter. Moreover, the best convergence (as indicated by the lowest $AC$-value) is observed at $t=2$, but looking at the chain means in panel A, there should be some signs of trending up-to iteration number seven. After consulting the documentation on stats::acf() [@R], we conclude that this implementation of $AC$ is not suitable for iterative imputation algorithms. There is a correction factor for a mathematical shortcut that works in the limit. According to @box15, the function works when the number of iterations > 50. The default number of iterations in iterative imputation, however, is often much lower. Therefore, we compute $AC$ manually, see panel D. The AC values in this plot do meet the requirements (...) and will therefore be used in the simulation study.**

**acf is niet bedoeld om weinig iteraties te... niet een fout in acf, alleen hier niet van toepassing**

```{r diagnostics, fig.cap="Convergence diagnostics pathological non-convergence versus typical convergence."}
load("./images/diagnostics_plot.Rdata")
diagnostics_plot
```

# Simulation study

**Add to this section: 1) explicit mention of simulation conditions; 2) emphasize that the plots are averages across repetitions, not within MICE; 3) sample effects due to single complete dataset**

The aim of the simulation study is to evaluate the impact of inducing non-convergence on several quantities of scientific interest $Q$, and to subsequently evaluate the performance of $\widehat{R}$ and $AC$ in identifying the non-convergence.

Non-convergence will be induced by: 1) varying the missingness proportions in $y$ ($p_{mis} =.05,.25,.50,.75,.95$), and 2) terminating the MICE algorithm at different imputation chain lengths ($t = 1, 2,..., 100$). The assumption underlying the different number of iterations is that the algorithm generally does reach convergence at $t=1$. The initial values in the first iteration are sampled randomly from the set of observed datapoints. **As the number of iterations goes up, the imputation chains become independent of the initial values until the point at which an extra added iteration does not lead to a more converged state.** The second set of experimental conditions under consideration--the missingness proportions--are chosen to reflect the difficulty of the missingness problem. The inherent assumption is that low missingness proportions will lead to quick algorithmic convergence, since there is a lot of information in the observed data. Higher missingness proportions then cause slower convergence. However, if the fraction of missing information is very high, there is so little information in the data that the random component in the algorithm will take the overhand and a stable but very uncertain (high variance) state will be reached.

## Hypothesis

We hypothesize that simulation conditions with a higer $p_miss$ and lower $t$ have inferior convergence. We assume that when $\bar{Q}$ is an unbiased and confidence valid estimate of $Q$, the algorithm is *sufficiently* converged.

**(bruggetje hier)** For multiple imputation algorithms, it holds that convergence is reached when there is no dependency between subsequent iterations of imputation chains ($AC = 0$), and chains intermingle such that the only difference between the chains is caused by the randomness induced by the algorithm ($\widehat{R} = 1$). **We expect that a completely converged state will not be reached.**

<!-- This study evaluates whether $\widehat{R}$ and $AC$ could diagnose convergence of multiple imputation algorithms. We assess the performance of the two convergence diagnostics against the recommended evaluation criteria for MI methods [i.e., average bias, average confidence interval width, and empirical coverage rate across simulations; @buur18, $\S$ 2.5.2]. **That is, there is no baseline measure available to evaluate performance against.** -->

<!-- We could also look at distributional characteristics, and plausibility of imputed values, see @vinknd. For now, this is outside of the scope of this study. -->

Based on an empirical finding [@lace07], we hypothesize that $\widehat{R}$ will over-estimate non-convergence of MI algorithms **explain why?**. Lacerda concluded that rhat dropped below 1.2 after at least 50 iterations. Therefore we hypothesize that Vehtari's proposed threshold of $\widehat{R} < 1.01$ will be too stringent for diagnosing convergence in iterative imputation. This over-estimation may, however, be diminished because $\widehat{R}$ can falsely diagnose convergence if initial values of the algorithm are not appropriately over-dispersed [@broo98, p. 437]. In e.g. `mice`, initial values are chosen randomly from the observed data. Therefore, we cannot be certain that the initial values are over-dispersed. We expect this to have little effect on the hypothesized performance of $\widehat{R}$. **(Add actual hypothesis: we thought that mice would converge sooner than $\widehat{R}$ would indicate that it did)** We hypothesize that high AC values are implausible in converged MI algorithms. **Added later: High $AC$-values are implausible in MI procedures. That is, the randomness induced by the MI algorithm effectively mitigates the risk of dependency within chains. But this doesn't seem the case in practice??**

## Set-up

Convergence of the MICE algorithm is investigated through model-based simulation in `R` [version `r paste(version$major,version$minor, sep = ".")`; @R]. The simulation set-up is summarized in the pseudo-code below. The complete `R` script of the simulation study is available from [github.com/gerkovink/shinyMice](https://github.com/gerkovink/shinyMice/tree/master/3.Thesis/1.SimulationStudy).

```
# pseudo-code of simulation 
1. simulate data 
for (number of simulation runs from 1 to 1000)
 for (missingness proportions 5%, 25%, 50%, 75% and 95%)
  2. create missingness
  for (number of iterations from 1 to 100)
   3. impute missingness
   4. perform analysis of scientific interest
   5. compute convergence diagnostics 
   6. pool results across imputations
   7. compute performance measures
 8. combine outcomes of all missingness proportions
9. aggregate outcomes across simulation runs

```

The data-generating mechanism is a multivariate normal distribution, representing person-data on three predictor variables (from an unspecified social scientific field of study). Let our predictor space be defined as 
\begin{align*}
\begin{pmatrix}X_1\\
X_2\\
X_3
\end{pmatrix} \sim \mathcal{N}
\begin{bmatrix}
\begin{pmatrix}
12\\
3\\
0.5
\end{pmatrix}\!\!,
\begin{pmatrix}
4 & 4 & 1.8 & 0\\
4 & 16 & 4.8 & 0\\
1.8 & 4.8 & 9 & 0
\end{pmatrix}
\end{bmatrix}\!\!\text{.}\\[2\jot]
\end{align*}

In this study, sampling variance is not of interest. Therefore, a single complete set may serve as comparative truth in all simulation runs [@vink14]. A finite population of $N=1000$ is simulated using the `mvtnorm` package [@mvtnorm]. Subsequently, a fourth variable is constructed  as outcome variable in a multiple linear regression problem. Let 
$$
Y_i = 1 + 2X_{1i} +.5X_{2i} - X_{3i} + \epsilon_i ,
$$
where $i = 1, 2,..., N$ and $\epsilon \sim \mathcal{N}(0, 100)$.

The complete data is *amputed* once for each simulation repetition with function `mice::ampute()`. The missingness is multivariate, and the probability to be missing is the same for all $n \times k$ cells in $y$ ['MCAR'; @rubin87]. Proper performance of the convergence diagnostics under at least 'missing completely at random' missingness mechanism is necessary to demonstrate the appropriateness of $\widehat{R}$ and $AC$ as convergence diagnostics. However, results may not be extrapolated to other missingness mechanisms. The missinness proportion varies between conditions ($p_{mis} =.05,.25,.5,.75,.95$).
<!-- This note only considers a 'missing completely at random' (MCAR) mechanism, where the probability of being missing is equal for all $n \times k$ cells in $y$ [@rubin87]. -->

Missing datapoints in $y$ are imputed with the `mice` package [@mice]. All MI procedures are performed with Bayesian linear regression imputation (`method = "norm"`), and five imputation chains (`m = 5`). The number of iterations varies between simulation conditions (`maxit = ` $1, 2, \dots, 100$). 
<!-- Each repetition of the simulation thus starts from 1 complete dataset, then splits into 5 missingness conditions ($p_{mis} =.05,.25,.5,.75,.95$), each of the amputed datasets is imputed 5 times ($m = 1, 2,..., 5$) for each iteration condition ($t = 1, 2,..., 100$). From these $5\times5\times100$ imputations, we extract the $\theta$s to apply $\widehat{R}$ and $AC$ on (**i.e., chain means, chain variances, $\beta$s per imputation, and the first eigenvalue of the variance-covariance matrix per imputation**).  -->

The methods under evaluation, $\widehat{R}$ and $AC$, are computed for four scalar summaries $\theta$: chain means (i.e., the mean in each $y_{imp, m, j}$), chain variances (i.e., the variance in each $y_{imp, m, j}$), the statistic of scientific interest $\hat{Q}$ in each {$y_{obs}, y_{imp, m}$}, and the first eigenvalue of the variance-covariance matrix in each {$y_{obs}, y_{imp, m}$}, $\lambda_1$. We calculate $\widehat{R}$ by implementing Vehtari et al.'s [-@veht19] recommendations, and $AC$ as the correlation between the $t^{th}$ and the $(t+1)^{th}$ iteration. For each $\theta$ we estimate the corresponding scientific estimand $Q$.

The performance of two methods on each of these scalar summaries will be evaluated with several estimands, i.e. quantities of scientific interest $Q$. For each $\theta$, one estimand $Q$ is defined that may be of interest in empirical research. The $\widehat{R}$ and $AC$ values with chain means as $\theta$s are evaluated against the bias in univariate mean estimates. Similarly, the performance measure for $\widehat{R}$ and $AC$ applied to chain variances is the bias in estimated standard deviations. To evaluate the performance of $\widehat{R}$ and $AC$ on the eigenvalues, we will use bias in the coefficient of determination, $R^2$. Additionally, we will evaluate the estimated regression coefficients, the coverage rate of the CI95% of the regression estimates, and the CI length (**see definitions in old Methods**). **remove the term performance measure here and add an extra paragraph to define performance measures as bias in all estimands, and coverage rate and CI length of regression coefficients.**

We consider four quantities of scientific interest, namely the descriptive statistics of all variables (mean and standard deviation), $\beta_2$, and the percentage of variance in $Y$ explained (coefficient of determination $R^2 \times 100$). We solve a multiple linear regression problem, where outcome $Y$ is regressed on predictors $X_1$, $X_2$ and $X_3$

$$Y \sim \beta_1 X_1 + \beta_2 X_2 + \beta_3 X_3.$$

The estimator for each $Q$ is $\bar{Q}$---the pooled aggregate of the $\hat{Q}$s across imputations. To calculate the $\bar{Q}$s, we combine the observed data $y_{obs}$ and the imputed data for each imputation $y_{imp,m}$ with the function `mice::complete()`. Descriptive statistics are computed as the average across imputations for each variable ($\mu_j$ and $\sigma_j$, where $j = Y, X_1, X_2, X_3$). Estimated regression coefficients are obtained with the function `stats::lm()` for each imputation, and then pooled conform @vink14. The coefficient of determination is estimated for each imputation, and pooled using `mice::pool.r.squared()`.

The performance of $\widehat{R}$ and $AC$ is assessed by comparing $\bar{Q}$s with $Q$s. For each $Q$, we compute bias as $\bar{Q} - Q$. For $Q=\beta_j$, we also compute the empirical coverage rate (CR). CR is defined as the percentage of simulation repetitions in which the 95\% confidence interval (CI) around $\bar{Q}$ covers the true estimand $Q$. Let

$$\text{CI} = \bar{Q} \pm t_{(M-1)} \times SE_{\bar{Q}},$$

where $t_{(M-1)}$ is the quantile of a $t$-distribution with $M-1$ degrees of freedom, and $SE_{\bar{Q}}$ is the square root of the pooled variance estimate. **Remove this? ** Under-estimating the variance of $\bar{Q}$ may yield spurious inferences. Confidence interval width (CIW) is defined as the difference between the lower and upper bound of the 95\% confidence interval (CI) around $\bar{Q}$ From bias and CIW, we calculate empirical coverage rates. Coverage rate is the proportion of simulations in which $Q$ is between the bounds of the CI95\% around $\bar{Q}$.

# Results

For reasons of brevity, we only discuss the convergence diagnostics for the $Q$ with the worst performance in terms of bias. For the descriptve statistics, the magnitude of the bias was the largest in $Y$ (i.e, $j = Y$ in $\mu_j$ and $\sigma_j$). For the $Q=\beta_j$, $j=2$ showed the worst perfomance (i.e, the effect of $X_1$ on $Y$.

As expected, conditions with a higher proportion of missingness and/or a lower number of iterations show more signs of non-convergence, as indicated by more extreme bias in the estimated $Q$s. Roughly speaking this means that MICE is indeed not converged as $t=1$, and converges gradually as $t$ increases. The point at which an additional iteration does not lead to an improvement of the estimates depends on the difficulty of the missingness problem---between XYZ and XYZ. This means that the algorithm has converged sufficiently (**under the current specifications**).

As required for a method to diagnose non-convergence, the values of $\widehat{R}$ follow a trend similar to the performance measures. $\widehat{R}$ values are generally lower in conditions with a higher number of iterations, and somewhat higher in conditions with a higher percentage of missing data. Autocorrelation values were indeed decreasing with a higher number of iterations.

Evaluation with the performance measures shows that $\widehat{R}$ and autocorrelation are conservative: they indicate signs of non-convergence in conditions where $\bar{Q}$s are unbiased and or confidence valid estimates of $Q$.

## Old Results section

**(Add more info about figure legends and axes.)**

```{r mean, fig.height = 6.8, fig.cap="Convergence diagnostics chain mean."}
mean_Rh + mean_AC + mean_bias + plot_layout(guides = "collect", ncol = 1) + plot_annotation(tag_levels = "A", tag_suffix = ".", title = "Performance of convergence diagnostics", subtitle = bquote(theta ~ "= chain mean of Y in" ~ y[imp][",m"] ~ "; Q = " ~ mu[Y] ~ "= 25.81"))
```

```{r sd, fig.height = 6.8, fig.cap="Convergence diagnostics chain variance."}
var_Rh + var_AC + sd_bias + plot_layout(guides = "collect", ncol = 1) + plot_annotation(tag_levels = "A", tag_suffix = ".", title = "Performance of convergence diagnostics", subtitle = bquote(theta ~ "= chain variance of Y in" ~ y[imp][",m"] ~ "; Q = " ~ sigma[Y] ~ "= 11.32"))
```
```{r est, fig.height = 6.8, fig.cap="Convergence diagnostics regression coefficient."}
est_Rh + est_AC + est_bias + plot_layout(guides = "collect", ncol = 1) + plot_annotation(tag_levels = "A", tag_suffix = ".", title = "Performance of convergence diagnostics", subtitle = bquote(theta ~ "= " ~ hat(beta)[2] ~ "in {" ~ y[obs] ~ "," ~ y[imp][",m"] ~ "}" ~ "; Q = " ~ beta[2] ~ "= 2.06"))
```

```{r pred, fig.height = 6.8, fig.cap="Convergence diagnostics regression coefficient."}
PCA_Rh + PCA_AC + Rsq_bias + plot_layout(guides = "collect", ncol = 1) + plot_annotation(tag_levels = "A", tag_suffix = ".", title = "Performance of convergence diagnostics", subtitle = bquote(theta ~ "= first eigenvalue of" ~ Sigma ~ "in {" ~ y[obs] ~ "," ~ y[imp][",m"] ~ "}" ~ "; Q = explained variance = 19.25"))

```

## Univariate estimates and convergence diagnostics

The bias in the estimates of the variable means shows little to no difference between simulation conditions. It doesn't seem to matter how many iterations you use in the mice algorithm, the estimates are unbiased. Similarly, the bias in the estimated variances is more or less stable across simulation conditions. These univariate quantities appear to be unaffected by the number of iterations.

When applied to the imputation chain means, $\widehat{R}$ indicates that the mice algorithm does not reach a converged state ($\widehat{R}$ = 1) in any of the simulation conditions. Neither is the most recent recommended threshold reached ($\widehat{R}$ < 1.01). **This is only for 75% missingness:** The conventional $\widehat{R}$ threshold of 1.2 is reached in simulation conditions $T = 3$ and $T > 6$. (With the default number of iterations (maxit = 5), this dip in $\widehat{R}$ values would be spotted, so it is no problem.) The point at which an extra iteration does not seem to improve the $\widehat{R}$ value is around $T=30$.

Autocorrelations indicate no sign of trending **(NOT correct anymore, fluke of the stats::acf() function!!)** within imputation chains. In most simulation conditions $AC$ is smaller than or about equal to zero. Across simulation conditions, however, the autocorrelation curve does not trend towards 0. Autocorrelation values plateau off at a value of around.1. This is a small positive autocorrelation, which would indicate some trending within chains.

The $\widehat{R}$ and $AC$ values for the imputation chain variances show equal trends to the chain means and are therefore not discussed separately. Taken together, univariate estimates seem robust across simulation conditions. There is no clear effect of the number of iterations on the bias in these estimates, while the convergence diagnostics indicate that the algorithm did not reach a completely converged state (yet).

## Multivariate estimates

**This is only true for 75% missingness:** There is a clear bias in the regression estimates in simulation conditions where the number of iterations is smaller than four. In simulation conditions where $T > 5$ there is little to no bias in the estimated regression coefficients. In most simulation conditions nominal coverage is obtained (i.e., coverage rates of.95) for the confidence intervals around the regression coefficients. Conditions with only one or two iterations show some under-coverage. Since confidence interval width is stable across conditions, the under-coverage may be attributed to the bias in the estimated regression coefficients.

**This is only true for 75% missingness:** If we look at the estimated proportion of explained variance in outcome variable Y we see that the coefficient of determination is underestimate estimated in conditions where the number of iterations is equal to two or less, and slightly overestimated in conditions where the number of iterations is equal to three or more.

In short, we see that the minimum number of iterations required to obtain unbiased, confidence valid regression estimates is 5. This value, however, is dependent on the percentage of missing values. E.g., with 95% of cases having missing data we need at least seven iterations to obtain unbiased results.

```{r thresh-Rh, include=FALSE}
# # chain means
# dat[-1, 6:9] %>% mutate(max_Rh = apply(., 1, max)) %>%.$max_Rh<1.2
# 
# # chain variances
# dat[-1, 26:29] %>% mutate(max_Rh = apply(., 1, max)) %>%.$max_Rh<1.2

```

# Discussion

Upon convergence, $\widehat{R}=1$ and $AC=0$, which are unlikely thresholds for MCMC algorithms, because of its convergence to a distribution. In practice, non-convergence is usually diagnosed when $\widehat{R}$ > 1.2 or 1.1 or even 1.01. And a t-test is performed to assess whether $AC$ is significantly different from zero.

As expected, complete convergence ($\widehat{R}=1$ and $AC=0$) is never observed. But what about the thresholds for practice? $\widehat{R}$ <1.2 is not stringent enough, because conditions in which $\widehat{R}$ is smaller than 1.2 do not all yield unbiased estimates of the $Q$s. Moreover, there is a dip in the $\widehat{R}$ values, after which $\widehat{R}$ > 1.2 again. If the algorithms would be terminated at the iteration where $\widehat{R}$ < 1.2 occurs for the first time, the increase in $\widehat{R}$ values might be missed. The threshold $\widehat{R}$ < 1.1 is somewhat conservative in comparison with the performance measures. Overall it seems OK. $\widehat{R}$ < 1.01 is too stringent compared to the performance measures. And it is not obtained for any number of iterations (**but may be necessary in more complex complete data models???**).

This note shows that convergence diagnostics $\widehat{R}$ and $AC$ may diagnose convergence of multiple imputation algorithms, but their performance differs from conventional applications to iterative algorithmic procedures. (**nope! it shows that MICE can lead to correct outcomes when they have not converged according to two common conv diags. This may be due to the measures (e.g., assumption of overdisp) or due to the Qs (lm reg coeff, not higher dimensional/more complex RQs). Add what %miss has to do with it.**)

$\widehat{R}$ and autocorrelation indicate that algorithmic convergence may only be reached after twenty or even forty iterations, while unbiased, confidence valid estimates may be obtained with as little as four iterations. These results are in agreement with the simulation hypothesis: $\widehat{R}$ over-estimates the severity of non-convergence when applied to MI procedures. %This may be due to the quantity of scientific interest chosen. More `complicated' $Q$s (e.g., higher-order effects or variance components) might show bias, under- or over-coverage at higher $T$.

According to this simulation study, the recently proposed threshold of $\widehat{R}<1.01$ may be too stringent for MI algorithms. (**This is only one of the goals: to give applied researchers a diagnostic to indicate that they should keep iterating. The other is the default in mice and other software packages, and yet another is... i forgot**) Under the relatively easy missing data problem of the current study, the threshold was not reached. The other extreme of the $\widehat{R}$-thresholds, the conventionally acceptable $\widehat{R} <1.2$, may be too lenient for MI procedures. Applying this threshold to the current data, lead to falsely diagnosing convergence at $T = 3$ (**because it goes up after, not because it is not converged enough**). It appears that the widely used threshold of $\widehat{R} < 1.1$ suits MI algorithms the best. We might, however, also formulate a new threshold, specifically for the evaluation of MI algorithms. The current study suggests that $\widehat{R} < 1.05$ may be implemented, since that is the level at which the $\widehat{R}$ stabilize (around $T = 20$) (**Not necessary for this Q, but maybe for more complicated Qs**).

The negative $AC$-values obtained in this study show no threat of non-stationarity. However, the initial dip in $AC$-values **(which disappeared!)** may have implications for the default number of iterations in `mice` (`maxit = 5`). Terminating the algorithm at $T=5$ may not be the most appropriate, since this lead to the worst convergence (**nope, only for Rh, not $AC$**), as indicated by $\widehat{R}$ and $AC$. Under the current specifications, $T>20$ would be more appropriate.

Further research is needed to investigate their performance under clear violation of convergence, e.g. dependency between predictors (predictors with very high correlations). Until then, we have only shown that the convergence diagnostics can diagnose non-convergence of MI algorithms that trend towards a converged state. Also for future research, look at developing a convergence diagnostic for substantive models, and implement a Wald test for $AC = 0$.

## Recommendations for empirical researchers

For empirical researchers: 1) Check trace plots for pathological non convergence and adjust imputation model if necessary. 2) use $\widehat{R}$ wait 1.1 ash threshold and autocorrelation with **?** as threshold. Keep iterating until these thresholds are reached. 3) Do not use the R function ACF. Instead, compute autocorrelations manually **(see e.g., [GitHub link])**. 4) Track your own scalar summary of interest. This is somewhat advanced but explained in Van Buren 2018. Compute $\widehat{R}$ and autocorrelation values for this scalar summary. 5) **Something** about the novel $\theta$ that is 'substantive model-independent'.

## Recommendations for future research

**Pmm, M(N)AR, Empirical data, Vignette, ShinyMICE, AC across imputations, not iterations (new convergence diagnostic unique for iterative imputation??).**

**If I don't include it in my study, also: Significance of $AC$ values.**

# Grading

Scientific contribution
(clear research question, convincing argumentation that the study adds to the scientific literature)

Theory
(use of theory, clear hypotheses, relevant literature)

Method
(correct data collection/handling, measures, choice and description of method of analysis)

Results
(correct interpretation of findings, clear presentation)
Discussion and conclusions
(summary, broader implications, limitations, future research)

Written presentation
(writing, structure, tables and figures, references)

<!-- # References -->