### 2020-04-30 3.Thesis/1.SimulationStudy/Writeup/manuscript.tex

\documentclass[Royal,times,sageh]{sagej}

\usepackage{moreverb,url,natbib, multirow, tabularx}
\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

\begin{document}

\title{Missing the Point: Non-Convergence in Iterative Imputation Algorithms}

\runninghead{Oberman}

\author{H. I. Oberman\affilnum{1}}

\affiliation{\affilnum{1}{Department of Methodology and Statistics, Utrecht University, Utrecht,
The Netherlands}}

\corrauth{Hanne Oberman, Sjoerd Groenman building, Utrecht Science Park, Utrecht,
The Netherlands.}

\email{\href{mailto:h.i.oberman@uu.nl}{\nolinkurl{h.i.oberman@uu.nl}}}

\begin{abstract}
Iterative imputation is a popular tool to accommodate missing data.
While it is widely accepted that valid inferences can be obtained with
this technique, these inferences all rely on algorithmic convergence.
There is no consensus on how to evaluate the convergence properties of
the method. This paper provides insight into identifying non-convergence
of iterative impuation algorithms.
\end{abstract}

\keywords{missing data, iterative imputation, non-convergence, mice}

\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Anyone who analyzes person-data may run into a missing data problem.
Missing data is not only ubiquitous, but treating it can also be
tedious. If a dataset contains just one incomplete observation,
statistical inferences are undefined and will not produce any results.
To circumvent this, many statistical packages employ list-wise deletion
by default (i.e., ignoring incomplete observations). Unfortunately, this
\emph{ad hoc} solution may yield wildly invalid results \citep{buur18}.
An alternative is to \emph{impute} (i.e., fill in) the missing values in
the incomplete observations. Subsequently, statistical inferences can be
performed on the completed dataset. By repeating this process several
times, a distribution of plausible results may be obtained, which
reflects the uncertainty in the data due to missingness. This technique
is known as `multiple imputation' \citep[MI;][]{rubin76}. MI has proven
to be a powerful technique to yield unbiased and confidence valid
estimates of the true---but missing---data inference under many
circumstances \citep{buur18}.

Figure \ref{fig:diagram} provides an overview of the steps involved with
MI---from incomplete data, to \(M\) multiply imputed datasets, to \(M\)
estimated quantities of interest \(\hat{Q}\)s, to a single pooled
estimate \(\bar{Q}\). Missing data in dataset \(y\) is imputed \(M\)
times. The imputed data \(y_{imp}\) is combined with the observed data
\(y_{obs}\) to create \(M\) completed datasets. On each completed
dataset, the analysis of scientific interest is performed. The quantity
of scientific interest (e.g., a regression coefficient) is denoted with
\(Q\). Since \(Q\) is estimated on each completed dataset, \(M\)
separate \(\hat{Q}\)-values are obtained. These \(M\) values are
combined into a single pooled estimate \(\bar{Q}\).

\begin{figure}

{\centering \includegraphics[width=\linewidth]{./images/diagram}

}

\caption{Scheme of the main steps in multiple imputation.}\label{fig:diagram}
\end{figure}

A popular method to obtain imputations is to use the `Multiple
Imputation by Chained Equations' algorithm, shorthand
`MICE'\citep{mice}. MICE is an iterative algorithmic procedure to draw
imputations from the posterior predictive distribution of the missing
values. This introduces a potential threat to the validity of the
imputations: What if the algorithm has not converged? Are the
implications then to be trusted? And can we rely on the inference
obtained on the completed data? These are all open questions, because
the convergence properties of iterative imputation algorithms have not
been systematically studied \citep{buur18}. Moreover, there is no
scientific consensus on how to evaluate convergence of MI algorithms
\citep{taka17}. Some default MICE techniques (e.g., `predictive mean
modeling') might not yield converged states at all \citep{murr18}.
Therefore, algorithmic convergence should be monitored carefully.

Currently, the recommended practice for evaluating convergence in
iterative imputation algorithms is to visually inspect imputations for
signs of non-convergence. \textbf{Add that this refers to chain means
and variances already? It's insufficient that these are univariate, so
multivariate state space of the algorithm may not be converged when
univariates are ok. And \citet{buur18}`s suggestion to track
user-defined statistics (e.g., regression coefficient) may be somewhat
advanced for empirical researchers. Moreover, this is not
model-independent, i.e., only apply to one substantive model, while one
of the advantages of MI is that missing data problem and scientific
problem are split. Then we can say 'we propose a novel\ldots{}' and
integrate this: Note that usually we only evaluate the convergence of
univariate scalar summaries (e.g., chain means or variances). With these
we cannot diagnose convergence of multivariable statistics (i.e.,
relations between scalar summaries). Van Buuren \citeyearpar[\(\S\)
4.5.2]{buur18} proposed to implement multivariable evaluation of the
MICE algorithm through eigenvalue decomposition building on the work of
\citet{mack03}. Eigenvalues of a covariance matrix are a measure of the
data's covariance.}.

This method is insufficient on two counts: 1) it may be challenging to
the untrained eye, and 2) only severely pathological cases of
non-convergence may be diagnosed \citep[\(\S\) 6.5.2]{buur18}.
Therefore, a quantitative, diagnostic evaluation of convergence would be
preferred. Yet, monitoring convergence of iterative imputation
algorithms diagnostically is challenging. Iterative imputation
algorithms such as MICE are Markov chain Monte Carlo (MCMC) methods. In
MCMC methods, convergence is not from a scalar to a point but from one
distribution to another. The values generated by the algorithm (e.g.,
imputed values) will vary even after convergence. Therefore, the aim of
convergence diagnostics for MCMC methods is not to establish the point
at which convergence is reached, but to monitor signs of non-convergence
\citep{hoff09}. Several of such diagnostics exist for MCMC methods, but
it is not known whether these are appropriate for iterative imputation
algorithms.

In this paper, we investigate how non-convergence in iterative
imputation algorithms may be diagnosed, how well these methods perform,
and at which point convergence may safely be assumed. For reasons of
brevity, we only focus on the iterative imputation algorithm implemented
in the popular \texttt{mice} package \citep{mice} in \texttt{R}
\citep{R}. The convergence properties of the MICE algorithm are
investigated through model-based simulation. The results of this
simulation study are guidelines for assessing convergence of MI
algorithms, which will aid applied researchers in drawing valid
inference from incomplete datasets.

\textbf{Include sub-questions!} How can non-convergence be identified
diagnostically? Are common MCMC non-convergence diagnostics appropriate
for MICE? And if so, which threshold should be used to diagnose
non-convergence? How many iterations are sufficient/needed to be able to
diagnose non-convergence? Are the default number of iterations
sufficient (i.e., 5 in mice, 10 in SPSS and Stata, 30 in mi)?
\textbf{How severe is it when the algorithm has not converged? And what
are guidelines for practice? Can the parameter of interest, estimand
\(Q\), be correct when the algorithm is not (yet) converged, and vice
versa?} \emph{(Maybe add these too? What are the effects of continued
iterations on estimates, predictions and inferences? Do the answers
differ with varying missingness proportions? That is, we vary the nr of
iterations and the missingness proportion because we assume that the
algorithm has not reached convergence at \(t=1\), and performs worse
with moderate to high missingness (not so much with little or a LOT of
missingness).)}

\hypertarget{some-notation}{%
\subsection{Some notation}\label{some-notation}}

Let \(y\) denote an \(n \times k\) matrix containing the data values on
\(k\) variables for all \(n\) units in a sample. The data value of unit
\(i\) (\(i = 1, 2, \dots, n\)) on variable \(j\)
(\(j = 1, 2, \dots, k\)) may be either observed or missing. \textbf{The
number of units \(i\) with at least one missing data value can be
divided by the total number of units \(n\) to obtain the `missingness
proportion' \(p_{mis}\).} The collection of observed data values in
\(y\) is denoted by \(y_{obs}\); the missing part of \(y\) is referred
to as \(y_{mis}\). For each datapoint in \(y_{mis}\), we sample
\(M \times T\) times plausible values, where \(M\) is the number of
imputations (\(m = 1, 2, \dots, M\)) and \(T\) is the number of
iterations (\(t = 1, 2, \dots, T\)). The collection of samples between
the initial value (at \(t=1\)) and the final imputed value (at \(t=T\))
will be referred to as an `imputation chain'. \textbf{Add: \(\theta\)s
are scalar summaries of interest in the iterative algorithm (e.g., chain
means; the average of the imputed values in each imputation chain). }

\hypertarget{identifying-non-convergence}{%
\section{Identifying
non-convergence}\label{identifying-non-convergence}}

The goal of imputation through e.g.~FCS/MICE is to pbtain a multivariate
distribution by iterating over the sequence of univariate imputations.
Traditionally, the inspection of the multivariate convergence is done by
visually inspecting the univariate chain means and variances over the
iterations, or by tracking the Rhat and autocorrelation. The downside to
these approaches is that they either focus on the univariate state
space, or primarily track the change over the iterations of a
multivariate outcome conform the scientific model of interest. Focusing
on the convergence of these outcome parameters may influence this
procedure in the sense that the model of evaluation favors the model of
interest. Ideally, one would like to evaluate a model-independent
parameter that summarizes the multivariate nature of the data. We
propose eigenvalue as such a parameter, because it\ldots.

Mixing and stationarity have histroically been inspected visually, by
evaluating traceplots of scalar summaries of interest (\(\theta\)s;
e.g., chain means and chain variances). As \citet{buur18} describes,
users can also specify a model-specific scalar summary (e.g., a
regression coefficient). A user-specified scalar summary, however, is
not universal to all complete data problems. Therefore, as inspired by
\citep{mack03}, we propose a \(\theta\) that summarizes the multivariate
state space of the algorithm. Namely, the first eigenvalue of the
variance-covariance matrix of the \(M\) completed datasets. \textbf{The
first eigenvalue has the appealing property that is not dependent on the
substantive model of interest.}

\textbf{Since convergence of iterative imputation algorithms is in
distribution, there is not a unique point at which the algorithm reaches
a converged state. We can only evaluate scalar summaries of the
multivariate state space of the algorithm, \(\theta\)s. The recommended
scalar summaries to evaluate are chain means and chain variances.
Additionally, researchers may ``monitor some statistic of scientific
interest'' \citep[\(\S\) 6.5.2]{buur18}. Because of the issues mentioned
above, we propose a novel \(\theta\) to monitor: the first eigenvalue of
the variance-covariance matrix of the completed data. Let
\(\lambda_1 \geq \lambda_2 \geq ... \geq \lambda_j\) be the eigenvalues
of \(\Sigma\) in each of the \(M\) completed datasets
\({y_{obs}, y_{imp}}\). \(\lambda_1\) is measure that summarises the
covariances in the completed datasets.}

There are two requirements for convergence of iterative algorithms:
mixing and stationarity \citep{gelm13}. Without mixing, imputation
chains do not intermingle nicely, \textbf{indicating that \ldots{}}.
Without stationarity, there is trending within imputation chains, which
implies that further iterations would yiled a different set of
imputations.

To illustrate what non-convergence looks like in MI, we reproduce the
example from van Buuren \citeyearpar[\(\S\) 6.5.2]{buur18}. Figure
\ref{fig:non-conv} shows the traceplot for one of the variables in the
example. we see the average of the imputed values for a variable \(j\)
in \(y_{imp}\). The average imputed values in the left hand plot are
about a magnitude 2 lower than the right hand plot. So non-convergence
has a impact on the average imputed value for \(j\) in \(y_{imp, m}\).
This difference (presumably) translates to bias in the pooled estimate
\(\bar{\mu}_j\) as well. And maybe also to higher order statistics,
e.g.~variances, covariances, etc. Therefore, it's important to converge.
\textbf{Explain what we see, namely example by \citet{buur18}
reproduced, showing the traceplots of chain means for some variable. The
first plot is typical convergence of MICE, the second is pathological
non-convergence because of a mis-specified imputation model. Each line
is an imputation. In the first plot, the chains intermingle nicely and
there is little to no trending. In the second plot, there is a lot of
trending and some chains do not intermingle. Importantly, the chain
means at the last iteration (the imputed value per \(m\)) are very
different between the two plots. The algorithm with the mis-specified
model yields imputed values that are on average a magnitude two larger
than those of the typically converged algorithm. This shows the
importance of reaching converged states in iterative imputation
algorithms.}

\hypertarget{methods-under-evaluation}{%
\subsection{Methods under evaluation}\label{methods-under-evaluation}}

Non-stationarity within chains may be diagnosed with e.g.,
autocorrelation \citep[\(AC\);][]{scha97, gelm13}, numeric standard
error \citep[`MC error';][]{gewe92}, or Raftery and Lewis's
\citeyearpar{raft91} procedure to determine the effect of trending on
the precision of estimates. A widely used diagnostic to monitor mixing
between chains is the potential scale reduction factor \(\widehat{R}\)
\citep[`Gelman-Rubin statistic';][]{gelm92}. With a recently proposed
adaptation, \(\widehat{R}\) might also serve to diagnose
non-stationarity, but this has not yet been thoroughly investigated
\citep{veht19}. Therefore, use we \(\widehat{R}\) and \(AC\) to evaluate
mixing and stationarity separately, as recommended by e.g.,
\citet{cowl96}.

\begin{figure}

{\centering \includegraphics[width=.49\linewidth]{manuscript_files/figure-latex/non-conv-1} \includegraphics[width=.49\linewidth]{manuscript_files/figure-latex/non-conv-2}

}

\caption{Typical convergence versus pathological non-convergence. Please note that this is the same data with a different imputation model, leading to different imputations (see the y axis!). Wel min of meer dezelfde origin op iteration 1.}\label{fig:non-conv}
\end{figure}

\hypertarget{potential-scale-reduction-factor}{%
\subsubsection{Potential scale reduction
factor}\label{potential-scale-reduction-factor}}

An updated version of \(\widehat{R}\) has been proposed by
\citet{veht19} (p.~5). \textbf{This version may be suitable for
iterative imputation.} Let \(M\) be the total number of chains, \(T\)
the number of iterations per chain, and \(\theta\) the scalar summary of
interest (e.g., chain mean or chain variance). For each chain
(\(m = 1, 2, \dots, M\)), we estimate the variance of \(\theta\), and
average these to obtain within-chain variance \(W\).

\begin{align*}
W&=\frac{1}{M} \sum_{m=1}^{M} s_{j}^{2}, \text { where } s_{m}^{2}=\frac{1}{T-1} \sum_{t=1}^{T}\left(\theta^{(t m)}-\bar{\theta}^{(\cdot m)}\right)^{2}. 
\end{align*}

We then estimate between-chain variance \(B\) as the variance of the
collection of average \(\theta\) per chain.

\begin{align*}
B&=\frac{T}{M-1} \sum_{m=1}^{M}\left(\bar{\theta}^{(\cdot m)}-\bar{\theta}^{(\cdot \cdot)}\right)^{2}, \text { where } \bar{\theta}^{(\cdot m)}=\frac{1}{T} \sum_{t=1}^{T} \theta^{(t m)} \text{, } \bar{\theta}^{(\cdot \cdot)}=\frac{1}{M} \sum_{m=1}^{M} \bar{\theta}^{(\cdot m)}. 
\end{align*}

From the between- and within-chain variances we compute a weighted
average, \(\widehat{\operatorname{var}}^{+}\), which over-estimates the
total variance of \(\theta\) \textbf{remove or explain why, or leave
in}. \(\widehat{R}\) is then obtained as a ratio between the
over-estimated total variance and the within-chain variance:

\begin{equation*}
\widehat{R}=\sqrt{\frac{\widehat{\operatorname{var}}^{+}(\theta | y)}{W}},
\text{ where } \widehat{\operatorname{var}}^{+}(\theta | y)=\frac{N-1}{N} W+\frac{1}{N} B.
\end{equation*}

We can interpret \(\widehat{R}\) as potential scale reduction factor
since it indicates by how much the variance of \(\theta\) could be
shrunken down if an infinite number of iterations per chain would be run
\citep{gelm92}. This interpretation assumes that chains are
`over-dispersed' at \(t=1\), and reach convergence as \(T \to \infty\).
Over-dispersion implies that the initial values of the chains are `far
away' from the target distribution and each other. When all chains
sample independent of their initial values, the mixing component of
convergence is satisfied, and \(\widehat{R}\)-values will be close to
one. High \(\widehat{R}\)-values thus indicate non-convergence. The
conventionally acceptable threshold for convergence was
\(\widehat{R} < 1.2\) \citep{gelm92}. More recently, \citet{veht19}
proposed a more stringent threshold of \(\widehat{R} < 1.01\).

\hypertarget{autocorrelation}{%
\subsubsection{Autocorrelation}\label{autocorrelation}}

Following the same notation, we define autocorrelation as the
correlation between two subsequent \(\theta\)-values within the same
chain \citep[p.~147]{lync07}. In this study, we only consider \(AC\) at
lag 1, i.e., the correlation between the \(t^{th}\) and \(t+1^{th}\)
iteration of the same chain.

\begin{equation*}
AC = \left( \frac{T}{T-1} \right) \frac{\sum_{t=1}^{T-1}(\theta_t - \bar{\theta}^{(\cdot m)})(\theta_{t+1} - \bar{\theta}^{(\cdot m)})}{\sum_{t=1}^{T}(\theta_t - \bar{\theta}^{(\cdot m)})^2}.
\end{equation*}

We can interpret \(AC\)-values as a measure of stationarity. If
\(AC\)-values are close to zero, there is no dependence between
subsequent samples within imputation chains. Negative \(AC\)-values
indicate divergence within imputation chains. \textbf{Subsequent sampled
values within each imputation chain are less alike.} Positive
\(AC\)-values indicate recurrence. If \(\theta\)-values of subsequent
iterations are similar, trending may occur. Negative \(AC\)-values show
no threat to the stationarity component of convergence. On the contrary
even---negative \(AC\)-values indicate that \(\theta\)-values of
subsequent iterations diverge from one another, which may increase the
variance of \(\theta\) and speed up convergence. As convergence
diagnostic, the interest is therefore in positive \(AC\)-values.
\textbf{Maybe remove:} Moreover, the magnitude of \(AC\)-values may be
evaluated statistically, but that is outside of our scope.

\hypertarget{in-practice}{%
\subsection{In practice}\label{in-practice}}

\textbf{Before using them in the simulation, the two methods show at
least be able to distinguish between the two algorithms plotted in
Figure 2. From visual inspection we know that the typical convergence
has some signs of non-mixing around iteration 2, and little trending. In
the non-convergence situation, there is a lot of trending upto iteration
6, after which the chains reach a somewhat more stationary state. Mixing
in this situation gets worse from the first iteration onward, and
gradually gets better around iteration 7. To assess whether
\(\widehat{R}\) and \(AC\) may be appropriate non-convergence
identifiers for iterative imputation algorithms, the following condition
must hold: the methods indicate worse performance for the mis-specified
model with pathological non-convergence (i.e., higher \(\widehat{R}\)-
and \(AC\)-values than the typical performance). And additionally the
methods should reflect the increasing convergence in the typical
convergence situation as the number of iterations goes up.}

\textbf{In Figure \ref{fig:diagnostics}A, the chain means from Figure
\ref{fig:non-conv} are plotted again, now together---as are the
non-convergence diagnostics. Panel B shows \(\widehat{R}\) as computed
by implementing \citet{veht19} 's recommendations. As required,
\(\widehat{R}\) indicates less signs of non-convergence as the number of
iterations goes up }in the typical convergence situation\textbf{. The
superior performance of the typical convergence over the pathological
non-convergence is less prominent, even flipped for \(t<4\). Panel C
displays the \(AC\) as computed with the R function
\texttt{stats::acf()}. When we look at this panel, we conclude something
weird. The \(AC\)-values indicate equal performance (up-to \(t=5\)) for
the typical convergence and the pathological non-convergence, while
there is obvious trending in the latter. Moreover, the best convergence
(as indicated by the lowest \(AC\)-value) is observed at \(t=2\), but
looking at the chain means in panel A, there should be some signs of
trending up-to iteration number seven. After consulting the
documentation on stats::acf() \citep{R}, we conclude that this
implementation of \(AC\) is not suitable for iterative imputation
algorithms. There is a correction factor for a mathematical shortcut
that works in the limit. According to \citet{box15}, the function works
when the number of iterations \textgreater{} 50. The default number of
iterations in iterative imputation, however, is often much lower.
Therefore, we compute \(AC\) manually, see panel D. The AC values in
this plot do meet the requirements (\ldots) and will therefore be used
in the simulation study.}

\textbf{acf is niet bedoeld om weinig iteraties te\ldots{} niet een fout
in acf, alleen hier niet van toepassing}

We assume that complete convergence, defined as \(\widehat{R} = 1\) and
\(AC = 0\), will not be observed. Because even in the most converged
state, the algorithm will show some signs of non-mixing and
non-stationarity. Upon \emph{sufficient} convergence, the imputation
chains will intermingle such that the only difference between the chains
is caused by the randomness induced by the algorithm
(\(\widehat{R} > 1\)), and there may be some dependency between
subsequent iterations of imputation chains (\(AC > 0\)).

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/diagnostics-1}

}

\caption{Convergence diagnostics applied on the imputation algorithms of Figure 2.}\label{fig:diagnostics}
\end{figure}

\hypertarget{simulation-study}{%
\section{Simulation study}\label{simulation-study}}

The aim of the simulation study is to evaluate the impact of inducing
non-convergence in the MICE algorithm on several quantities of
scientific interest \(Q\). Subsequently we will evaluate how well
\(\widehat{R}\) and \(AC\) perform in identifying the effects of
non-convergence. And finally, we will formulate an informed advice on
the requirements to safely assume \emph{sufficient} convergence in
practice. That is, we assume that the algorithm is \emph{sufficiently}
converged when each estimate \(\bar{Q}\) is an unbiased and confidence
valid estimate of the corresponding estimand \(Q\).

Non-convergence will be induced by: 1) increasing the missingness
proportion \(p_{mis}\) in dataset \(y\) incrementally, and 2)
terminating the imputation algorithm at a varying number of iterations
\(T\). The first set of simulation conditions---the missingness
proportions---is chosen to reflect the difficulty of the missingness
problem. The underlying assumption is that low missingness proportions
lead to quick algorithmic convergence, since there is a lot of
information in the observed data. Higher missingness proportions should
yield slower convergence. Unless, however, the fraction of missing
information is so high that the random component in the imputation
algorithm outweighs the information in the observed data. Then,
convergence to a stable but highly variable state may be reached
instanteneously. We assume that the incremental missingness proportions
in our study will result in a corresponding increase in signs of
non-convergence.

The assumption inherent to the second set of simulation conditions---the
number of iterations---is that terminating the imputation algorithm too
early causes non-convergence. Generally, the algorithm will not reach
convergence if \(T=1\), because the imputed values in the first
iteration (at \(t=1\)) depend on the initial values of the algorithm
(which are sampled randomly from the set of observed datapoints). As the
number of iterations increases, the imputation chains should become
independent of the initial values, until the point at which adding an
extra iteration does not lead to a more converged state. We assume that
we can induce non-convergence at least in conditions where \(T\) is
smaller than the default number of iterations in \texttt{mice}, \(T=5\).

\hypertarget{hypotheses}{%
\subsection{Hypotheses}\label{hypotheses}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  We expect that simulation conditions with a high missingness
  proportion \(p_{mis}\) and a low number of iterations \(T\) will
  result in biased, invalid estimates of the quantities of scientific
  interest, \(Q\)s.
\item
  We hypothesize that \(\widehat{R}\) and \(AC\) will correctly identify
  signs of non-convergence in those simulation conditions where the
  \(\bar{Q}\)s are \emph{not} unbiased and confidence valid estimates of
  the \(Q\)s.
\item
  We hypothesize that the recommended thresholds to diagnose
  non-convergence with \(\widehat{R}\) (\(\widehat{R} > 1.2\),
  \(\widehat{R} > 1.1\), and \(\widehat{R} > 1.01\)) may be too
  stringent for iterative imputation applications. In an empirical
  study, where \(\widehat{R}\) was used to inform the required
  imputation chain length, it took as many as 50 iterations to overcome
  the conventional non-convergence threshold \(\widehat{R}>1.2\). Yet,
  scientific estimates were insensitive to continued iteration after
  \(t>5\) \citep{lace07}. We therefore suspect that \(\widehat{R}\) may
  over-estimate signs of non-convergence in iterative imputation
  algorithms. In contrast to this, it may occur that the signs of
  non-convergence are under-estimated by \(\widehat{R}\), in exceptional
  cases where the initial values of the algorithm are not appropriately
  over-dispersed \citep[p.~437]{broo98}. In e.g.~\texttt{mice}, initial
  values are chosen randomly from the observed data, hence we cannot be
  certain of over-dispersion in the initial values. In practice, we do
  not expect this to cause problems for identifying non-convergence with
  \(\widehat{R}\).
\item
  We expect that high \(AC\) values are implausible in iterative
  imputation algorithms with typical convergence. That is, after only a
  few iterations, the randomness induced by the algorithm will
  effectively mitigate the risk of dependency within chains.
\end{enumerate}

\hypertarget{set-up}{%
\subsection{Set-up}\label{set-up}}

Convergence of the MICE algorithm is investigated through model-based
simulation in \texttt{R} \citep[version 3.6.3;][]{R}. The simulation
set-up is summarized in the pseudo-code below. The complete \texttt{R}
script of the simulation study is available from
\href{https://github.com/gerkovink/shinyMice/tree/master/3.Thesis/1.SimulationStudy}{github.com/gerkovink/shinyMice}.

\begin{verbatim}
# pseudo-code of simulation 
1. simulate data 
for (number of simulation runs from 1 to 1000)
 for (missingness proportions 5%, 25%, 50%, 75% and 95%)
  2. create missingness
  for (number of iterations from 1 to 100)
   3. impute missingness
   4. perform analysis of scientific interest
   5. compute non-convergence diagnostics 
   6. pool results across imputations
   7. compute performance measures
 8. combine outcomes of all missingness proportions
9. aggregate outcomes across simulation runs 
\end{verbatim}

\hypertarget{data-generating-mechanism.}{%
\subsubsection{Data-generating
mechanism.}\label{data-generating-mechanism.}}

In this study, sampling variance is not of interest. Therefore, a single
complete dataset may serve as comparative truth in all simulation
repetitions \citep{vink14}. The data-generating mechanism is a
multivariate normal distribution, representing person-data on three
predictor variables in a multiple linear regression problem
\textbf{(from an unspecified social scientific field of study)}. Let the
predictor space be defined as \[
\begin{aligned}
\begin{pmatrix}X_1\\
X_2\\
X_3
\end{pmatrix} \sim \mathcal{N}
\begin{bmatrix}
\begin{pmatrix}
12\\
3\\
0.5
\end{pmatrix}\!\!,
\begin{pmatrix}
4 & 4 & 1.8 & 0\\
4 & 16 & 4.8 & 0\\
1.8 & 4.8 & 9 & 0
\end{pmatrix}
\end{bmatrix}\!\!\text{.}\\[2\jot]
\end{aligned}
\] A finite population of \(N=1000\) is simulated using the
\texttt{mvtnorm} package \citep{mvtnorm}. Subsequently, a fourth
variable is constructed as outcome \(Y\). For each unit
\(i = 1, 2,..., N\), let \[
Y_i = 1 + 2X_{1i} +.5X_{2i} - X_{3i} + \epsilon_i ,
\] where \(\epsilon \sim \mathcal{N}(0, 100)\). From the complete set,
we obtain the true values of each quantity of scientific interest \(Q\).

\hypertarget{scientific-estimands.}{%
\subsubsection{Scientific estimands.}\label{scientific-estimands.}}

We consider four types of \(Q\)s that are often of interest in empirical
research. Namely, two descriptive statistics: the mean \(\mu_j\) and
standard deviation \(\sigma_j\) of each variable
\(j = Y, X_1, X_2, X_3\). We also consider the regression coefficients
of the predictors: \(\beta_1\), \(\beta_2\), and \(\beta_3\) in
regression equation
\[Y' = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \beta_3 X_3,\] where \(Y'\)
is the expected value of the outcome. And finally, the proportion of
variance explained by the set of predictors: coefficient of
determination \(r^2 \times 100\) \textbf{(note that lower case \(r\) is
used to avoid confusion with non-convergence diagnostic
\(\widehat{R}\))}. The true values of these scientific estimands are
calculated with the R functions \texttt{base::mean()},
\texttt{stats::sd()}, and \texttt{stats::lm()} \citep{R}. Each of these
\(Q\)s is estimated with a corresponding \(\bar{Q}\), after the complete
dataset is first \emph{amputed} (with a varying missingness proportion
\(p_{mis}\)) and subsequently imputed (with a varying number of
iterations \(T\)).

\hypertarget{methods.}{%
\subsubsection{Methods.}\label{methods.}}

The complete dataset is amputed with the \texttt{mice} package
\citep[function \texttt{mice::ampute()};][]{mice}. All missingness is
multivariate, and conform a `missing completely at random' missingness
mechanism \citep{rubin87}. I.e., the probability to be missing is the
same for all \(N \times k\) cells in \(y\), conditional on the
missingness proportion (\(p_{mis} =.05,.25,.5,.75,.95\)).

Missing datapoints in \(y\) are imputed with the \texttt{mice()}
function \citep{mice}. All imputation procedures are performed with
Bayesian linear regression imputation, and five imputation chains
(\(M=5\)). The number of iterations varies between simulation conditions
(\(T=1,2,...,100\)).

After each run of the \texttt{mice} algorithm, the function
\texttt{mice::complete()} is used to combine the imputed data
\(y_{imp,m}\) with the observed data \(y_{obs}\), resulting in five
completed datasets \{\(y_{obs}, y_{imp, m}\)\}, where
\(m = 1, 2, ..., 5\).

\begin{itemize}
\item
  on the 5 imputations \(y_{imp,m}\): get mean and var within imp chains
  as theta to compute conv diag --\textgreater{} then complete with
  \texttt{mice::complete()}
\item
  on the 5 completed datasets \{\(y_{obs}, y_{imp, m}\)\}: 1) estimate
  Qhats for mu and sigma --\textgreater{} pool as mean()
  --\textgreater{} get Qbars of mu and sigma; 2) compute lambda as theta
  for conv diag; 3) perform lm to get Qhats for betas and r\^{}2
  --\textgreater{} use Qhats of beta as theta to compute conv diag
  --\textgreater{} pool Qhats of beta and r\^{}2 to get Qbars
\end{itemize}

The estimator for each \(Q\) is \(\bar{Q}\)---the pooled aggregate of
the \(\hat{Q}\)s across imputations. To calculate the \(\bar{Q}\)s, we
combine the observed data \(y_{obs}\) and the imputed data for each
imputation \(y_{imp,m}\) with the function \texttt{mice::complete()}.
Descriptive statistics are computed as the average across imputations
for each variable (\(\mu_j\) and \(\sigma_j\), where
\(j = Y, X_1, X_2, X_3\)). Estimated regression coefficients are
obtained with the function \texttt{stats::lm()} for each imputation, and
then pooled conform \citet{vink14}. The coefficient of determination is
estimated for each imputation, and pooled using
\texttt{mice::pool.r.squared()}.

For each run of the imputation algorithm, we compute non-convergence
diagnostics \(\widehat{R}\) and \(AC\). We calculate \(\widehat{R}\) by
implementing Vehtari et al.'s \citeyearpar{veht19} recommendations, and
\(AC\) as the correlation between the \(t^{th}\) and the \((t+1)^{th}\)
iteration. The two methods are applied to four scalar summaries
\(\theta\): chain means (i.e., the mean in each \(y_{imp, m, j}\)),
chain variances (i.e., the variance in each \(y_{imp, m, j}\)), a
statistic of scientific interest \(\hat{Q}\) in each
\{\(y_{obs}, y_{imp, m}\)\}, and the first eigenvalue of the
variance-covariance matrix in each \{\(y_{obs}, y_{imp, m}\)\},
\(\lambda_1\).

The \(\widehat{R}\) and \(AC\) values with chain means as \(\theta\)s
are evaluated against the bias in univariate mean estimates. Similarly,
the performance measure for \(\widehat{R}\) and \(AC\) applied to chain
variances is the bias in estimated standard deviations. To evaluate the
performance of \(\widehat{R}\) and \(AC\) on the eigenvalues, we will
use bias in the coefficient of determination, \(R^2\). Additionally, we
will evaluate the estimated regression coefficients, the coverage rate
of the CI95\% of the regression estimates, and the CI length
(\textbf{see definitions in old Methods}). \textbf{remove the term
performance measure here and add an extra paragraph to define
performance measures as bias in all estimands, and coverage rate and CI
length of regression coefficients.}

\hypertarget{performance-measures.}{%
\subsubsection{Performance measures.}\label{performance-measures.}}

The impact of inducing non-convergence in the iterative imputation
algorithm is assessed by comparing \(\bar{Q}\)s with \(Q\)s. For each
\(Q\), we compute bias as \(\bar{Q} - Q\). For the regression
coefficients \(Q=\beta_{1,2,3}\), we also compute the empirical coverage
rate (CR). CR is defined as the percentage of simulation repetitions in
which the 95\% confidence interval (CI) around \(\bar{Q}\) covers the
true estimand \(Q\). Let
\[\text{CI} = \bar{Q} \pm t_{(M-1)} \times SE_{\bar{Q}},\] where
\(t_{(M-1)}\) is the quantile of a \(t\)-distribution with \(M-1\)
degrees of freedom, and \(SE_{\bar{Q}}\) is the square root of the
pooled variance estimate. Since under-estimating the variance of
\(\bar{Q}\) may yield spurious inferences, we also inspect CI width
(CIW): the difference between the lower and upper bound of the 95\%
confidence interval around \(\bar{Q}\).

The performance of two methods is evaluated against the bias and
confidence validity of the estimated \(Q\)s.

\hypertarget{results}{%
\section{Results}\label{results}}

\textbf{Add to this section: 1) emphasize that the plots are averages
across repetitions, not within MICE; 2) sample effects due to single
complete dataset.}

For reasons of brevity, we only discuss the convergence diagnostics for
the \(Q\) with the worst performance in terms of bias. For the
descriptve statistics, the magnitude of the bias was the largest in
\(Y\) (i.e, \(j = Y\) in \(\mu_j\) and \(\sigma_j\)). For the
\(Q=\beta_j\), \(j=2\) showed the worst perfomance (i.e, the effect of
\(X_1\) on \(Y\).

As expected, conditions with a higher proportion of missingness and/or a
lower number of iterations show more signs of non-convergence, as
indicated by more extreme bias in the estimated \(Q\)s. Roughly speaking
this means that MICE is indeed not converged as \(t=1\), and converges
gradually as \(t\) increases. The point at which an additional iteration
does not lead to an improvement of the estimates depends on the
difficulty of the missingness problem---between XYZ and XYZ. This means
that the algorithm has converged sufficiently (\textbf{under the current
specifications}).

As required for a method to diagnose non-convergence, the values of
\(\widehat{R}\) follow a trend similar to the performance measures.
\(\widehat{R}\) values are generally lower in conditions with a higher
number of iterations, and somewhat higher in conditions with a higher
percentage of missing data. Autocorrelation values were indeed
decreasing with a higher number of iterations.

Evaluation with the performance measures shows that \(\widehat{R}\) and
autocorrelation are conservative: they indicate signs of non-convergence
in conditions where \(\bar{Q}\)s are unbiased and or confidence valid
estimates of \(Q\).

\hypertarget{old-results-section}{%
\subsection{Old Results section}\label{old-results-section}}

\textbf{(Add more info about figure legends and axes.)}

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/mean-1}

}

\caption{Convergence diagnostics chain mean.}\label{fig:mean}
\end{figure}

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/sd-1}

}

\caption{Convergence diagnostics chain variance.}\label{fig:sd}
\end{figure}
\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/est-1}

}

\caption{Convergence diagnostics regression coefficient.}\label{fig:est}
\end{figure}

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/pred-1}

}

\caption{Convergence diagnostics regression coefficient.}\label{fig:pred}
\end{figure}

\hypertarget{univariate-estimates-and-convergence-diagnostics}{%
\subsection{Univariate estimates and convergence
diagnostics}\label{univariate-estimates-and-convergence-diagnostics}}

The bias in the estimates of the variable means shows little to no
difference between simulation conditions. It doesn't seem to matter how
many iterations you use in the mice algorithm, the estimates are
unbiased. Similarly, the bias in the estimated variances is more or less
stable across simulation conditions. These univariate quantities appear
to be unaffected by the number of iterations.

When applied to the imputation chain means, \(\widehat{R}\) indicates
that the mice algorithm does not reach a converged state
(\(\widehat{R}\) = 1) in any of the simulation conditions. Neither is
the most recent recommended threshold reached (\(\widehat{R}\)
\textless{} 1.01). \textbf{This is only for 75\% missingness:} The
conventional \(\widehat{R}\) threshold of 1.2 is reached in simulation
conditions \(T = 3\) and \(T > 6\). (With the default number of
iterations (maxit = 5), this dip in \(\widehat{R}\) values would be
spotted, so it is no problem.) The point at which an extra iteration
does not seem to improve the \(\widehat{R}\) value is around \(T=30\).

Autocorrelations indicate no sign of trending \textbf{(NOT correct
anymore, fluke of the stats::acf() function!!)} within imputation
chains. In most simulation conditions \(AC\) is smaller than or about
equal to zero. Across simulation conditions, however, the
autocorrelation curve does not trend towards 0. Autocorrelation values
plateau off at a value of around.1. This is a small positive
autocorrelation, which would indicate some trending within chains.

The \(\widehat{R}\) and \(AC\) values for the imputation chain variances
show equal trends to the chain means and are therefore not discussed
separately. Taken together, univariate estimates seem robust across
simulation conditions. There is no clear effect of the number of
iterations on the bias in these estimates, while the convergence
diagnostics indicate that the algorithm did not reach a completely
converged state (yet).

\hypertarget{multivariate-estimates}{%
\subsection{Multivariate estimates}\label{multivariate-estimates}}

\textbf{This is only true for 75\% missingness:} There is a clear bias
in the regression estimates in simulation conditions where the number of
iterations is smaller than four. In simulation conditions where
\(T > 5\) there is little to no bias in the estimated regression
coefficients. In most simulation conditions nominal coverage is obtained
(i.e., coverage rates of.95) for the confidence intervals around the
regression coefficients. Conditions with only one or two iterations show
some under-coverage. Since confidence interval width is stable across
conditions, the under-coverage may be attributed to the bias in the
estimated regression coefficients.

\textbf{This is only true for 75\% missingness:} If we look at the
estimated proportion of explained variance in outcome variable Y we see
that the coefficient of determination is underestimate estimated in
conditions where the number of iterations is equal to two or less, and
slightly overestimated in conditions where the number of iterations is
equal to three or more.

In short, we see that the minimum number of iterations required to
obtain unbiased, confidence valid regression estimates is 5. This value,
however, is dependent on the percentage of missing values. E.g., with
95\% of cases having missing data we need at least seven iterations to
obtain unbiased results.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\emph{Proper performance under a `missing completely at random'
missingness mechanism is necessary to demonstrate the appropriateness of
\(\widehat{R}\) and \(AC\) as non-convergence diagnostics. However,
results may not be extrapolated to other missingness mechanisms.}

Upon convergence, \(\widehat{R}=1\) and \(AC=0\), which are unlikely
thresholds for MCMC algorithms, because of its convergence to a
distribution. In practice, non-convergence is usually diagnosed when
\(\widehat{R}\) \textgreater{} 1.2 or 1.1 or even 1.01. And a t-test is
performed to assess whether \(AC\) is significantly different from zero.

As expected, complete convergence (\(\widehat{R}=1\) and \(AC=0\)) is
never observed. But what about the thresholds for practice?
\(\widehat{R}\) \textless1.2 is not stringent enough, because conditions
in which \(\widehat{R}\) is smaller than 1.2 do not all yield unbiased
estimates of the \(Q\)s. Moreover, there is a dip in the \(\widehat{R}\)
values, after which \(\widehat{R}\) \textgreater{} 1.2 again. If the
algorithms would be terminated at the iteration where \(\widehat{R}\)
\textless{} 1.2 occurs for the first time, the increase in
\(\widehat{R}\) values might be missed. The threshold \(\widehat{R}\)
\textless{} 1.1 is somewhat conservative in comparison with the
performance measures. Overall it seems OK. \(\widehat{R}\) \textless{}
1.01 is too stringent compared to the performance measures. And it is
not obtained for any number of iterations (\textbf{but may be necessary
in more complex complete data models???}).

This note shows that convergence diagnostics \(\widehat{R}\) and \(AC\)
may diagnose convergence of multiple imputation algorithms, but their
performance differs from conventional applications to iterative
algorithmic procedures. (\textbf{nope! it shows that MICE can lead to
correct outcomes when they have not converged according to two common
conv diags. This may be due to the measures (e.g., assumption of
overdisp) or due to the Qs (lm reg coeff, not higher dimensional/more
complex RQs). Add what \%miss has to do with it.})

\(\widehat{R}\) and autocorrelation indicate that algorithmic
convergence may only be reached after twenty or even forty iterations,
while unbiased, confidence valid estimates may be obtained with as
little as four iterations. These results are in agreement with the
simulation hypothesis: \(\widehat{R}\) over-estimates the severity of
non-convergence when applied to MI procedures. \%This may be due to the
quantity of scientific interest chosen. More `complicated' \(Q\)s (e.g.,
higher-order effects or variance components) might show bias, under- or
over-coverage at higher \(T\).

According to this simulation study, the recently proposed threshold of
\(\widehat{R}<1.01\) may be too stringent for MI algorithms.
(\textbf{This is only one of the goals: to give applied researchers a
diagnostic to indicate that they should keep iterating. The other is the
default in mice and other software packages, and yet another is\ldots{}
i forgot}) Under the relatively easy missing data problem of the current
study, the threshold was not reached. The other extreme of the
\(\widehat{R}\)-thresholds, the conventionally acceptable
\(\widehat{R} <1.2\), may be too lenient for MI procedures. Applying
this threshold to the current data, lead to falsely diagnosing
convergence at \(T = 3\) (\textbf{because it goes up after, not because
it is not converged enough}). It appears that the widely used threshold
of \(\widehat{R} < 1.1\) suits MI algorithms the best. We might,
however, also formulate a new threshold, specifically for the evaluation
of MI algorithms. The current study suggests that \(\widehat{R} < 1.05\)
may be implemented, since that is the level at which the \(\widehat{R}\)
stabilize (around \(T = 20\)) (\textbf{Not necessary for this Q, but
maybe for more complicated Qs}).

The negative \(AC\)-values obtained in this study show no threat of
non-stationarity. However, the initial dip in \(AC\)-values
\textbf{(which disappeared!)} may have implications for the default
number of iterations in \texttt{mice} (\texttt{maxit\ =\ 5}).
Terminating the algorithm at \(T=5\) may not be the most appropriate,
since this lead to the worst convergence (\textbf{nope, only for Rh, not
\(AC\)}), as indicated by \(\widehat{R}\) and \(AC\). Under the current
specifications, \(T>20\) would be more appropriate.

Further research is needed to investigate their performance under clear
violation of convergence, e.g.~dependency between predictors (predictors
with very high correlations). Until then, we have only shown that the
convergence diagnostics can diagnose non-convergence of MI algorithms
that trend towards a converged state. Also for future research, look at
developing a convergence diagnostic for substantive models, and
implement a Wald test for \(AC = 0\).

\hypertarget{recommendations-for-empirical-researchers}{%
\subsection{Recommendations for empirical
researchers}\label{recommendations-for-empirical-researchers}}

For empirical researchers: 1) Check trace plots for pathological non
convergence and adjust imputation model if necessary. 2) use
\(\widehat{R}\) wait 1.1 ash threshold and autocorrelation with
\textbf{?} as threshold. Keep iterating until these thresholds are
reached. 3) Do not use the R function ACF. Instead, compute
autocorrelations manually \textbf{(see e.g., {[}GitHub link{]})}. 4)
Track your own scalar summary of interest. This is somewhat advanced but
explained in Van Buren 2018. Compute \(\widehat{R}\) and autocorrelation
values for this scalar summary. 5) \textbf{Something} about the novel
\(\theta\) that is `substantive model-independent'.

\hypertarget{recommendations-for-future-research}{%
\subsection{Recommendations for future
research}\label{recommendations-for-future-research}}

\textbf{Pmm, M(N)AR, Empirical data, Vignette, ShinyMICE, AC across
imputations, not iterations (new convergence diagnostic unique for
iterative imputation??).}

\textbf{If I don't include it in my study, also: Significance of \(AC\)
values.}

\hypertarget{grading}{%
\section{Grading}\label{grading}}

Scientific contribution (clear research question, convincing
argumentation that the study adds to the scientific literature)

Theory (use of theory, clear hypotheses, relevant literature)

Method (correct data collection/handling, measures, choice and
description of method of analysis)

Results (correct interpretation of findings, clear presentation)
Discussion and conclusions (summary, broader implications, limitations,
future research)

Written presentation (writing, structure, tables and figures,
references)

\bibliographystyle{sageh}
\bibliography{thesis.bib}

\end{document}