### 2020-05-03 3.Thesis/1.SimulationStudy/Writeup/manuscript.tex

\documentclass[Royal,times,sageh]{sagej}

\usepackage{moreverb,url,natbib, multirow, tabularx}
\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

\begin{document}

\title{Missing the Point: Non-Convergence in Iterative Imputation Algorithms}

\runninghead{Oberman}

\author{H. I. Oberman\affilnum{1}}

\affiliation{\affilnum{1}{Department of Methodology and Statistics, Utrecht University, Utrecht,
The Netherlands}}

\corrauth{Hanne Oberman, Sjoerd Groenman building, Utrecht Science Park, Utrecht,
The Netherlands.}

\email{\href{mailto:h.i.oberman@uu.nl}{\nolinkurl{h.i.oberman@uu.nl}}}

\begin{abstract}
Iterative imputation is a popular tool to accommodate missing data.
While it is widely accepted that valid inferences can be obtained with
this technique, these inferences all rely on algorithmic convergence.
There is no consensus on how to evaluate the convergence properties of
the method. This paper provides insight into identifying non-convergence
of iterative impuation algorithms.
\end{abstract}

\keywords{missing data, iterative imputation, non-convergence, mice}

\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Anyone who analyzes person-data may run into a missing data problem.
Missing data is not only ubiquitous, but treating it can also be
tedious. If a dataset contains just one incomplete observation,
statistical inferences are undefined and will not produce any results.
To circumvent this, many statistical packages employ list-wise deletion
by default (i.e., ignoring incomplete observations). Unfortunately, this
\emph{ad hoc} solution may yield wildly invalid results \citep{buur18}.
An alternative is to \emph{impute} (i.e., fill in) the missing values in
the incomplete observations. Subsequently, statistical inferences can be
performed on the completed dataset. By repeating this process several
times, a distribution of plausible results may be obtained, which
reflects the uncertainty in the data due to missingness. This technique
is known as `multiple imputation' \citep[MI;][]{rubin76}. MI has proven
to be a powerful technique to draw valid inferences from incomplete data
under many circumstances \citep{buur18}.

Figure \ref{fig:diagram} provides an overview of the steps involved with
MI. Missing data \(y_{mis}\) in an incomplete dataset is imputed \(M\)
times. The imputed data \(y_{imp}\) is combined with the observed data
\(y_{obs}\) to create \(M\) completed datasets. On each completed
dataset, an analysis of scientific interest is performed. The quantity
of scientific interest (e.g., a regression coefficient) is denoted with
\(Q\). Since \(Q\) is estimated on each completed dataset, \(M\)
separate \(\hat{Q}\)-values are obtained. These \(M\) values are
combined into a single pooled estimate \(\bar{Q}\). The premise of
multiple imputation is that \(\bar{Q}\) is an unbiased and confidence
valid estimate of the true---but missing---data inference.

\begin{figure}

{\centering \includegraphics[width=\linewidth]{./images/diagram}

}

\caption{Scheme of the main steps in multiple imputation (where $M=3$)---from an incomplete dataset, to $M$ multiply imputed datasets, to $M$ estimated quantities of scientific interest $\hat{Q}$s, to a single pooled estimate $\bar{Q}$.}\label{fig:diagram}
\end{figure}

A popular method to obtain imputations is to use the `Multiple
Imputation by Chained Equations' algorithm, shorthand
`MICE'\citep{mice}. With MICE, imputed values \(y_{imp}\) are drawn from
the posterior predictive distribution of the missing values \(y_{mis}\).
The algorithm is named after the iterative algorithmic procedure by
which imputed values are generated: a multivariate distribution is
obtained by iterating over a sequence of univariate imputations. The
iterative nature of algorithms like MICE introduces a potential threat
to the validity of the imputations: What if the algorithm has not
converged? Are the implications then to be trusted? And can we rely on
the inference obtained on the completed data? These are all open
questions, because the convergence properties of iterative imputation
algorithms have not been systematically studied \citep{buur18}.
Moreover, there is no scientific consensus on how to evaluate
convergence of imputation algorithms \citep{taka17}. Some default MICE
techniques (e.g., `predictive mean modeling') might not yield converged
states at all \citep{murr18}. Therefore, algorithmic convergence should
be monitored carefully.

Currently, the recommended practice for evaulating the convergence of
iterative imputation algorithms is through visual inspection. That is,
the state space of the algorithm is monitored by plotting a scalar
summary \(\theta\) against the iteration number. \textbf{{[}Segway to
WHAT TO MONITOR? Also, this paragraph is more or less the same as the
third paragraph in the second section. Merge or remove one of the
two?{]}} Typically, the \(\theta\)s under evaluation are chain means and
variances (\textbf{define!}). These \(\theta\)s may be insufficient
because they are univariate summaries of the state space of the
algorithm, while MICE is not only concerned with univariate estimates,
but the entire multivariate distribution of \(y_{imp}\). A suggestion by
\citet{buur18} for a multivariate \(\theta\) to monitor is the quantity
of scientific interest \(Q\) (e.g., a regression coefficient).
Implementing this, however, might be somewhat advanced for empirical
researchers. Moreover, this scalar summary is not model-independent,
i.e., it only applies to one model of scientific interest. Yet, one of
the advantages of MI is that the missing data problem and scientific
problem are solved independently. Therefore, this \(\theta\) is not
sufficient either. Van Buuren \citeyearpar[\(\S\) 4.5.2]{buur18} also
proposed multivariable evaluation of the MICE algorithm through
eigenvalue decomposition, building on the work of \citet{mack03}.
Eigenvalues of a variance-covariance matrix are a measure of the data's
total covariance. This would be a model-independent, multivariate scalar
summary to monitor, but it has not yet been implemented.
\textbf{{[}Segway to HOW TO MONITOR?{]}} Monitoring convergence visually
is insufficient on two counts: 1) it may be challenging to the untrained
eye, and 2) only severely pathological cases of non-convergence may be
diagnosed \citep[\(\S\) 6.5.2]{buur18}. Therefore, a quantitative,
diagnostic evaluation of convergence would be preferred---although not
straightforward. Iterative imputation algorithms such as MICE are Markov
chain Monte Carlo (MCMC) methods. In MCMC methods, convergence is not
from a scalar to a point, but from one distribution to another. The
values generated by the algorithm (e.g., imputed values) will vary even
after convergence. Since MCMC algorithms do not reach a unique point at
which convergence is established, diagnostics may only identify signs of
\emph{non}-convergence \citep{hoff09}. Several of such non-convergence
diagnostics exist, but it is not known whether these are appropriate for
iterative imputation algorithms.

In this paper, we investigate how non-convergence in iterative
imputation algorithms may be diagnosed. First, we define which
non-convergence diagnostics will be considered, and evaluate how these
diagnostics could be appropriate for iterative imputation applications.
In the second part of this paper, we will investigate the impact of
inducing non-convergence in iterative imputation algorithms by means of
model-based simulation in \texttt{R} \citep{R}. The aim of the
simulation is to determine whether unbiased, confidence valid inferences
may be obtained under different levels of non-convergence (\textbf{and
at which point convergence may safely be assumed}). Additionally, we
will assess how well different non-convergence diagnostics perform in
identifying signs of non-convergence. The results of this study are
guidelines for identifying non-convergence in iterative imptation
algorithms, which will aid applied researchers in drawing valid
inference from incomplete datasets. \textbf{{[}add more societal
relevance: what is the consequence of non-convergence? e.g., biased
estimates, invalid inference, etc.{]}} For reasons of brevity, we only
focus on the iterative imputation algorithm implemented in the popular
\texttt{mice} package \citep{mice} in \texttt{R} \citep{R}.

\textbf{{[}Include sub-questions and make structure of the paper
clear!{]}} How can non-convergence be identified diagnostically? Are
common MCMC non-convergence diagnostics appropriate for MICE? And if so,
which threshold should be used to diagnose non-convergence? How many
iterations are sufficient/needed to be able to diagnose non-convergence?
Are the default number of iterations sufficient (i.e., 5 in mice, 10 in
SPSS and Stata, 30 in mi)? How severe is it when the algorithm has not
converged? And what are guidelines for practice? Can the parameter of
interest \(Q\) be correct when the algorithm is not (yet) converged, and
vice versa?

\hypertarget{some-notation}{%
\subsection{Some notation}\label{some-notation}}

Let \(y\) denote an \(n \times k\) matrix containing the data values on
\(k\) variables for all \(n\) units in a sample. The data value of unit
\(i\) (\(i = 1, 2, \dots, n\)) on variable \(j\)
(\(j = 1, 2, \dots, k\)) may be either observed or missing. The number
of units \(i\) in dataset \(y\) with at least one missing value, divided
by the total number of units \(n\), is called the missingness proportion
\(p_{mis}\). The collection of observed data values in \(y\) is denoted
by \(y_{obs}\); the missing part of \(y\) is referred to as \(y_{mis}\).
For each datapoint in \(y_{mis}\), we sample \(M \times T\) plausible
values, where \(M\) is the number of imputations
(\(m = 1, 2, \dots, M\)) and \(T\) is the number of iterations
(\(t = 1, 2, \dots, T\)) in the imputation algorithm. The collection of
samples between the initial value (at \(t=1\)) and the final imputed
value (at \(t=T\)) will be referred to as an `imputation chain'.
\textbf{\(\theta\)s are scalar summaries of the algorithm at a certain
iteration (e.g., chain means; the average of the imputed values in each
imputation chain). }

\hypertarget{identifying-non-convergence}{%
\section{Identifying
non-convergence}\label{identifying-non-convergence}}

There are two requirements for convergence of iterative algorithms:
mixing and stationarity \citep{gelm13}. Without mixing, imputation
chains do not intermingle nicely, indicating that the distribution of
imputed values differs across imputations. Chains may be `stuck' at a
local optimum, instead of sampling imputed values from the entire
predictive posterior distribution of the missing values. This may cause
under-estimation of the variance within chains, which results in
spurious, invalid inferences. Without stationarity, there is trending
within imputation chains. Trending implies that further iterations would
yield a different set of imputations. Iterative imputation algorithms
that have not yet reached stationarity, may thus yield biased estimates.

\textbf{{[}Move to INTRO???{]}} To illustrate what non-convergence looks
like in iterative imputation algorithms, we reproduce an example from
van Buuren \citeyearpar[\(\S\) 6.5.2]{buur18}. Figure \ref{fig:non-conv}
displays a subset of the example: a variable that we will call \(j\). We
see two traceplots of the chain means for \(j\)---i.e., the average of
the imputed values for \(j\) in each imputation \(y_{imp,m}\), plotted
against the iteration number. Each line portrays one imputation. The
plot on the lefthand side of the figure shows typical convergence of an
iterative imputation algorithm. The righthand side displays pathological
non-convergence, induced by purposefully mis-specifying the imputation
model. \textbf{{[}Leave this in if it's after mixing and stationarity
are defined, otherwise remove?{]} In the typical convergence situation,
the imputation chains intermingle nicely and there is little to no
trending. In the non-convergence plot, there is a lot of trending and
some chains do not intermingle.} Importantly, the chain means at the
last iteration (the imputed value per \(m\)) are very different between
the two plots. The algorithm with the mis-specified model yields imputed
values that are on average a factor two larger than those of the
typically converged algorithm. That means that non-convergence has a
impact on the mean of \(j\) in \(y_{imp, m}\). This difference
(presumably) translates to bias in the \(M\) completed data estimates
\(\hat{\mu}_{j,m}\), and consequently also to the pooled estimate
\(\bar{\mu}_j\). Non-convergence thus leads to biased estimates. This
shows the importance of reaching converged states in iterative
imputation algorithms.

\begin{figure}

{\centering \includegraphics[width=.49\linewidth]{manuscript_files/figure-latex/non-conv-1} \includegraphics[width=.49\linewidth]{manuscript_files/figure-latex/non-conv-2}

}

\caption{Typical convergence versus pathological non-convergence. Please note that this is the same data with a different imputation model, leading to different imputations (e.g., see the units on the y-axis).}\label{fig:non-conv}
\end{figure}

\textbf{{[}Move to INTRO???{]}} Convergence has historically been
inspected visually, by monitoring the imputation algorithm over
iterations. This is typically done through traceplots. In a traceplot,
the iteraton number is plotted against a certain \(\theta\). \(\theta\)s
are scalar summaries of the state space of the algorithm at a specific
iteration. The recommended scalar summaries to evaluate are chain means
and chain variances. Additionally, researchers may ``monitor some
statistic of scientific interest'' \citep[\(\S\) 6.5.2]{buur18}. As
\citet{buur18} describes, researchers can specify their quantity of
scientific interest \(Q\) (e.g., a regression coefficient) as scalar
summary \(\theta\). Such a user-defined scalar summary, however, may be
somewhat advanced for empirical researchers. And moreover, it is not
universal to all complete data problems. Focusing on the convergence of
outcome parameters may influence the iterative imputation procedure in
the sense that the model of evaluation favors the model of interest. The
downside to these two approaches is that they either focus on the
univariate state space, or primarily track the change over the
iterations of a multivariate outcome conform the scientific model of
interest. Ideally, one would like to evaluate a model-independent
parameter that summarizes the multivariate nature of the data.
Therefore, we propose a \(\theta\) that summarizes the multivariate
state space of the algorithm, but is independent from the model of
scientific interest. We propose \(\lambda_1\) as such a scalar summary.
We define \(\lambda_1\) as the first eigenvalue of the
variance-covariance matrix of the completed data. Let
\(\lambda_1 \geq \lambda_2 \geq ... \geq \lambda_j\) be the eigenvalues
of \(\Sigma\) in each of the \(M\) completed datasets
\({y_{obs}, y_{imp}}\). \(\lambda_1\) is measure that summarises the
covariances in the completed datasets. \textbf{The first eigenvalue has
the appealing property that is not dependent on the substantive model of
interest. Eigenvalue decomposition is inspired by \citet{mack03}.}

\hypertarget{diagnostics-under-evaluation}{%
\subsection{Diagnostics under
evaluation}\label{diagnostics-under-evaluation}}

Non-convergence diagnostics for MCMC algorithms typically identify
problems in either one of the two requirements for converegnce: mixing
or stationarity. Non-stationarity within chains may be diagnosed with
e.g., autocorrelation \citep[\(AC\);][]{scha97, gelm13}, numeric
standard error \citep[`MC error';][]{gewe92}, or Raftery and Lewis's
\citeyearpar{raft91} procedure to determine the effect of trending on
the precision of estimates. A widely used diagnostic to monitor mixing
between chains is the potential scale reduction factor \(\widehat{R}\)
\citep[`Gelman-Rubin statistic';][]{gelm92}. With a recently proposed
adaptation, \(\widehat{R}\) might also serve to diagnose
non-stationarity, but this has not yet been thoroughly investigated
\citep{veht19}. Therefore, we use \(\widehat{R}\) and \(AC\) to evaluate
mixing and stationarity separately, as recommended by e.g.,
\citet{cowl96}.

\hypertarget{potential-scale-reduction-factor}{%
\subsubsection{Potential scale reduction
factor}\label{potential-scale-reduction-factor}}

An updated version of \(\widehat{R}\) has been proposed by Vehtari et
al.~\citeyearpar[p.~5]{veht19}. \textbf{This version may be suitable for
iterative imputation.} Let \(M\) be the total number of chains, \(T\)
the number of iterations per chain, and \(\theta\) the scalar summary of
interest (e.g., chain means). For each chain (\(m = 1, 2, \dots, M\)),
we estimate the variance of \(\theta\), and average these to obtain
within-chain variance \(W\).

\begin{align*}
W&=\frac{1}{M} \sum_{m=1}^{M} s_{j}^{2}, \text { where } s_{m}^{2}=\frac{1}{T-1} \sum_{t=1}^{T}\left(\theta^{(t m)}-\bar{\theta}^{(\cdot m)}\right)^{2}. 
\end{align*}

We then estimate between-chain variance \(B\) as the variance of the
collection of average \(\theta\) per chain.

\begin{align*}
B&=\frac{T}{M-1} \sum_{m=1}^{M}\left(\bar{\theta}^{(\cdot m)}-\bar{\theta}^{(\cdot \cdot)}\right)^{2}, \text { where } \bar{\theta}^{(\cdot m)}=\frac{1}{T} \sum_{t=1}^{T} \theta^{(t m)} \text{, } \bar{\theta}^{(\cdot \cdot)}=\frac{1}{M} \sum_{m=1}^{M} \bar{\theta}^{(\cdot m)}. 
\end{align*}

From the between- and within-chain variances we compute a weighted
average, \(\widehat{\operatorname{var}}^{+}\), which over-estimates the
total variance of \(\theta\) \textbf{remove or explain why, or leave
in}. \(\widehat{R}\) is then obtained as a ratio between the
over-estimated total variance and the within-chain variance:

\begin{equation*}
\widehat{R}=\sqrt{\frac{\widehat{\operatorname{var}}^{+}(\theta | y)}{W}},
\text{ where } \widehat{\operatorname{var}}^{+}(\theta | y)=\frac{N-1}{N} W+\frac{1}{N} B.
\end{equation*}

We can interpret \(\widehat{R}\) as potential scale reduction factor
since it indicates by how much the variance of \(\theta\) could be
shrunken down if an infinite number of iterations per chain would be run
\citep{gelm92}. This interpretation assumes that chains are
`over-dispersed' at \(t=1\), and reach convergence as \(T \to \infty\).
Over-dispersion implies that the initial values of the chains are `far
away' from the target distribution and each other. When all chains
sample independent of their initial values, the mixing component of
convergence is satisfied, and \(\widehat{R}\)-values will be close to
one. High \(\widehat{R}\)-values thus indicate non-convergence. The
conventionally acceptable threshold for convergence was
\(\widehat{R} < 1.2\) \citep{gelm92}. More recently, \citet{veht19}
proposed a more stringent threshold of \(\widehat{R} < 1.01\).

\hypertarget{autocorrelation}{%
\subsubsection{Autocorrelation}\label{autocorrelation}}

Following the same notation, we define autocorrelation as the
correlation between two subsequent \(\theta\)-values within the same
chain \citep[p.~147]{lync07}. In this study, we only consider \(AC\) at
lag 1, i.e., the correlation between the \(t^{th}\) and \(t+1^{th}\)
iteration of the same chain.

\begin{equation*}
AC = \left( \frac{T}{T-1} \right) \frac{\sum_{t=1}^{T-1}(\theta_t - \bar{\theta}^{(\cdot m)})(\theta_{t+1} - \bar{\theta}^{(\cdot m)})}{\sum_{t=1}^{T}(\theta_t - \bar{\theta}^{(\cdot m)})^2}.
\end{equation*}

We can interpret \(AC\)-values as a measure of stationarity. If
\(AC\)-values are close to zero, there is no dependence between
subsequent samples within imputation chains. Negative \(AC\)-values
indicate divergence within imputation chains. \textbf{Subsequent sampled
values within each imputation chain are less alike.} Positive
\(AC\)-values indicate recurrence. If \(\theta\)-values of subsequent
iterations are similar, trending may occur. Negative \(AC\)-values show
no threat to the stationarity component of convergence. On the contrary
even---negative \(AC\)-values indicate that \(\theta\)-values of
subsequent iterations diverge from one another, which may increase the
variance of \(\theta\) and speed up convergence. As convergence
diagnostic, the interest is therefore in positive \(AC\)-values. The
magnitude of \(AC\)-values may be evaluated statistically {[}\textbf{add
reference}{]}.

\hypertarget{thresholds}{%
\subsubsection{Thresholds}\label{thresholds}}

Upon convergence, \(\widehat{R}=1\) and \(AC=0\), which are unlikely
thresholds for MCMC algorithms, because of its convergence to a
distribution. Even in the most converged state, the algorithm will show
some signs of non-mixing and non-stationarity. Upon \emph{sufficient}
convergence, the imputation chains will intermingle such that the only
difference between the chains is caused by the randomness induced by the
algorithm (\(\widehat{R} > 1\)), and there may be some dependency
between subsequent iterations of imputation chains (\(AC > 0\)). In
practice, non-convergence is usually diagnosed when \(\widehat{R}\)
\textgreater{} 1.2 or 1.1 or even 1.01. And a t-test is performed to
assess whether \(AC\) is significantly different from zero.

\hypertarget{in-practice}{%
\subsection{In practice}\label{in-practice}}

Before we evaluate the performance of the non-convergence diagnostics
\(\widehat{R}=1\) and \(AC=0\) quantitatively through simulation, we
apply them to the example of pathological non-convergence by
\citet{buur18} that we reproduced above. We want to be able to draw the
same conclusions as we did from visually inspecting the two algorithms.
But \(\widehat{R}=1\) and \(AC=0\) should at least be able to
distinguish between the two algorithms plotted in Figure
\ref{fig:non-conv}. From visual inspection, we know that the typically
converged algorithm initially portrayed some signs of non-mixing (around
\(t=2\)), but intermigled nicely overall. Additionally, there was very
little trending. The algorithm with pathological non-convergence showed
severe non-mixing, although this gradually improved (from \(t=7\)).
Moreover, there was a lot of trending as well initially (up-to \(t=6\)),
after which the chains reached a somewhat more stationary state. To
assess whether \(\widehat{R}\) and \(AC\) may be appropriate
non-convergence identifiers for iterative imputation algorithms, they
should identify the same trends as the visual inspection. But at least,
the following condition must hold: the methods should indicate worse
performance for the mis-specified model with pathological
non-convergence than the typically converged algorithm (i.e., higher
\(\widehat{R}\)- and \(AC\)-values).

In Figure \ref{fig:diagnostics}A, the chain means from Figure
\ref{fig:non-conv} are plotted again---now together. Panel B shows two
versions of calculating \(AC\)s applied to the \(\theta\)s of panel A:
the default calculation with \texttt{R} function \texttt{stats::acf()}
\citep{R}, and manual calculation as the correlation between \(\theta\)
in iteration \(t\) and \(\theta\) in iteration \(t+1\). Panel C shows
the traditional computation of \(\widehat{R}\), and panel D shows
\(\widehat{R}\) as computed by implementing \citet{veht19} 's
recommendations.

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/diagnostics-1}

}

\caption{Convergence diagnostics applied on the imputation algorithms of Figure 2. $\theta$ = chain mean in $y_{imp, m}$.}\label{fig:diagnostics}
\end{figure}

When we look at panel B, we conclude something weird. The \(AC\)-values
calculated with the default function indicate equal performance for the
typical convergence and the pathological non-convergence (up-to
\(t=5\)), while there is obvious trending in the \(\theta\)s of the
latter. Moreover, the best convergence (as indicated by the lowest
\(AC\)-value) is observed at \(t=2\), but looking at the chain means in
panel A, there should be some signs of trending up-to iteration number
seven. After consulting the documentation on \texttt{stats::acf()}
\citep{R}, we conclude that this \(AC\) function is not suited for
iterative imputation algorithms. The function is optimized for
performance when \(t\geq50\) \citep{box15}, while the default number of
iterations in iterative imputation is often much lower. Therefore, we
compute \(AC\) manually, see the solid line. These \(AC\) values
indicate non-stationarity as expected. We therefore calculate \(AC\)
manually.

\textbf{Describe panel C here and why it is not optimal{]}} The
\(\widehat{R}\)-values in panel D do meet the requirements specified
above. \(\widehat{R}\) as computed conform \citet{veht19} indeed
indicate less signs of non-convergence as the number of iterations goes
up.

\hypertarget{simulation-study}{%
\section{Simulation study}\label{simulation-study}}

The aim of the simulation study is to evaluate the impact of inducing
non-convergence in the MICE algorithm on several quantities of
scientific interest \(Q\). Subsequently we will evaluate how well
\(\widehat{R}\) and \(AC\) perform in identifying the effects of
non-convergence \textbf{{[}and performance of different thetas{]}}. And
finally, we will formulate an informed advice on the requirements to
safely assume \emph{sufficient} convergence in practice. That is, we
assume that the algorithm is \emph{sufficiently} converged when each
estimate \(\bar{Q}\) is an unbiased and confidence valid estimate of the
corresponding estimand \(Q\). \textbf{{[}and formulate advice on
thetas?{]}}

Non-convergence will be induced by: 1) increasing the missingness
proportion \(p_{mis}\) in dataset \(y\) incrementally, and 2)
terminating the imputation algorithm at a varying number of iterations
\(T\). The first set of simulation conditions---the missingness
proportions---is chosen to reflect the difficulty of the missingness
problem. The underlying assumption is that low missingness proportions
lead to quick algorithmic convergence, since there is a lot of
information in the observed data. Higher missingness proportions should
yield slower convergence. Unless, however, the fraction of missing
information is so high that the random component in the imputation
algorithm outweighs the information in the observed data. Then,
convergence to a stable but highly variable state may be reached
instanteneously. We assume that the incremental missingness proportions
in our study will result in a corresponding increase in signs of
non-convergence.

The assumption inherent to the second set of simulation conditions---the
number of iterations---is that terminating the imputation algorithm too
early causes non-convergence. Generally, the algorithm will not reach
convergence if \(T=1\), because the imputed values in the first
iteration (at \(t=1\)) depend on the initial values of the algorithm
(which are sampled randomly from the set of observed datapoints). As the
number of iterations increases, the imputation chains should become
independent of the initial values, until the point at which adding an
extra iteration does not lead to a more converged state. We assume that
we can induce non-convergence at least in conditions where \(T\) is
smaller than the default number of iterations in \texttt{mice}, \(T=5\).

\hypertarget{hypotheses}{%
\subsection{Hypotheses}\label{hypotheses}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  We expect that simulation conditions with a high missingness
  proportion \(p_{mis}\) and a low number of iterations \(T\) will
  result in biased, invalid estimates of the quantities of scientific
  interest, \(Q\)s.
\item
  We hypothesize that \(\widehat{R}\) and \(AC\) will correctly identify
  signs of non-convergence in those simulation conditions where the
  \(\bar{Q}\)s are \emph{not} unbiased and confidence valid estimates of
  the \(Q\)s.
\item
  We hypothesize that the recommended thresholds to diagnose
  non-convergence with \(\widehat{R}\) (\(\widehat{R} > 1.2\),
  \(\widehat{R} > 1.1\), and \(\widehat{R} > 1.01\)) may be too
  stringent for iterative imputation applications. In an empirical
  study, where \(\widehat{R}\) was used to inform the required
  imputation chain length, it took as many as 50 iterations to overcome
  the conventional non-convergence threshold \(\widehat{R}>1.2\). Yet,
  scientific estimates were insensitive to continued iteration from
  \(t=5\) onwards \citep{lace07}. We therefore suspect that
  \(\widehat{R}\) may over-estimate signs of non-convergence in
  iterative imputation algorithms, compared to the bias the estimates.
  In contrast to this, it may occur that signs of non-convergence are
  under-estimated by \(\widehat{R}\), in exceptional cases where the
  initial values of the algorithm are not appropriately over-dispersed
  \citep[p.~437]{broo98}. In e.g.~\texttt{mice}, initial values are
  chosen randomly from the observed data, hence we cannot be certain of
  over-dispersion in the initial values. In practice, we do not expect
  this to cause problems for identifying non-convergence with
  \(\widehat{R}\).
\item
  We expect that high \(AC\) values are implausible in iterative
  imputation algorithms with typical convergence. That is, after only a
  few iterations, the randomness induced by the algorithm should
  effectively mitigate the risk of dependency within chains.
\item
  \textbf{Something about thetas}
\end{enumerate}

\hypertarget{set-up}{%
\subsection{Set-up}\label{set-up}}

Convergence of the MICE algorithm is investigated through model-based
simulation in \texttt{R} \citep[version 3.6.3;][]{R}. The simulation
set-up is summarized in the pseudo-code below. The complete \texttt{R}
script of the simulation study is available from
\href{https://github.com/gerkovink/shinyMice/tree/master/3.Thesis/1.SimulationStudy}{github.com/gerkovink/shinyMice}.

\begin{verbatim}
# pseudo-code of simulation 
1. simulate data 
for (number of simulation runs from 1 to 1000)
 for (missingness proportions 5%, 25%, 50%, 75% and 95%)
  2. create missingness
  for (number of iterations from 1 to 100)
   3. impute missingness
   4. perform analysis of scientific interest
   5. compute non-convergence diagnostics 
   6. pool results across imputations
   7. compute performance measures
 8. combine outcomes of all missingness proportions
9. aggregate outcomes of all simulation runs 
\end{verbatim}

\hypertarget{data-generating-mechanism.}{%
\subsubsection{Data-generating
mechanism.}\label{data-generating-mechanism.}}

In this study, sampling variance is not of interest. Therefore, a single
complete dataset may serve as comparative truth in all simulation
repetitions \citep{vink14}. The data-generating mechanism is a
multivariate normal distribution, representing person-data on three
predictor variables in a multiple linear regression problem
\textbf{(from an unspecified social scientific field of study)}. Let the
predictor space be defined as \[
\begin{aligned}
\begin{pmatrix}X_1\\
X_2\\
X_3
\end{pmatrix} \sim \mathcal{N}
\begin{bmatrix}
\begin{pmatrix}
12\\
3\\
0.5
\end{pmatrix}\!\!,
\begin{pmatrix}
4 & 4 & 1.8 & 0\\
4 & 16 & 4.8 & 0\\
1.8 & 4.8 & 9 & 0
\end{pmatrix}
\end{bmatrix}\!\!\text{.}\\[2\jot]
\end{aligned}
\] A finite population of \(N=1000\) is simulated using the
\texttt{mvtnorm} package \citep{mvtnorm}. Subsequently, a fourth
variable is constructed as outcome \(Y\). For each unit
\(i = 1, 2,..., N\), let \[
Y_i = 1 + 2X_{1i} +.5X_{2i} - X_{3i} + \epsilon_i ,
\] where \(\epsilon \sim \mathcal{N}(0, 100)\). From the complete set,
we obtain the true values of several quantities of scientific interest,
\(Q\)s. \textbf{{[}MOVE THIS NEXT PART TO TOP OF SIMULATION SECTION?{]}}
\(Q\)s are the targets that will be estimated in each simulation
repetition, after first \emph{amputing} the complete data with varying
missingness proportions, then imputing the missingness with a varying
number of iterations, then estimating the \(Q\)s on the \(M\) completed
datasets, and finally pooling the \(M\) estimates \(\hat{Q}\) to obtain
a single \(\bar{Q}\) for each \(Q\). \textbf{We try to predict the
effect of non-converegence (i.e., the difference between \(Q\) and
\(\bar{Q}\)) with the convergence diagnostics \(\widehat{R}\) and
\(AC\), by appliying these methods to several scalar summaries of the
state space of the algorithm, \(\theta\)s.}

\hypertarget{scientific-estimands.}{%
\subsubsection{Scientific estimands.}\label{scientific-estimands.}}

We consider four types of \(Q\)s that are often of interest in empirical
research. Namely, two descriptive statistics: the mean \(\mu_j\) and
standard deviation \(\sigma_j\) of each variable
\(j = Y, X_1, X_2, X_3\). We also consider the regression coefficients
of the predictors: \(\beta_1\), \(\beta_2\), and \(\beta_3\) in
regression equation
\[Y' = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \beta_3 X_3,\] where \(Y'\)
is the expected value of the outcome. And finally, the proportion of
variance explained by the set of predictors: coefficient of
determination \(r^2 \times 100\) \textbf{(note that lower case \(r\) is
used to avoid confusion with non-convergence diagnostic
\(\widehat{R}\))}. The true values of these scientific estimands are
calculated on the complete data with the \texttt{R} functions
\texttt{base::mean()}, \texttt{stats::sd()}, and \texttt{stats::lm()}
\citep{R}.

\hypertarget{methods.}{%
\subsubsection{Methods.}\label{methods.}}

In each simulation repetition, we \emph{ampute} the complete dataset
five times to obtain five different incomplete datasets, \(y\)s. The
ratio between \(y_{obs}\) and \(y_{mis}\) in each \(y\) depends on the
missingness proportion: \(p_{mis} =.05,.25,.5,.75,.95\). We ampute the
complete data with the \texttt{mice} package \citep[function
\texttt{mice::ampute()};][]{mice}. We consider all possible univariate
and multivariate patterns of missingness, conform a `missing completely
at random' missingness mechanism \citep{rubin87}. I.e., the probability
to be missing is the same for all \(N \times k\) cells in \(y\),
conditional on the missingness proportion \(p_{mis}\).

Missing datapoints in each incomplete dataset \(y\) are imputed with the
\texttt{mice()} function \citep{mice}. All imputation procedures are
performed with Bayesian linear regression imputation, and five
imputation chains (\(M=5\)). Because \(M=5\), each run of the
\texttt{mice()} function results in five sets of imputations:
\(y_{imp,m}\), where \(m = 1, 2, ..., 5\). The number of iterations in
each run of the algorithm varies between simulation conditions
(\(T=1,2,...,100\)).

From the 5 imputations \(y_{imp,m}\) we obtain the chain means and chain
variances for each variable \(j\). These will serve as scalar summary
\(\theta\) to apply non-convergence diagnostics \(\widehat{R}\) and
\(AC\) on. We then use \texttt{mice::complete()} to combine the imputed
data \(y_{imp,m}\) with the observed data \(y_{obs}\), resulting in five
completed datasets \{\(y_{obs}, y_{imp, m}\)\}.

On the 5 completed datasets \{\(y_{obs}, y_{imp, m}\)\}, we:

\begin{itemize}
\item
  Estimate \(\hat{Q}\)s for \(Q= \mu\) and \(Q=\sigma\) with
  \texttt{base::mean()}, and \texttt{stats::sd()} \citep{R}. We pool the
  \(\hat{Q}\)s with \texttt{base::mean()} to get \(\bar{Q}\)s.
\item
  Compute \(\lambda_1\), to use as \(\theta\) in the two convergence
  diagnostics. By definition, the first eigenvalue of the
  variance-covariance matrix, \(\lambda_1\), is equal to the variance of
  the principal component solution of each \{\(y_{obs}, y_{imp, m}\)\},
  which we calculate with \texttt{stats::princomp()} \citep{R}.
\item
  Perform multiple linear regression with \texttt{stats::lm()} \citep{R}
  to get the \(\hat{Q}\)s for \(Q=\beta\) and \(Q=r^2\). We pool the
  \(\hat{Q}\)s conform \citet{vink14} to get \(\bar{Q}\)s for the
  \(\beta\)s and \(r^2\) \textbf{{[}check pooling of \(r^2\), instead of
  \texttt{mice::pool.r.squared()}{]}}. Additionally, we use the
  \(\hat{Q}\)s of \(Q=\beta\) as \(\theta\) to apply the convergence
  diagnostics to.
\end{itemize}

We apply non-convergence diagnostics \(\widehat{R}\) and \(AC\) to each
scalar summary \(\theta\). We calculate \(\widehat{R}\) by implementing
Vehtari et al.'s \citeyearpar{veht19} recommendations, and \(AC\) as the
correlation between the \(t^{th}\) and the \((t+1)^{th}\) iteration.

\hypertarget{performance-measures.}{%
\subsubsection{Performance measures.}\label{performance-measures.}}

We assess the impact of inducing non-convergence in the iterative
imputation algorithm by comparing \(\bar{Q}\)s with \(Q\)s. For each
\(Q\), we calculate bias as \(\bar{Q} - Q\). For the regression
coefficients \(Q=\beta_{1,2,3}\), we also compute the empirical coverage
rate (CR). CR is defined as the percentage of simulation repetitions in
which the 95\% confidence interval (CI) around \(\bar{Q}\) covers the
true estimand \(Q\). Let
\[\text{CI} = \bar{Q} \pm t_{(M-1)} \times SE_{\bar{Q}},\] where
\(t_{(M-1)}\) is the quantile of a \(t\)-distribution with \(M-1\)
degrees of freedom, and \(SE_{\bar{Q}}\) is the square root of the
pooled variance estimate. If we obtain nominal coverage (CR = 95\%), we
can conclude that the \(\bar{Q}\)s are conficence valid estimates of the
\(Q\)s. Finally, we inspect CI width (CIW): the difference between the
lower and upper bound of the 95\% confidence interval around
\(\bar{Q}\). CIW is of interest because too short CIs indicate
under-estimation of the variance in \(\bar{Q}\), which may yield
spurious inferences.

With these performance measures, we also evaluate how well the
non-convergence diagnostics \(\widehat{R}\) and \(AC\) can identify
simulation conditions in which the \(\bar{Q}\)s are \emph{not} unbiased,
confidence valid estimates of the \(Q\)s. \textbf{Rephrase and add
corresponding \(Q\)s:} We measure the performance of \(\widehat{R}\) and
\(AC\) by the four scalar summaries \(\theta\): chain means (i.e., the
mean in each \(y_{imp, m, j}\)) versus \(Q=\mu\), chain variances (i.e.,
the variance in each \(y_{imp, m, j}\)) versus \(Q=\sigma\), the
estimated regression coefficients \(\beta\) in each
\{\(y_{obs}, y_{imp, m}\)\} (i.e., a user-defined statistic of
scientific interest \(\hat{Q}\)) against \(Q=\beta\), and the first
eigenvalue of the variance-covariance matrix in each
\{\(y_{obs}, y_{imp, m}\)\}, \(\lambda_1\) versus \(Q=r^2\).

\hypertarget{results}{%
\section{Results}\label{results}}

\textbf{{[}Add to this section: 1) emphasize that the iterations in the
plots are averages across repetitions, not within one run of MICE; 2)
sample effects due to single complete dataset?? 3) Add more info about
figure legends and axes.{]}} Primarily, we want to know what the effect
is of inducing non-convergence on the estimated \(Q\)s. Specifically, we
are interested in the bias of \(\bar{Q}\) for all \(Q\)s and the
confidence validity of \(\bar{Q}\) for \(Q=\beta\). Therefore, we plot
these performance measures against the simulation conditions (the number
of iterations \(T\) and the missingness proportion \(p_{mis}\)). We look
at averages across simulation repetitions.

Secondly, we want to know if conditions in which the \(Q\)s are affected
by non-convergence are also identified by the non-convergance
diagnostics \(\widehat{R}\) and \(AC\). We do this, again, by looking at
the performance measures for the \(Q\)s, but now we evaluate these
against what \(\widehat{R}\) and \(AC\) indicate about signs of
non-convergence in the \(\theta\)s.

\textbf{{[}In short:{]}} - Is there more bias in conditions where
non-convergence was induced? (NOPE for univariates, YES for multivar.).

\begin{itemize}
\item
  Is the point of \emph{sufficient} convergence (at which the
  \(\bar{Q}\)s are unbiased and confidence valid estimates of the
  \(Q\)s) the same point as identified by the non-convergence
  diagnostics? (NOPE, they are too conservative).
\item
  Is one of the thetas better? (NAH, doesn't really matter, probably
  multivariate is better).
\item
  What should the thresholds be? (LOWER than recommended).
\item
  And the default nr of iterarions? (MAYBE increase to 10?).
\end{itemize}

For reasons of brevity, we only discuss the convergence diagnostics for
the \(Q\)s with the worst performance in terms of bias. Bias in the
descriptve statistics (\(Q=\mu\) and \(Q=\sigma\)) was most pronounced
in the outcome variable \(Y\). Of the regression coefficients,
\(Q=\beta\), the largest bias was observed in the effect of \(X_1\) on
\(Y\), \(\beta_1\). Results for \(Q=r^2\) are reported as-is. In the
figures, however, we only present simulation conditions where
\(T\leq50\). Full results are available from \textbf{{[}link github
repo{]}}.

\hypertarget{in-general}{%
\subsection{In general}\label{in-general}}

As expected, conditions with a higher proportion of missingness and/or a
lower number of iterations show more signs of non-convergence. I.e., on
average these conditions portray more extreme bias in the estimated
\(Q\)s, and non-nominal coverages. Roughly speaking this means that MICE
is indeed not converged at \(t=1\), and converges gradually as \(t\)
increases to \(T\). The point at which an additional iteration does not
lead to an improvement of the estimates depends on the difficulty of the
missingness problem---between XYZ and XYZ. This means that the algorithm
has converged sufficiently (\textbf{under the current specifications}).

As required for a method to diagnose non-convergence, the values of
\(\widehat{R}\) follow a trend similar to the performance measures.
\textbf{Add split between convergence diagnostics versus performance
measures, and convergence diagnostics versus simulation conditions.}
\(\widehat{R}\) values are generally lower in conditions with a higher
number of iterations, and somewhat higher in conditions with a higher
percentage of missing data. Autocorrelation values were indeed
decreasing with a higher number of iterations, and lower in conditions
with less missingness.

Evaluation with the performance measures shows that \(\widehat{R}\) is
conservative: it identifies signs of non-convergence in conditions where
the \(\bar{Q}\)s are unbiased and/or confidence valid estimates of the
\(Q\)s.

\hypertarget{means}{%
\subsection{Means}\label{means}}

\textbf{Message: after one iteration (\(T\)=2), the algorithm is as well
converged as the missingness proportion allows. Sooner even with lower
missingness proportions (\(T\)=1 might be ok for \(p_{mis}\) = 5 and
25).} Number of it has little influence on bias in \(\hat{Q}\), only in
first iteration some bias in missingness conditions of 25 and up. There
is a clear different between the missingness conditions. The higher the
missingness prop, the larger the magnitude of the bias. For this \(Q\),
t=2 would be sufficient. And only for \(p_{mis}\) = 5 the \(\bar{Q}\)s
are unbiased. All other \(p_{mis}\) conditions have bias irrespective of
the number of iterations.

\(\widehat{R}\) differs a lot across iteration conditions. But not
structurally for missingness proportions. Highest \(\widehat{R}\)s for
the condition \(T\)=2, while the most bias is for (\(T\)=1), but there
is no \(\widehat{R}\) for that by definition. But for missingness, 25\%
seems the most signs of non-conv, while it has one of the least bias.
Threshold of 1.2 reached in condition it = 3, \(p_{mis}\) = 5 and 75,
and then for it\textgreater6, \(p_{mis}\) = 5 and 50, and
\(T\)\textgreater10 for all \(p_{mis}\). Threshold 1.1 reached in in
\(T\)=12, \(p_{mis}\) = all but 75, 1.01 only reached in \(T\) = 93,
\(p_{mis}\) = 75. While for this \(Q\), \(T\)=2 would be sufficient.

\(AC\) mainly shows effects of the set of iteration conditions. \(AC\)
indicates less signs of non-convergencein conditions with higher number
of iterations. For iteration conditions where \(T\)\textless10, there
are also some effects of the missingness proportions. The \(p_{mis}\)
condtions 25 and 95 \% indicate the most signs of non-convergence. For
\(p_{mis}\) = 95\% this corresponds to the bias in Qbar. For 25\% miss
it doesn't. The thresholds defined by the critical value is only met in
the conditions it=3, \(p_{mis}\) = 5,75, and 95\%.

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/mean-1}

}

\caption{Convergence diagnostics chain mean.}\label{fig:mean}
\end{figure}

\hypertarget{standard-deviations}{%
\subsection{Standard deviations}\label{standard-deviations}}

\textbf{Message: even with possible effect of initial values (\(T\)=1),
the algorithm is as well converged as the missingness proportion
allows.} For \(Q=\sigma_Y\) there is no effect visible of the iteration
conditions, but clearly for the \(p_{mis}\) conditions. Most bias for
\(p_{mis}\) = 95, followed by \(p_{mis}\) = 50 and 75. For the least
possible bias we may use any number of iterations \(T\geq1\), but none
of the \(p_{mis}\) results in unbiased \(\bar{Q}\)s.

There is a clear effect of the number of iterations on \(\widehat{R}\).
Conditions with lower Ts have more signs of non-conv, as identified by
\(\widehat{R}\). There is little difference between missingness
conditions. Only for conditions where \(T<5\) the missingness
proportions seem to affect whether \(\widehat{R}\) indicates
non-convergence The bias in \(\bar{Q}\) seems unrelated to the magnitude
of \(\widehat{R}\). Conditions with more bias (\(p_{mis}\) = 50,75,95)
are not identified to have more signs of non-convergence. The thresholds
are therefore useless (because depend mostly on \(T\), not so much
\(p_{mis}\)).

\(AC\) is affected by both t and \(p_{mis}\) conditions. But conditions
with worst bias (\(p_{mis}\) = 50, 75, 95) are not identified as worst
converging according to AC. Moreover, only for \(T\) =3 the \(AC\) is
significantly different from zero.

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/sd-1}

}

\caption{Convergence diagnostics chain variance.}\label{fig:sd}
\end{figure}

\hypertarget{regression-coefficients}{%
\subsection{Regression coefficients}\label{regression-coefficients}}

\textbf{Message: after four iterations (\(T\)=5), the algorithm is as
well converged as the missingness proportion allows. Sooner even with
lower missingness proportions (\(T\)=1 might be ok for \(p_{mis}\) = 5
and 25).} The figure for \(Q=\beta_1\) shows that the \(\bar{Q}\)s are
affected by both the number of iterations and the missingness
proportion. For \(T\) =1, the magnitude of the bias (almost) follows the
order of the missingness proportions. More missingness causes more bias
in the \(\bar{Q}\). In general, a lower nr of iterations results in more
bias, but this depends on the missingness proportion. For \(p_{mis}\) =
95, there is more bias in conditions where \(T\)\textless6. When
\(p_{mis}\) = 50, it's \(T\)\textless3. For \(p_{mis}\) = 75 and 25,
it's only for \(T\)\textless2. And conditions where \(p_{mis}\) = 5 are
unaffected on average. However, none of the conditions result in
completely unbiased estimates of \(Q\).

We should see this in the conv diag: higher \(\widehat{R}\) and \(AC\)
for \(p_{mis}\) = 75 and 95, and for condtions where \(T<2-6\),
depending on \(p_{mis}\).

\(\widehat{R}\) depends somewhat on the missingness proportion. For
conditions where \(T\)\textless10, the \(\widehat{R}\) values follow the
\(p_{mis}\) more or less. But what we actually want to know is whether
they are indicators of bias. Conditions with the worst bias (high
\(p_{mis}\), low T) have the largest \(\widehat{R}\) values. The
thresholds are not appropriate to diagnose sufficient convergence. All
thresholds over-estimate the signs of non-convergence compared to the
bias in Qbar. However, a higher threshold than 1.2 may result in
terminating the algorithm after 3 iterations, which is insufficient for
the \(p_{mis}\) = 95 condition. Threshold of 1.25 or 1.3 could work for
the other \(p_{mis}\) conditions.

\(AC\) is very affected by the p\_mis. The two conditions with the most
bias (\(p_{mis}\) = 75 and 95) are the only ones identified as
non-converging, as they should. However, the \(p_{mis}\) = 75 condition
is only identified in the \(T\) =3 condtion, while the bias is the least
pronounced there. And the \(AC\) only identifies non-convergence in the
\(p_{mis}\) = 95 conditions where \(T\)\textgreater10. In the default
situation, \(T\) =5, this would be overlooked. Also, the \(AC\) shows
the opposite of what you would expect from a theta that is based on the
observed data: shouldn't there be higher autocorrelations in conditions
with comparatively more observed data in the completed datasets? I don't
understand this.

\textbf{Add how interesting it is because we literally track \(Q\) and
are still not ok.}

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/est-1}

}

\caption{Convergence diagnostics regression coefficient.}\label{fig:est}
\end{figure}

\hypertarget{variance-explained}{%
\subsection{Variance explained}\label{variance-explained}}

\textbf{Message: after two iterations (\(T\)=3), the algorithm is as
well converged as the missingness proportion allows. Sooner even with
lower missingness proportions (\(T\)=1 might be ok for \(p_{mis}\) = 5
and 25).} Bias is the worst for conditions where \(p_{mis}\) = 95, \(T\)
\textless3. Followed by \(p_{mis}\) = 75, \(T\) =1; and \(p_{mis}\) =
50, \(T\) = 1. Conditions where \(p_{mis}\) = 25 or 5 seem unaffected by
the number of iterations. Minimal bias is obtained in all iteration
conditions. We should see this in \(\widehat{R}\) and AC: worst
convergence in \(p_{mis}\) = 95 and 50; some non-convergencein
conditions where \(p_{mis}\) = 75 and \(T\) =1, and no non-convergencein
conditions where \(p_{mis}\) = 25 or 5. The sign of the bias is logical:
when there is less info in the data, the relations between the
predictors and the outcome is under-estimated.

\(\widehat{R}\) behaves oppositely to what we wanted. The lowest
\(\widehat{R}\) values for conditions where \(T\)\textless10 are
observed for the two missingness conditions with the worst bias.
However, the highest \(\widehat{R}\) observed is indeed for the
missingness condition with the worst bias, \(p_{mis}\) =95. But only at
\(T\) =5 (while bias is no worse than anywhere else in this \(p_{mis}\)
condition, conditional on iterations condition \(T\)\textgreater2). So
neither of the thresholds works perfectly.

\(AC\) shows a similar trend to the biases: more signs of
non-convergencefor higher \(p_{mis}\) and lower T. The threshold,
however, performs less well. It identifies signs of non-convergence for
all conditions where \(T=3\), irrespective of p\_mis. Then nothing, and
then for \(p_{mis}\) =95, \(T\)\textgreater7, and \(p_{mis}\) = 50,
\(T\) \textgreater14, and \(p_{mis}\) = 75, \(T\)\textgreater20. This
does not corresond to the bias in \(\bar{Q}\). Maybe use
AC\textgreater.5 at \(T\) =5 as threshold?

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/pred-1}

}

\caption{Convergence diagnostics regression coefficient.}\label{fig:pred}
\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

In short, univariate estimates seem robust against terminating the
algorithm early: There is no clear effect of the number of iterations on
the bias in the estimated means and standard deviations. The bias in
these \(Q\)s only depends on the \(p_{mis}\), with higher \(p_{mis}\)
resulting in more bias. Yet, the convergence diagnostics indicate that
the algorithm did not reach a completely converged state (yet).

\(\widehat{R}\) and autocorrelation indicate that algorithmic
convergence may only be reached after twenty or even forty iterations,
while unbiased, confidence valid estimates may be obtained with as
little as one iteration. These results are in agreement with the
simulation hypothesis: \(\widehat{R}\) over-estimates the severity of
non-convergence when applied to MI procedures.

With that, we show that MICE can lead to correct outcomes when they have
not converged according to two common conv diags. This may be due to the
methods (and their thresholds) or due to the \(Q\)s (descriptives and
multivariate linear regression, not higher dimensional/more complex
\(Q\)s). More `complicated' \(Q\)s (e.g., higher-order effects or
variance components) might show bias, under- or over-coverage at higher
\(T\), as indiated by \(\widehat{R}\) and AC. Add what \%miss has to do
with it.

\hypertarget{recommendations-for-empirical-researchers}{%
\subsection{Recommendations for empirical
researchers}\label{recommendations-for-empirical-researchers}}

For empirical researchers:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Check trace plots for pathological non convergence and adjust
  imputation model if necessary.
\item
  use \(\widehat{R}\) wait 1.1 ash threshold and autocorrelation with
  \textbf{?} as threshold. Keep iterating until these thresholds are
  reached.
\item
  Do not use the R function ACF. Instead, compute autocorrelations
  manually \textbf{(see e.g., {[}GitHub link{]})}.
\item
  Track your own scalar summary of interest. This is somewhat advanced
  but explained in Van Buren 2018. Compute \(\widehat{R}\) and
  autocorrelation values for this scalar summary.
\item
  \textbf{Something} about the novel \(\theta\) that is `substantive
  model-independent'.
\end{enumerate}

\hypertarget{recommendations-for-future-research}{%
\subsection{Recommendations for future
research}\label{recommendations-for-future-research}}

Further research is needed to investigate their performance under clear
violation of convergence, e.g.~dependency between predictors (predictors
with very high correlations). Also:

\begin{itemize}
\item
  Induce non-convergence with mis-specified imputation model.
\item
  More difficult missingness problems, i.e., M(N)AR instead of MCAR.
  Proper performance under a `missing completely at random' missingness
  mechanism is necessary to demonstrate the appropriateness of
  \(\widehat{R}\) and \(AC\) as non-convergence diagnostics. However,
  results may not be extrapolated to other missingness mechanisms.
\item
  Imputation technique with questionable convergence properties, e.g.,
  PMM.
\item
  Apply recommendations to empirical data and write a vignette for
  applied researchers.
\item
  Implement recommendations in ShinyMICE.
\item
  Develop an additional convergence diagnostic unique for iterative
  imputation: \(AC\) across imputations, instead of iterations.
\end{itemize}

Uitsmijter: To conclude, in the case of drawing inference from
incomplete data, convergence of the iterative imputation algorithm is
often a convenience but not a necessity.

\hypertarget{grading}{%
\section{Grading}\label{grading}}

\begin{itemize}
\item
  Scientific contribution (clear research question, convincing
  argumentation that the study adds to the scientific literature)
\item
  Theory (use of theory, clear hypotheses, relevant literature)
\item
  Method (correct data collection/handling, measures, choice and
  description of method of analysis)
\item
  Results (correct interpretation of findings, clear presentation)
\item
  Discussion and conclusions (summary, broader implications,
  limitations, future research)
\item
  Written presentation (writing, structure, tables and figures,
  references)
\end{itemize}

\bibliographystyle{sageh}
\bibliography{thesis.bib}

\end{document}