### 2020-05-04 3.Thesis/1.SimulationStudy/Writeup/manuscript.tex
Add requirements thesis submission

\documentclass[Royal,times,sageh]{sagej}

\usepackage{moreverb,url,natbib, multirow, tabularx}
\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

\begin{document}

\title{Missing the Point: Non-Convergence in Iterative Imputation Algorithms}

\runninghead{Oberman}

\author{H. I. Oberman\affilnum{1}}

\affiliation{\affilnum{1}{Department of Methodology and Statistics, Utrecht University, Utrecht, The Netherlands}}

\corrauth{Hanne Oberman, Sjoerd Groenman building, Utrecht Science Park, Utrecht, The Netherlands.}

\email{\href{mailto:h.i.oberman@uu.nl}{\nolinkurl{h.i.oberman@uu.nl}}}

\begin{abstract}
Iterative imputation is a popular tool to accommodate missing data. While it is widely accepted that valid inferences can be obtained with this technique, these inferences all rely on algorithmic convergence. There is no consensus on how to evaluate the convergence properties of the method. This paper provides insight into identifying non-convergence of iterative impuation algorithms.
\end{abstract}

\keywords{missing data, iterative imputation, non-convergence, mice}

\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Anyone who analyzes person-data may run into a missing data problem. Missing data is not only ubiquitous, but treating it can also be tedious. If a dataset contains just one incomplete observation, statistical inferences are undefined and will not produce any results. To circumvent this, many statistical packages employ list-wise deletion by default (i.e., ignoring incomplete observations). Unfortunately, this \emph{ad hoc} solution may yield wildly invalid results \citep{buur18}. An alternative is to \emph{impute} (i.e., fill in) the missing values in the incomplete observations. Subsequently, statistical inferences can be performed on the completed dataset. By repeating this process several times, a distribution of plausible results may be obtained, which reflects the uncertainty in the data due to missingness. This technique is known as `multiple imputation' \citep[MI;][]{rubin76}. MI has proven to be a powerful technique to draw valid inferences from incomplete data under many circumstances \citep{buur18}.

Figure \ref{fig:diagram} provides an overview of the steps involved with MI. Missing data \(y_{mis}\) in an incomplete dataset is imputed \(M\) times. The imputed data \(y_{imp}\) is combined with the observed data \(y_{obs}\) to create \(M\) completed datasets. On each completed dataset, an analysis of scientific interest is performed. The quantity of scientific interest (e.g., a regression coefficient) is denoted with \(Q\). Since \(Q\) is estimated on each completed dataset, \(M\) separate \(\hat{Q}\)-values are obtained. These \(M\) values are combined into a single pooled estimate \(\bar{Q}\). The premise of multiple imputation is that \(\bar{Q}\) is an unbiased and confidence valid estimate of the true---but missing---data inference.

\begin{figure}

{\centering \includegraphics[width=\linewidth]{./images/diagram}

}

\caption{Scheme of the main steps in multiple imputation (where $M=3$)---from an incomplete dataset, to $M$ multiply imputed datasets, to $M$ estimated quantities of scientific interest $\hat{Q}$s, to a single pooled estimate $\bar{Q}$.}\label{fig:diagram}
\end{figure}

A popular method to obtain imputations is to use the `Multiple Imputation by Chained Equations' algorithm, shorthand `MICE'\citep{mice}. With MICE, imputed values \(y_{imp}\) are drawn from the posterior predictive distribution of the missing values \(y_{mis}\). The algorithm is named after the iterative algorithmic procedure by which imputed values are generated: a multivariate distribution is obtained by iterating over a sequence of univariate imputations. The iterative nature of algorithms like MICE introduces a potential threat to the validity of the imputations: What if the algorithm has not converged? Are the implications then to be trusted? And can we rely on the inference obtained on the completed data? These are all open questions, because the convergence properties of iterative imputation algorithms have not been systematically studied \citep{buur18}. Moreover, there is no scientific consensus on how to evaluate convergence of imputation algorithms \citep{taka17}. Some default MICE techniques (e.g., `predictive mean modeling') might not yield converged states at all \citep{murr18}. Therefore, algorithmic convergence should be monitored carefully.

Currently, the recommended practice for evaulating the convergence of iterative imputation algorithms is through visual inspection. That is, the state space of the algorithm is monitored by plotting a scalar summary \(\theta\) against the iteration number.
\textbf{{[}Segway to WHAT TO MONITOR? Also, this paragraph is more or less the same as the third paragraph in the second section. Merge or remove one of the two?{]}} Typically, the \(\theta\)s under evaluation are chain means and variances (\textbf{define!}). These \(\theta\)s may be insufficient because they are univariate summaries of the state space of the algorithm, while MICE is not only concerned with univariate estimates, but the entire multivariate distribution of \(y_{imp}\). A suggestion by \citet{buur18} for a multivariate \(\theta\) to monitor is the quantity of scientific interest \(Q\) (e.g., a regression coefficient). Implementing this, however, might be somewhat advanced for empirical researchers. Moreover, this scalar summary is not model-independent, i.e., it only applies to one model of scientific interest. Yet, one of the advantages of MI is that the missing data problem and scientific problem are solved independently. Therefore, this \(\theta\) is not sufficient either. Van Buuren \citeyearpar[\(\S\) 4.5.2]{buur18} also proposed multivariable evaluation of the MICE algorithm through eigenvalue decomposition, building on the work of \citet{mack03}. Eigenvalues of a variance-covariance matrix are a measure of the data's total covariance. This would be a model-independent, multivariate scalar summary to monitor, but it has not yet been implemented.
\textbf{{[}Segway to HOW TO MONITOR?{]}} Monitoring convergence visually is insufficient on two counts: 1) it may be challenging to the untrained eye, and 2) only severely pathological cases of non-convergence may be diagnosed \citep[\(\S\) 6.5.2]{buur18}. Therefore, a quantitative, diagnostic evaluation of convergence would be preferred---although not straightforward. Iterative imputation algorithms such as MICE are Markov chain Monte Carlo (MCMC) methods. In MCMC methods, convergence is not from a scalar to a point, but from one distribution to another. The values generated by the algorithm (e.g., imputed values) will vary even after convergence. Since MCMC algorithms do not reach a unique point at which convergence is established, diagnostics may only identify signs of \emph{non}-convergence \citep{hoff09}. Several of such non-convergence diagnostics exist, but it is not known whether these are appropriate for iterative imputation algorithms.

In this paper, we investigate how non-convergence in iterative imputation algorithms may be diagnosed. First, we define which non-convergence diagnostics will be considered, and evaluate how these diagnostics could be appropriate for iterative imputation applications. In the second part of this paper, we will investigate the impact of inducing non-convergence in iterative imputation algorithms by means of model-based simulation in \texttt{R} \citep{R}. The aim of the simulation is to determine whether unbiased, confidence valid inferences may be obtained under different levels of non-convergence (\textbf{and at which point convergence may safely be assumed}). Additionally, we will assess how well different non-convergence diagnostics perform in identifying signs of non-convergence. The results of this study are guidelines for identifying non-convergence in iterative imptation algorithms, which will aid applied researchers in drawing valid inference from incomplete datasets. \textbf{{[}add more societal relevance: what is the consequence of non-convergence? e.g., biased estimates, invalid inference, etc.{]}} For reasons of brevity, we only focus on the iterative imputation algorithm implemented in the popular \texttt{mice} package \citep{mice} in \texttt{R} \citep{R}.

\textbf{{[}Include sub-questions and make structure of the paper clear!{]}} How can non-convergence be identified diagnostically? Are common MCMC non-convergence diagnostics appropriate for MICE? And if so, which threshold should be used to diagnose non-convergence? How many iterations are sufficient/needed to be able to diagnose non-convergence? Are the default number of iterations sufficient (i.e., 5 in mice, 10 in SPSS and Stata, 30 in mi)? How severe is it when the algorithm has not converged? And what are guidelines for practice? Can the parameter of interest \(Q\) be correct when the algorithm is not (yet) converged, and vice versa?

\hypertarget{some-notation}{%
\subsection{Some notation}\label{some-notation}}

Let \(y\) denote an \(n \times k\) matrix containing the data values on \(k\) variables for all \(n\) units in a sample. The data value of unit \(i\) (\(i = 1, 2, \dots, n\)) on variable \(j\) (\(j = 1, 2, \dots, k\)) may be either observed or missing. The number of units \(i\) in dataset \(y\) with at least one missing value, divided by the total number of units \(n\), is called the missingness proportion \(p_{mis}\). The collection of observed data values in \(y\) is denoted by \(y_{obs}\); the missing part of \(y\) is referred to as \(y_{mis}\). For each datapoint in \(y_{mis}\), we sample \(M \times T\) plausible values, where \(M\) is the number of imputations (\(m = 1, 2, \dots, M\)) and \(T\) is the number of iterations (\(t = 1, 2, \dots, T\)) in the imputation algorithm. The collection of samples between the initial value (at \(t=1\)) and the final imputed value (at \(t=T\)) will be referred to as an `imputation chain'. \textbf{\(\theta\)s are scalar summaries of the algorithm at a certain iteration (e.g., chain means; the average of the imputed values in each imputation chain). }

\hypertarget{identifying-non-convergence}{%
\section{Identifying non-convergence}\label{identifying-non-convergence}}

There are two requirements for convergence of iterative algorithms: mixing and stationarity \citep{gelm13}. Without mixing, imputation chains do not intermingle nicely, indicating that the distribution of imputed values differs across imputations. Chains may be `stuck' at a local optimum, instead of sampling imputed values from the entire predictive posterior distribution of the missing values. This may cause under-estimation of the variance within chains, which results in spurious, invalid inferences. Without stationarity, there is trending within imputation chains. Trending implies that further iterations would yield a different set of imputations. Iterative imputation algorithms that have not yet reached stationarity, may thus yield biased estimates.

\textbf{{[}Move to INTRO???{]}} To illustrate what non-convergence looks like in iterative imputation algorithms, we reproduce an example from van Buuren \citeyearpar[\(\S\) 6.5.2]{buur18}. Figure \ref{fig:non-conv} displays a subset of the example: a variable that we will call \(j\). We see two traceplots of the chain means for \(j\)---i.e., the average of the imputed values for \(j\) in each imputation \(y_{imp,m}\), plotted against the iteration number. Each line portrays one imputation. The plot on the lefthand side of the figure shows typical convergence of an iterative imputation algorithm. The righthand side displays pathological non-convergence, induced by purposefully mis-specifying the imputation model. \textbf{{[}Leave this in if it's after mixing and stationarity are defined, otherwise remove?{]} In the typical convergence situation, the imputation chains intermingle nicely and there is little to no trending. In the non-convergence plot, there is a lot of trending and some chains do not intermingle.} Importantly, the chain means at the last iteration (the imputed value per \(m\)) are very different between the two plots. The algorithm with the mis-specified model yields imputed values that are on average a factor two larger than those of the typically converged algorithm. That means that non-convergence has a impact on the mean of \(j\) in \(y_{imp, m}\). This difference (presumably) translates to bias in the \(M\) completed data estimates \(\hat{\mu}_{j,m}\), and consequently also to the pooled estimate \(\bar{\mu}_j\). Non-convergence thus leads to biased estimates. This shows the importance of reaching converged states in iterative imputation algorithms.

\begin{figure}

{\centering \includegraphics[width=.49\linewidth]{manuscript_files/figure-latex/non-conv-1} \includegraphics[width=.49\linewidth]{manuscript_files/figure-latex/non-conv-2}

}

\caption{Typical convergence versus pathological non-convergence. Please note that this is the same data with a different imputation model, leading to different imputations (e.g., see the units on the y-axis).}\label{fig:non-conv}
\end{figure}

\textbf{{[}Move to INTRO???{]}} Convergence has historically been inspected visually, by monitoring the imputation algorithm over iterations. This is typically done through traceplots. In a traceplot, the iteraton number is plotted against a certain \(\theta\). \(\theta\)s are scalar summaries of the state space of the algorithm at a specific iteration. The recommended scalar summaries to evaluate are chain means and chain variances. Additionally, researchers may ``monitor some statistic of scientific interest'' \citep[\(\S\) 6.5.2]{buur18}. As \citet{buur18} describes, researchers can specify their quantity of scientific interest \(Q\) (e.g., a regression coefficient) as scalar summary \(\theta\). Such a user-defined scalar summary, however, may be somewhat advanced for empirical researchers. And moreover, it is not universal to all complete data problems. Focusing on the convergence of outcome parameters may influence the iterative imputation procedure in the sense that the model of evaluation favors the model of interest. The downside to these two approaches is that they either focus on the univariate state space, or primarily track the change over the iterations of a multivariate outcome conform the scientific model of interest. Ideally, one would like to evaluate a model-independent parameter that summarizes the multivariate nature of the data. Therefore, we propose a \(\theta\) that summarizes the multivariate state space of the algorithm, but is independent from the model of scientific interest. We propose \(\lambda_1\) as such a scalar summary. We define \(\lambda_1\) as the first eigenvalue of the variance-covariance matrix of the completed data. Let \(\lambda_1 \geq \lambda_2 \geq ... \geq \lambda_j\) be the eigenvalues of \(\Sigma\) in each of the \(M\) completed datasets \({y_{obs}, y_{imp}}\). \(\lambda_1\) is measure that summarises the covariances in the completed datasets. \textbf{The first eigenvalue has the appealing property that is not dependent on the substantive model of interest. Eigenvalue decomposition is inspired by \citet{mack03}.}

\hypertarget{diagnostics-under-evaluation}{%
\subsection{Diagnostics under evaluation}\label{diagnostics-under-evaluation}}

Non-convergence diagnostics for MCMC algorithms typically identify problems in either one of the two requirements for converegnce: mixing or stationarity. Non-stationarity within chains may be diagnosed with e.g., autocorrelation \citep[\(AC\);][]{scha97, gelm13}, numeric standard error \citep[`MC error';][]{gewe92}, or Raftery and Lewis's \citeyearpar{raft91} procedure to determine the effect of trending on the precision of estimates. A widely used diagnostic to monitor mixing between chains is the potential scale reduction factor \(\widehat{R}\) \citep[`Gelman-Rubin statistic';][]{gelm92}. With a recently proposed adaptation, \(\widehat{R}\) might also serve to diagnose non-stationarity, but this has not yet been thoroughly investigated \citep{veht19}. Therefore, we use \(\widehat{R}\) and \(AC\) to evaluate mixing and stationarity separately, as recommended by e.g., \citet{cowl96}.

\hypertarget{potential-scale-reduction-factor}{%
\subsubsection{Potential scale reduction factor}\label{potential-scale-reduction-factor}}

An updated version of \(\widehat{R}\) has been proposed by Vehtari et al.~\citeyearpar[p.~5]{veht19}. \textbf{This version may be suitable for iterative imputation.} Let \(M\) be the total number of chains, \(T\) the number of iterations per chain, and \(\theta\) the scalar summary of interest (e.g., chain means). For each chain (\(m = 1, 2, \dots, M\)), we estimate the variance of \(\theta\), and average these to obtain within-chain variance \(W\).

\begin{align*}
W&=\frac{1}{M} \sum_{m=1}^{M} s_{j}^{2}, \text { where } s_{m}^{2}=\frac{1}{T-1} \sum_{t=1}^{T}\left(\theta^{(t m)}-\bar{\theta}^{(\cdot m)}\right)^{2}. 
\end{align*}

We then estimate between-chain variance \(B\) as the variance of the collection of average \(\theta\) per chain.

\begin{align*}
B&=\frac{T}{M-1} \sum_{m=1}^{M}\left(\bar{\theta}^{(\cdot m)}-\bar{\theta}^{(\cdot \cdot)}\right)^{2}, \text { where } \bar{\theta}^{(\cdot m)}=\frac{1}{T} \sum_{t=1}^{T} \theta^{(t m)} \text{, } \bar{\theta}^{(\cdot \cdot)}=\frac{1}{M} \sum_{m=1}^{M} \bar{\theta}^{(\cdot m)}. 
\end{align*}

From the between- and within-chain variances we compute a weighted average, \(\widehat{\operatorname{var}}^{+}\), which over-estimates the total variance of \(\theta\) \textbf{remove or explain why, or leave in}. \(\widehat{R}\) is then obtained as a ratio between the over-estimated total variance and the within-chain variance:

\begin{equation*}
\widehat{R}=\sqrt{\frac{\widehat{\operatorname{var}}^{+}(\theta | y)}{W}},
\text{ where } \widehat{\operatorname{var}}^{+}(\theta | y)=\frac{N-1}{N} W+\frac{1}{N} B.
\end{equation*}

We can interpret \(\widehat{R}\) as potential scale reduction factor since it indicates by how much the variance of \(\theta\) could be shrunken down if an infinite number of iterations per chain would be run \citep{gelm92}. This interpretation assumes that chains are `over-dispersed' at \(t=1\), and reach convergence as \(T \to \infty\). Over-dispersion implies that the initial values of the chains are `far away' from the target distribution and each other. When all chains sample independent of their initial values, the mixing component of convergence is satisfied, and \(\widehat{R}\)-values will be close to one. High \(\widehat{R}\)-values thus indicate non-convergence. The conventionally acceptable threshold for convergence was \(\widehat{R} < 1.2\) \citep{gelm92}. More recently, \citet{veht19} proposed a more stringent threshold of \(\widehat{R} < 1.01\).

\hypertarget{autocorrelation}{%
\subsubsection{Autocorrelation}\label{autocorrelation}}

Following the same notation, we define autocorrelation as the correlation between two subsequent \(\theta\)-values within the same chain \citep[p.~147]{lync07}. In this study, we only consider \(AC\) at lag 1, i.e., the correlation between the \(t^{th}\) and \(t+1^{th}\) iteration of the same chain.

\begin{equation*}
AC = \left( \frac{T}{T-1} \right) \frac{\sum_{t=1}^{T-1}(\theta_t - \bar{\theta}^{(\cdot m)})(\theta_{t+1} - \bar{\theta}^{(\cdot m)})}{\sum_{t=1}^{T}(\theta_t - \bar{\theta}^{(\cdot m)})^2}.
\end{equation*}

We can interpret \(AC\)-values as a measure of stationarity. If \(AC\)-values are close to zero, there is no dependence between subsequent samples within imputation chains. Negative \(AC\)-values indicate divergence within imputation chains. \textbf{Subsequent sampled values within each imputation chain are less alike.} Positive \(AC\)-values indicate recurrence. If \(\theta\)-values of subsequent iterations are similar, trending may occur. Negative \(AC\)-values show no threat to the stationarity component of convergence. On the contrary even---negative \(AC\)-values indicate that \(\theta\)-values of subsequent iterations diverge from one another, which may increase the variance of \(\theta\) and speed up convergence. As convergence diagnostic, the interest is therefore in positive \(AC\)-values. The magnitude of \(AC\)-values may be evaluated statistically {[}\textbf{add reference}{]}.

\hypertarget{thresholds}{%
\subsubsection{Thresholds}\label{thresholds}}

Upon convergence, \(\widehat{R}=1\) and \(AC=0\), which are unlikely thresholds for MCMC algorithms, because of its convergence to a distribution. Even in the most converged state, the algorithm will show some signs of non-mixing and non-stationarity. Upon \emph{sufficient} convergence, the imputation chains will intermingle such that the only difference between the chains is caused by the randomness induced by the algorithm (\(\widehat{R} > 1\)), and there may be some dependency between subsequent iterations of imputation chains (\(AC > 0\)). In practice, non-convergence is usually diagnosed when \(\widehat{R}\) \textgreater{} 1.2 or 1.1 or even 1.01. And a t-test is performed to assess whether \(AC\) is significantly different from zero.

\hypertarget{in-practice}{%
\subsection{In practice}\label{in-practice}}

Before we evaluate the performance of the non-convergence diagnostics \(\widehat{R}=1\) and \(AC=0\) quantitatively through simulation, we apply them to the example of pathological non-convergence by \citet{buur18} that we reproduced above. We want to be able to draw the same conclusions as we did from visually inspecting the two algorithms. But \(\widehat{R}=1\) and \(AC=0\) should at least be able to distinguish between the two algorithms plotted in Figure \ref{fig:non-conv}. From visual inspection, we know that the typically converged algorithm initially portrayed some signs of non-mixing (around \(t=2\)), but intermigled nicely overall. Additionally, there was very little trending. The algorithm with pathological non-convergence showed severe non-mixing, although this gradually improved (from \(t=7\)). Moreover, there was a lot of trending as well initially (up-to \(t=6\)), after which the chains reached a somewhat more stationary state. To assess whether \(\widehat{R}\) and \(AC\) may be appropriate non-convergence identifiers for iterative imputation algorithms, they should identify the same trends as the visual inspection. But at least, the following condition must hold: the methods should indicate worse performance for the mis-specified model with pathological non-convergence than the typically converged algorithm (i.e., higher \(\widehat{R}\)- and \(AC\)-values).

In Figure \ref{fig:diagnostics}A, the chain means from Figure \ref{fig:non-conv} are plotted again---now together. Panel B shows two versions of calculating \(AC\)s applied to the \(\theta\)s of panel A: the default calculation with \texttt{R} function \texttt{stats::acf()} \citep{R}, and manual calculation as the correlation between \(\theta\) in iteration \(t\) and \(\theta\) in iteration \(t+1\). Panel C shows the traditional computation of \(\widehat{R}\), and panel D shows \(\widehat{R}\) as computed by implementing \citet{veht19} 's recommendations.

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/diagnostics-1}

}

\caption{Convergence diagnostics applied on the imputation algorithms of Figure 2. $\theta$ = chain mean in $y_{imp, m}$.}\label{fig:diagnostics}
\end{figure}

When we look at panel B, we conclude something weird. The \(AC\)-values calculated with the default function indicate equal performance for the typical convergence and the pathological non-convergence (up-to \(t=5\)), while there is obvious trending in the \(\theta\)s of the latter. Moreover, the best convergence (as indicated by the lowest \(AC\)-value) is observed at \(t=2\), but looking at the chain means in panel A, there should be some signs of trending up-to iteration number seven. After consulting the documentation on \texttt{stats::acf()} \citep{R}, we conclude that this \(AC\) function is not suited for iterative imputation algorithms. The function is optimized for performance when \(t\geq50\) \citep{box15}, while the default number of iterations in iterative imputation is often much lower. Therefore, we compute \(AC\) manually, see the solid line. These \(AC\) values indicate non-stationarity as expected. We therefore calculate \(AC\) manually.

\textbf{Describe panel C here and why it is not optimal{]}} The \(\widehat{R}\)-values in panel D do meet the requirements specified above. \(\widehat{R}\) as computed conform \citet{veht19} indeed indicate less signs of non-convergence as the number of iterations goes up.

\hypertarget{simulation-study}{%
\section{Simulation study}\label{simulation-study}}

The aim of the simulation study is to evaluate the impact of inducing non-convergence in the MICE algorithm on several quantities of scientific interest \(Q\). Subsequently we will evaluate how well \(\widehat{R}\) and \(AC\) perform in identifying the effects of non-convergence \textbf{{[}and performance of different thetas{]}}. And finally, we will formulate an informed advice on the requirements to safely assume \emph{sufficient} convergence in practice. That is, we assume that the algorithm is \emph{sufficiently} converged when each estimate \(\bar{Q}\) is an unbiased and confidence valid estimate of the corresponding estimand \(Q\). \textbf{{[}and formulate advice on thetas?{]}}

Non-convergence will be induced by: 1) increasing the missingness proportion \(p_{mis}\) in dataset \(y\) incrementally, and 2) terminating the imputation algorithm at a varying number of iterations \(T\). The first set of simulation conditions---the missingness proportions---is chosen to reflect the difficulty of the missingness problem. The underlying assumption is that low missingness proportions lead to quick algorithmic convergence, since there is a lot of information in the observed data. Higher missingness proportions should yield slower convergence. Unless, however, the fraction of missing information is so high that the random component in the imputation algorithm outweighs the information in the observed data. Then, convergence to a stable but highly variable state may be reached instanteneously. We assume that the incremental missingness proportions in our study will result in a corresponding increase in signs of non-convergence.

The assumption inherent to the second set of simulation conditions---the number of iterations---is that terminating the imputation algorithm too early causes non-convergence. Generally, the algorithm will not reach convergence if \(T=1\), because the imputed values in the first iteration (at \(t=1\)) depend on the initial values of the algorithm (which are sampled randomly from the set of observed datapoints). As the number of iterations increases, the imputation chains should become independent of the initial values, until the point at which adding an extra iteration does not lead to a more converged state. We assume that we can induce non-convergence at least in conditions where \(T\) is smaller than the default number of iterations in \texttt{mice}, \(T=5\).

\hypertarget{hypotheses}{%
\subsection{Hypotheses}\label{hypotheses}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  We expect that simulation conditions with a high missingness proportion \(p_{mis}\) and a low number of iterations \(T\) will result in biased, invalid estimates of the quantities of scientific interest, \(Q\)s.
\item
  We hypothesize that \(\widehat{R}\) and \(AC\) will correctly identify signs of non-convergence in those simulation conditions where the \(\bar{Q}\)s are \emph{not} unbiased and confidence valid estimates of the \(Q\)s.
\item
  We hypothesize that the recommended thresholds to diagnose non-convergence with \(\widehat{R}\) (\(\widehat{R} > 1.2\), \(\widehat{R} > 1.1\), and \(\widehat{R} > 1.01\)) may be too stringent for iterative imputation applications. In an empirical study, where \(\widehat{R}\) was used to inform the required imputation chain length, it took as many as 50 iterations to overcome the conventional non-convergence threshold \(\widehat{R}>1.2\). Yet, scientific estimates were insensitive to continued iteration from \(t=5\) onwards \citep{lace07}. We therefore suspect that \(\widehat{R}\) may over-estimate signs of non-convergence in iterative imputation algorithms, compared to the bias the estimates. In contrast to this, it may occur that signs of non-convergence are under-estimated by \(\widehat{R}\), in exceptional cases where the initial values of the algorithm are not appropriately over-dispersed \citep[p.~437]{broo98}. In e.g.~\texttt{mice}, initial values are chosen randomly from the observed data, hence we cannot be certain of over-dispersion in the initial values. In practice, we do not expect this to cause problems for identifying non-convergence with \(\widehat{R}\).
\item
  We expect that high \(AC\) values are implausible in iterative imputation algorithms with typical convergence. That is, after only a few iterations, the randomness induced by the algorithm should effectively mitigate the risk of dependency within chains.
\item
  \textbf{Something about thetas}
\end{enumerate}

\hypertarget{set-up}{%
\subsection{Set-up}\label{set-up}}

Convergence of the MICE algorithm is investigated through model-based simulation in \texttt{R} \citep[version 3.6.3;][]{R}. The simulation set-up is summarized in the pseudo-code below. The complete \texttt{R} script of the simulation study is available from \href{https://github.com/gerkovink/shinyMice/tree/master/3.Thesis/1.SimulationStudy}{github.com/gerkovink/shinyMice}.

\begin{verbatim}
# pseudo-code of simulation 
1. simulate data 
for (number of simulation runs from 1 to 1000)
 for (missingness proportions 5%, 25%, 50%, 75% and 95%)
  2. create missingness
  for (number of iterations from 1 to 100)
   3. impute missingness
   4. perform analysis of scientific interest
   5. compute non-convergence diagnostics 
   6. pool results across imputations
   7. compute performance measures
 8. combine outcomes of all missingness proportions
9. aggregate outcomes of all simulation runs 
\end{verbatim}

\hypertarget{data-generating-mechanism.}{%
\subsubsection{Data-generating mechanism.}\label{data-generating-mechanism.}}

In this study, sampling variance is not of interest. Therefore, a single complete dataset may serve as comparative truth in all simulation repetitions \citep{vink14}. The data-generating mechanism is a multivariate normal distribution, representing person-data on three predictor variables in a multiple linear regression problem \textbf{(from an unspecified social scientific field of study)}. Let the predictor space be defined as
\[
\begin{aligned}
\begin{pmatrix}X_1\\
X_2\\
X_3
\end{pmatrix} \sim \mathcal{N}
\begin{bmatrix}
\begin{pmatrix}
12\\
3\\
0.5
\end{pmatrix}\!\!,
\begin{pmatrix}
4 & 4 & 1.8 & 0\\
4 & 16 & 4.8 & 0\\
1.8 & 4.8 & 9 & 0
\end{pmatrix}
\end{bmatrix}\!\!\text{.}\\[2\jot]
\end{aligned}
\]
A finite population of \(N=1000\) is simulated using the \texttt{mvtnorm} package \citep{mvtnorm}. Subsequently, a fourth variable is constructed as outcome \(Y\). For each unit \(i = 1, 2,..., N\), let
\[
Y_i = 1 + 2X_{1i} +.5X_{2i} - X_{3i} + \epsilon_i ,
\]
where \(\epsilon \sim \mathcal{N}(0, 100)\). From the complete set, we obtain the true values of several quantities of scientific interest, \(Q\)s. \textbf{{[}MOVE THIS NEXT PART TO TOP OF SIMULATION SECTION?{]}} \(Q\)s are the targets that will be estimated in each simulation repetition, after first \emph{amputing} the complete data with varying missingness proportions, then imputing the missingness with a varying number of iterations, then estimating the \(Q\)s on the \(M\) completed datasets, and finally pooling the \(M\) estimates \(\hat{Q}\) to obtain a single \(\bar{Q}\) for each \(Q\). \textbf{We try to predict the effect of non-converegence (i.e., the difference between \(Q\) and \(\bar{Q}\)) with the convergence diagnostics \(\widehat{R}\) and \(AC\), by appliying these methods to several scalar summaries of the state space of the algorithm, \(\theta\)s.}

\hypertarget{scientific-estimands.}{%
\subsubsection{Scientific estimands.}\label{scientific-estimands.}}

We consider four types of \(Q\)s that are often of interest in empirical research. Namely, two descriptive statistics: the mean \(\mu_j\) and standard deviation \(\sigma_j\) of each variable \(j = Y, X_1, X_2, X_3\). We also consider the regression coefficients of the predictors: \(\beta_1\), \(\beta_2\), and \(\beta_3\) in regression equation
\[Y' = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \beta_3 X_3,\]
where \(Y'\) is the expected value of the outcome. And finally, the proportion of variance explained by the set of predictors: coefficient of determination \(r^2 \times 100\) \textbf{(note that lower case \(r\) is used to avoid confusion with non-convergence diagnostic \(\widehat{R}\))}. The true values of these scientific estimands are calculated on the complete data with the \texttt{R} functions \texttt{base::mean()}, \texttt{stats::sd()}, and \texttt{stats::lm()} \citep{R}.

\hypertarget{methods.}{%
\subsubsection{Methods.}\label{methods.}}

In each simulation repetition, we \emph{ampute} the complete dataset five times to obtain five different incomplete datasets, \(y\)s. The ratio between \(y_{obs}\) and \(y_{mis}\) in each \(y\) depends on the missingness proportion: \(p_{mis} =.05,.25,.5,.75,.95\). We ampute the complete data with the \texttt{mice} package \citep[function \texttt{mice::ampute()};][]{mice}. We consider all possible univariate and multivariate patterns of missingness, conform a `missing completely at random' missingness mechanism \citep{rubin87}. I.e., the probability to be missing is the same for all \(N \times k\) cells in \(y\), conditional on the missingness proportion \(p_{mis}\).

Missing datapoints in each incomplete dataset \(y\) are imputed with the \texttt{mice()} function \citep{mice}. All imputation procedures are performed with Bayesian linear regression imputation, and five imputation chains (\(M=5\)). Because \(M=5\), each run of the \texttt{mice()} function results in five sets of imputations: \(y_{imp,m}\), where \(m = 1, 2, ..., 5\). The number of iterations in each run of the algorithm varies between simulation conditions (\(T=1,2,...,100\)).

From the 5 imputations \(y_{imp,m}\) we obtain the chain means and chain variances for each variable \(j\). These will serve as scalar summary \(\theta\) to apply non-convergence diagnostics \(\widehat{R}\) and \(AC\) on. We then use \texttt{mice::complete()} to combine the imputed data \(y_{imp,m}\) with the observed data \(y_{obs}\), resulting in five completed datasets \{\(y_{obs}, y_{imp, m}\)\}.

On the 5 completed datasets \{\(y_{obs}, y_{imp, m}\)\}, we:

\begin{itemize}
\item
  Estimate \(\hat{Q}\)s for \(Q= \mu\) and \(Q=\sigma\) with \texttt{base::mean()}, and \texttt{stats::sd()} \citep{R}. We pool the \(\hat{Q}\)s with \texttt{base::mean()} to get \(\bar{Q}\)s.
\item
  Compute \(\lambda_1\), to use as \(\theta\) in the two convergence diagnostics. By definition, the first eigenvalue of the variance-covariance matrix, \(\lambda_1\), is equal to the variance of the principal component solution of each \{\(y_{obs}, y_{imp, m}\)\}, which we calculate with \texttt{stats::princomp()} \citep{R}.
\item
  Perform multiple linear regression with \texttt{stats::lm()} \citep{R} to get the \(\hat{Q}\)s for \(Q=\beta\) and \(Q=r^2\). We pool the \(\hat{Q}\)s conform \citet{vink14} to get \(\bar{Q}\)s for the \(\beta\)s and \(r^2\) \textbf{{[}check pooling of \(r^2\), instead of \texttt{mice::pool.r.squared()}{]}}. Additionally, we use the \(\hat{Q}\)s of \(Q=\beta\) as \(\theta\) to apply the convergence diagnostics to.
\end{itemize}

We apply non-convergence diagnostics \(\widehat{R}\) and \(AC\) to each scalar summary \(\theta\). We calculate \(\widehat{R}\) by implementing Vehtari et al.'s \citeyearpar{veht19} recommendations, and \(AC\) as the correlation between the \(t^{th}\) and the \((t+1)^{th}\) iteration.

\hypertarget{performance-measures.}{%
\subsubsection{Performance measures.}\label{performance-measures.}}

We assess the impact of inducing non-convergence in the iterative imputation algorithm by comparing \(\bar{Q}\)s with \(Q\)s. For each \(Q\), we calculate bias as \(\bar{Q} - Q\). For the regression coefficients \(Q=\beta_{1,2,3}\), we also compute the empirical coverage rate (CR). CR is defined as the percentage of simulation repetitions in which the 95\% confidence interval (CI) around \(\bar{Q}\) covers the true estimand \(Q\). Let
\[\text{CI} = \bar{Q} \pm t_{(M-1)} \times SE_{\bar{Q}},\]
where \(t_{(M-1)}\) is the quantile of a \(t\)-distribution with \(M-1\) degrees of freedom, and \(SE_{\bar{Q}}\) is the square root of the pooled variance estimate. If we obtain nominal coverage (CR = 95\%), we can conclude that the \(\bar{Q}\)s are conficence valid estimates of the \(Q\)s. Finally, we inspect CI width (CIW): the difference between the lower and upper bound of the 95\% confidence interval around \(\bar{Q}\). CIW is of interest because too short CIs indicate under-estimation of the variance in \(\bar{Q}\), which may yield spurious inferences.

With these performance measures, we also evaluate how well the non-convergence diagnostics \(\widehat{R}\) and \(AC\) can identify simulation conditions in which the \(\bar{Q}\)s are \emph{not} unbiased, confidence valid estimates of the \(Q\)s. \textbf{Rephrase and add corresponding \(Q\)s:} We measure the performance of \(\widehat{R}\) and \(AC\) by the four scalar summaries \(\theta\): chain means (i.e., the mean in each \(y_{imp, m, j}\)) versus \(Q=\mu\), chain variances (i.e., the variance in each \(y_{imp, m, j}\)) versus \(Q=\sigma\), the estimated regression coefficients \(\beta\) in each \{\(y_{obs}, y_{imp, m}\)\} (i.e., a user-defined statistic of scientific interest \(\hat{Q}\)) against \(Q=\beta\), and the first eigenvalue of the variance-covariance matrix in each \{\(y_{obs}, y_{imp, m}\)\}, \(\lambda_1\) versus \(Q=r^2\).

\hypertarget{results}{%
\section{Results}\label{results}}

\textbf{{[}Add to this section: 1) emphasize that the iterations in the plots are averages across repetitions, not within one run of MICE; 2) sample effects due to single complete dataset?? 3) Add more info about figure legends and axes 4) the ac and rhat are independent of missingness, not the thetas.{]}}
Primarily, we want to know what the effect is of inducing non-convergence on the estimated \(Q\)s. Specifically, we are interested in the bias of \(\bar{Q}\) for all \(Q\)s and the confidence validity of \(\bar{Q}\) for \(Q=\beta\). Therefore, we plot these performance measures against the simulation conditions (the number of iterations \(T\) and the missingness proportion \(p_{mis}\)). We look at averages across simulation repetitions.

Secondly, we want to know if conditions in which the \(Q\)s are affected by non-convergence are also identified by the non-convergance diagnostics \(\widehat{R}\) and \(AC\). We do this, again, by looking at the performance measures for the \(Q\)s, but now we evaluate these against what \(\widehat{R}\) and \(AC\) indicate about signs of non-convergence in the \(\theta\)s.

\textbf{{[}In short:{]}}
- Is there more bias in conditions where non-convergence was induced? (NOPE for univariates, YES for multivar.).

\begin{itemize}
\item
  Is the point of \emph{sufficient} convergence (at which the \(\bar{Q}\)s are unbiased and confidence valid estimates of the \(Q\)s) the same point as identified by the non-convergence diagnostics? (NOPE, they are too conservative).
\item
  Is one of the thetas better? (NAH, doesn't really matter, probably multivariate is better).
\item
  What should the thresholds be? (LOWER than recommended).
\item
  And the default nr of iterarions? (MAYBE increase to 10?).
\end{itemize}

For reasons of brevity, we only discuss the convergence diagnostics for the \(Q\)s with the worst performance in terms of bias. Bias in the descriptve statistics (\(Q=\mu\) and \(Q=\sigma\)) was most pronounced in the outcome variable \(Y\). Of the regression coefficients, \(Q=\beta\), the largest bias was observed in the effect of \(X_1\) on \(Y\), \(\beta_1\). Results for \(Q=r^2\) are reported as-is. In the figures, however, we only present simulation conditions where \(T\leq50\). Full results are available from \textbf{{[}link github repo{]}}.

\hypertarget{in-general}{%
\subsection{In general}\label{in-general}}

As expected, conditions with a higher proportion of missingness and/or a lower number of iterations show more signs of non-convergence. I.e., on average these conditions portray more extreme bias in the estimated \(Q\)s, and non-nominal coverages. Roughly speaking this means that MICE is indeed not converged at \(t=1\), and converges gradually as \(t\) increases to \(T\). The point at which an additional iteration does not lead to an improvement of the estimates depends on the difficulty of the missingness problem---between XYZ and XYZ. This means that the algorithm has converged sufficiently (\textbf{under the current specifications}).

As required for a method to diagnose non-convergence, the values of \(\widehat{R}\) follow a trend similar to the performance measures. \textbf{Add split between convergence diagnostics versus performance measures, and convergence diagnostics versus simulation conditions.} \(\widehat{R}\) values are generally lower in conditions with a higher number of iterations, and somewhat higher in conditions with a higher percentage of missing data. Autocorrelation values were indeed decreasing with a higher number of iterations, and lower in conditions with less missingness.

Evaluation with the performance measures shows that \(\widehat{R}\) is conservative: it identifies signs of non-convergence in conditions where the \(\bar{Q}\)s are unbiased and/or confidence valid estimates of the \(Q\)s.

\hypertarget{means}{%
\subsection{Means}\label{means}}

\textbf{Message: after one iteration (\(T\)=2), the algorithm is as well converged as the missingness proportion allows. Sooner even with lower missingness proportions (\(T\)=1 might be ok for \(p_{mis}\) = 5 and 25).} Number of it has little influence on bias in \(\hat{Q}\), only in first iteration some bias in missingness conditions of 25 and up. There is a clear different between the missingness conditions. The higher the missingness prop, the larger the magnitude of the bias. For this \(Q\), t=2 would be sufficient. And only for \(p_{mis}\) = 5 the \(\bar{Q}\)s are unbiased. All other \(p_{mis}\) conditions have bias irrespective of the number of iterations.

\(\widehat{R}\) differs a lot across iteration conditions. But not structurally for missingness proportions. Highest \(\widehat{R}\)s for the condition \(T\)=2, while the most bias is for (\(T\)=1), but there is no \(\widehat{R}\) for that by definition. But for missingness, 25\% seems the most signs of non-conv, while it has one of the least bias. Threshold of 1.2 reached in condition it = 3, \(p_{mis}\) = 5 and 75, and then for it\textgreater6, \(p_{mis}\) = 5 and 50, and \(T\)\textgreater10 for all \(p_{mis}\). Threshold 1.1 reached in in \(T\)=12, \(p_{mis}\) = all but 75, 1.01 only reached in \(T\) = 93, \(p_{mis}\) = 75. While for this \(Q\), \(T\)=2 would be sufficient.

\(AC\) mainly shows effects of the set of iteration conditions. \(AC\) indicates less signs of non-convergencein conditions with higher number of iterations. For iteration conditions where \(T\)\textless10, there are also some effects of the missingness proportions. The \(p_{mis}\) condtions 25 and 95 \% indicate the most signs of non-convergence. For \(p_{mis}\) = 95\% this corresponds to the bias in Qbar. For 25\% miss it doesn't. The thresholds defined by the critical value is only met in the conditions it=3, \(p_{mis}\) = 5,75, and 95\%.

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/mean-1}

}

\caption{Convergence diagnostics chain mean.}\label{fig:mean}
\end{figure}

\hypertarget{standard-deviations}{%
\subsection{Standard deviations}\label{standard-deviations}}

\textbf{Message: even with possible effect of initial values (\(T\)=1), the algorithm is as well converged as the missingness proportion allows.} For \(Q=\sigma_Y\) there is no effect visible of the iteration conditions, but clearly for the \(p_{mis}\) conditions. Most bias for \(p_{mis}\) = 95, followed by \(p_{mis}\) = 50 and 75. For the least possible bias we may use any number of iterations \(T\geq1\), but none of the \(p_{mis}\) results in unbiased \(\bar{Q}\)s.

There is a clear effect of the number of iterations on \(\widehat{R}\). Conditions with lower Ts have more signs of non-conv, as identified by \(\widehat{R}\). There is little difference between missingness conditions. Only for conditions where \(T<5\) the missingness proportions seem to affect whether \(\widehat{R}\) indicates non-convergence The bias in \(\bar{Q}\) seems unrelated to the magnitude of \(\widehat{R}\). Conditions with more bias (\(p_{mis}\) = 50,75,95) are not identified to have more signs of non-convergence. The thresholds are therefore useless (because depend mostly on \(T\), not so much \(p_{mis}\)).

\(AC\) is affected by both t and \(p_{mis}\) conditions. But conditions with worst bias (\(p_{mis}\) = 50, 75, 95) are not identified as worst converging according to AC. Moreover, only for \(T\) =3 the \(AC\) is significantly different from zero.

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/sd-1}

}

\caption{Convergence diagnostics chain variance.}\label{fig:sd}
\end{figure}

\hypertarget{regression-coefficients}{%
\subsection{Regression coefficients}\label{regression-coefficients}}

\textbf{Message: after four iterations (\(T\)=5), the algorithm is as well converged as the missingness proportion allows. Sooner even with lower missingness proportions (\(T\)=1 might be ok for \(p_{mis}\) = 5 and 25).} The figure for \(Q=\beta_1\) shows that the \(\bar{Q}\)s are affected by both the number of iterations and the missingness proportion. For \(T\) =1, the magnitude of the bias (almost) follows the order of the missingness proportions. More missingness causes more bias in the \(\bar{Q}\). In general, a lower nr of iterations results in more bias, but this depends on the missingness proportion. For \(p_{mis}\) = 95, there is more bias in conditions where \(T\)\textless6. When \(p_{mis}\) = 50, it's \(T\)\textless3. For \(p_{mis}\) = 75 and 25, it's only for \(T\)\textless2. And conditions where \(p_{mis}\) = 5 are unaffected on average. However, none of the conditions result in completely unbiased estimates of \(Q\).

We should see this in the conv diag: higher \(\widehat{R}\) and \(AC\) for \(p_{mis}\) = 75 and 95, and for condtions where \(T<2-6\), depending on \(p_{mis}\).

\(\widehat{R}\) depends somewhat on the missingness proportion. For conditions where \(T\)\textless10, the \(\widehat{R}\) values follow the \(p_{mis}\) more or less. But what we actually want to know is whether they are indicators of bias. Conditions with the worst bias (high \(p_{mis}\), low T) have the largest \(\widehat{R}\) values. The thresholds are not appropriate to diagnose sufficient convergence. All thresholds over-estimate the signs of non-convergence compared to the bias in Qbar. However, a higher threshold than 1.2 may result in terminating the algorithm after 3 iterations, which is insufficient for the \(p_{mis}\) = 95 condition. Threshold of 1.25 or 1.3 could work for the other \(p_{mis}\) conditions.

\(AC\) is very affected by the p\_mis. The two conditions with the most bias (\(p_{mis}\) = 75 and 95) are the only ones identified as non-converging, as they should. However, the \(p_{mis}\) = 75 condition is only identified in the \(T\) =3 condtion, while the bias is the least pronounced there. And the \(AC\) only identifies non-convergence in the \(p_{mis}\) = 95 conditions where \(T\)\textgreater10. In the default situation, \(T\) =5, this would be overlooked. Also, the \(AC\) shows the opposite of what you would expect from a theta that is based on the observed data: shouldn't there be higher autocorrelations in conditions with comparatively more observed data in the completed datasets? I don't understand this.

\textbf{Add how interesting it is because we literally track \(Q\) and are still not ok.}

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/est-1}

}

\caption{Convergence diagnostics regression coefficient.}\label{fig:est}
\end{figure}

\hypertarget{variance-explained}{%
\subsection{Variance explained}\label{variance-explained}}

\textbf{Message: after two iterations (\(T\)=3), the algorithm is as well converged as the missingness proportion allows. Sooner even with lower missingness proportions (\(T\)=1 might be ok for \(p_{mis}\) = 5 and 25).} Bias is the worst for conditions where \(p_{mis}\) = 95, \(T\) \textless3. Followed by \(p_{mis}\) = 75, \(T\) =1; and \(p_{mis}\) = 50, \(T\) = 1. Conditions where \(p_{mis}\) = 25 or 5 seem unaffected by the number of iterations. Minimal bias is obtained in all iteration conditions. We should see this in \(\widehat{R}\) and AC: worst convergence in \(p_{mis}\) = 95 and 50; some non-convergencein conditions where \(p_{mis}\) = 75 and \(T\) =1, and no non-convergencein conditions where \(p_{mis}\) = 25 or 5. The sign of the bias is logical: when there is less info in the data, the relations between the predictors and the outcome is under-estimated.

\(\widehat{R}\) behaves oppositely to what we wanted. The lowest \(\widehat{R}\) values for conditions where \(T\)\textless10 are observed for the two missingness conditions with the worst bias. However, the highest \(\widehat{R}\) observed is indeed for the missingness condition with the worst bias, \(p_{mis}\) =95. But only at \(T\) =5 (while bias is no worse than anywhere else in this \(p_{mis}\) condition, conditional on iterations condition \(T\)\textgreater2). So neither of the thresholds works perfectly.

\(AC\) shows a similar trend to the biases: more signs of non-convergencefor higher \(p_{mis}\) and lower T. The threshold, however, performs less well. It identifies signs of non-convergence for all conditions where \(T=3\), irrespective of p\_mis. Then nothing, and then for \(p_{mis}\) =95, \(T\)\textgreater7, and \(p_{mis}\) = 50, \(T\) \textgreater14, and \(p_{mis}\) = 75, \(T\)\textgreater20. This does not corresond to the bias in \(\bar{Q}\). Maybe use AC\textgreater.5 at \(T\) =5 as threshold?

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/pred-1}

}

\caption{Convergence diagnostics regression coefficient.}\label{fig:pred}
\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

In short, univariate estimates seem robust against terminating the algorithm early: There is no clear effect of the number of iterations on the bias in the estimated means and standard deviations. The bias in these \(Q\)s only depends on the \(p_{mis}\), with higher \(p_{mis}\) resulting in more bias. Yet, the convergence diagnostics indicate that the algorithm did not reach a completely converged state (yet).

\(\widehat{R}\) and autocorrelation indicate that algorithmic convergence may only be reached after twenty or even forty iterations, while unbiased, confidence valid estimates may be obtained with as little as one iteration. These results are in agreement with the simulation hypothesis: \(\widehat{R}\) over-estimates the severity of non-convergence when applied to MI procedures.

With that, we show that MICE can lead to correct outcomes when they have not converged according to two common conv diags. This may be due to the methods (and their thresholds) or due to the \(Q\)s (descriptives and multivariate linear regression, not higher dimensional/more complex \(Q\)s). More `complicated' \(Q\)s (e.g., higher-order effects or variance components) might show bias, under- or over-coverage at higher \(T\), as indiated by \(\widehat{R}\) and AC. Add what \%miss has to do with it.

\hypertarget{recommendations-for-empirical-researchers}{%
\subsection{Recommendations for empirical researchers}\label{recommendations-for-empirical-researchers}}

For empirical researchers:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Check trace plots for pathological non convergence and adjust imputation model if necessary.
\item
  use \(\widehat{R}\) wait 1.1 ash threshold and autocorrelation with \textbf{?} as threshold. Keep iterating until these thresholds are reached.
\item
  Do not use the R function ACF. Instead, compute autocorrelations manually \textbf{(see e.g., {[}GitHub link{]})}.
\item
  Track your own scalar summary of interest. This is somewhat advanced but explained in Van Buren 2018. Compute \(\widehat{R}\) and autocorrelation values for this scalar summary.
\item
  \textbf{Something} about the novel \(\theta\) that is `substantive model-independent'.
\end{enumerate}

\hypertarget{recommendations-for-future-research}{%
\subsection{Recommendations for future research}\label{recommendations-for-future-research}}

Further research is needed to investigate their performance under clear violation of convergence, e.g.~dependency between predictors (predictors with very high correlations). Also:

\begin{itemize}
\item
  Induce non-convergence with mis-specified imputation model.
\item
  More difficult missingness problems, i.e., M(N)AR instead of MCAR. Proper performance under a `missing completely at random' missingness mechanism is necessary to demonstrate the appropriateness of \(\widehat{R}\) and \(AC\) as non-convergence diagnostics. However, results may not be extrapolated to other missingness mechanisms.
\item
  Imputation technique with questionable convergence properties, e.g., PMM.
\item
  Apply recommendations to empirical data and write a vignette for applied researchers.
\item
  Implement recommendations in ShinyMICE.
\item
  Develop an additional convergence diagnostic unique for iterative imputation: \(AC\) across imputations, instead of iterations.
\end{itemize}

Uitsmijter: To conclude, in the case of drawing inference from incomplete data, convergence of the iterative imputation algorithm is often a convenience but not a necessity.

\hypertarget{grading}{%
\section{Grading}\label{grading}}

\begin{itemize}
\item
  Scientific contribution
  (clear research question, convincing argumentation that the study adds to the scientific literature)
\item
  Theory
  (use of theory, clear hypotheses, relevant literature)
\item
  Method
  (correct data collection/handling, measures, choice and description of method of analysis)
\item
  Results
  (correct interpretation of findings, clear presentation)
\item
  Discussion and conclusions
  (summary, broader implications, limitations, future research)
\item
  Written presentation
  (writing, structure, tables and figures, references)
\end{itemize}

\bibliographystyle{sageh}
\bibliography{thesis.bib}

\end{document}