### 2020-05-07 3.Thesis/Figures/DiagramMI.Rnw
Feedback Stef

\documentclass{standalone}
%% recommended packages

%% more packages 
\usepackage{amsmath, tikz} % for table, equations, and flow chart figure (respectively)
\usetikzlibrary{fit, positioning, shapes} % for flow chart figure

\begin{document}

\begin{tikzpicture}
\tikzstyle{main}=[rectangle, minimum width = 22mm, minimum height = 9mm, thick, draw =black!80, node distance = 14mm]
\tikzstyle{connect}=[draw, -latex, thick, out=180,in=270, looseness=0]
\node[main, fill = white!100] (data) [label=below:Incomplete data] {$\{y_{\rm obs}, y_{\rm mis}\}$};
\node[main] (mids) [right=of data] { $\{y_{\rm obs}, \dot y_{\rm imp, \ell}\}$};
\node[main] (mids2) [above=10mm of mids] {$\{y_{\rm obs}, \dot y_{\rm imp, \ell}\}$ };
\node[main] (mids3) [below=10mm of mids,label=below: $m$ completed datasets] { $\{y_{\rm obs}, \dot y_{\rm imp, \ell}\}$};
\node[main] (mira) [right=of mids] {$\hat{Q}_\ell$};
\node[main] (mira2) [above=10mm of mira] { $\hat{Q}_\ell$ };
\node[main] (mira3) [below=10mm of mira,label=below: $m$ results] { $\hat{Q}_\ell$ };
\node[main, fill = white!100] (mipo) [right=of mira,label=below:Pooled result] {$\bar{Q}$ };
\path      
    (data.east) edge [connect] node [right] {} (mids.west)
(data.east) edge [connect] node [right] {} (mids2.west)
(data.east) edge [connect] node [right] {} (mids3.west)
      (mids.east) edge [connect] node [right] {} (mira.west)
      (mids2.east) edge [connect] node [right] {} (mira2.west)
      (mids3.east) edge [connect] node [right] {} (mira3.west)
		(mira.east) edge [connect] node [right] {} (mipo.west)
		(mira2.east) edge [connect] node [right] {} (mipo.west)
		(mira3.east) edge [connect] node [right] {} (mipo.west)
;
\end{tikzpicture}
%\caption{Scheme of the main steps in multiple imputation \citep[$m = 3$; adapted from][\S~1.4.1]{buur18}. {\footnotesize Missing data in dataset $Y$ is `imputed' (i.e., filled in) $m$ times. The imputed data is combined with the observed data ($Y_{obs}$) to create $m$ completed datasets. On each completed dataset, the analysis of scientific interest (or `substantive model') is performed. The quantity of scientific interest (e.g., a regression coefficient) is denoted with $Q$. Since $Q$ is estimated on each completed dataset, $m$ separate $\hat{Q}$-values are obtained. These $m$ values are combined into a single pooled estimate $\bar{Q}$. $\{y_{obs}, y_{mis}\}$ }}

\end{document}