### 2020-05-09 3.Thesis/Writeup/2.Manuscript.Rmd
Missing only some parts of the methods and discussion

---
title: "Missing the Point: Non-Convergence in Iterative Imputation Algorithms"
runninghead: Oberman
author:
  name: H. I. Oberman
  num: 1
address:
  num: 1
  org: Department of Methodology and Statistics, Utrecht University, Utrecht, The Netherlands
corrauth: "Hanne Oberman, Sjoerd Groenman building, Utrecht Science Park, Utrecht, The Netherlands."
email: h.i.oberman@uu.nl
abstract: "Iterative imputation is a popular tool to accommodate missing data. While it is widely accepted that valid inferences can be obtained with this technique, these inferences all rely on algorithmic convergence. There is no consensus on how to evaluate the convergence properties of the method. This paper provides insight into identifying non-convergence of iterative imputation algorithms. We conclude that in the case of drawing inference from incomplete data, convergence of the iterative imputation algorithm is often a convenience but not a necessity."
keywords: missing data, iterative imputation, non-convergence, mice
classoption:
 - Royal
 - times
bibliography: thesis.bib
bibliographystyle: sageh
output:
  bookdown::pdf_book:
    base_format: rticles::sage_article
    keep_tex: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 4.9, echo = FALSE, fig.align = "center") 
# knitr::opts_chunk$set(fig.width = 4.8, fig.height = 3.6, echo = FALSE, fig.align = "center")

# packages
library(mice)
library(dplyr)
library(ggplot2)
library(patchwork)

# load example non-convergence and figures
load("../Results/example_nonconv.Rdata")
load("../Results/example_conv.Rdata")
load("../Results/example_diagnostics.Rdata")
source("../3.PlotExampleNon-convergence.R")

# load results and figures
load("../Results/complete.Rdata") 
results <- results %>% filter(t<51)
source("../4.PlotResults.R")

# default graphing behavior
theme_update(
 plot.title = element_text(hjust = 0.5),
 plot.subtitle = element_text(hjust = 0.5),
 panel.border = element_blank(),
 panel.grid.major = element_blank(),
 panel.grid.minor = element_blank(),
 panel.background = element_blank(),
 axis.line = element_line(colour = "black"),
 legend.key = element_blank(),
 legend.position = "bottom",
 legend.margin = margin(0,0,0,0),
 # adjust text size
 text = element_text(size = 8),
 title = element_text(size = 8),
 legend.title = element_text(size = 7)
)
```

<!-- # Overview of sections -->

<!-- Intro -->

<!-- - This is missingness -->

<!-- - This is how we solve missingness -->

<!-- - For that we need an algorithm, but what if it has not converged? -->

<!-- - That is what we solve in this paper -->

<!-- Algorithmic convergence -->

<!-- - What is algorithmic convergence? -->

<!-- - When is it converged? Stationarity versus mixing. -->

<!-- - How is convergence currently investigated? And how should it be investigated? Multivariate versus univariate. -->

<!-- Methods -->

<!-- - Which methods are there? And how do they relate to mixing and stationarity? -->

<!-- - $\widehat{R}$ -->

<!-- - AC -->

<!-- - We can use these on any $\theta$ -->

<!-- Then Simulation study, Results, Discussion. -->

# Introduction

Anyone who analyzes person-data may run into a missing data problem. Missing data is not only ubiquitous across science, but treating it can also be a tedious task. If a dataset contains just one incomplete observation, calculations are not defined, and statistical models cannot be fitted to the data. To circumvent this, many statistical packages employ list-wise deletion by default (i.e., ignoring incomplete observations). Unfortunately, this *ad hoc* solution may yield invalid results [@buur18]. An alternative is to apply imputation. With imputation, we 'fill in' the missing values in an incomplete dataset. Subsequently, the model of scientific interest can be fitted to the completed dataset. By repeating this process several times, a distribution of plausible results may be obtained, which reflects the uncertainty in the data due to missingness. This technique is known as 'multiple imputation' [MI; @rubin76]. MI has proven to be a powerful technique to draw valid inferences from incomplete data under many circumstances [@buur18].

Figure \ref{fig:diagram} provides an overview of the steps involved with MI. The missing part $y_{\rm mis}$ of an incomplete dataset is imputed $m$ times. This creates $m$ sets of imputed data $\dot y_{\rm imp, \ell}$, where $\ell = 1, 2, ..., m$. The imputed data is then combined with the observed data $y_{\rm obs}$ to create $m$ completed datasets. On each of these datasets the analysis of scientific interest is performed to estimate $Q$: the quantity of scientific interest (e.g., a regression coefficient). Since $Q$ is estimated on each completed dataset, $m$ separate $\hat{Q}_\ell$-values are obtained. Finally, the $\hat{Q}_\ell$-values are combined into a single pooled estimate $\bar{Q}$. The premise of multiple imputation is that $\bar{Q}$ is an unbiased and confidence-valid estimate of the true---but unobserved---scientific estimand $Q$ [@rubin96].

<!-- $Q$ is quantity of scientific interest: "A scientific estimand $Q$ is a quantity of scientific interest that we can calculate if we would observe the entire population" [@buur18, par 2.3.1] -->

```{r diagram, out.width="\\linewidth", include=TRUE, fig.cap="Scheme of the main steps in multiple imputation---from an incomplete dataset, to $m=3$ multiply imputed datasets, to $m=3$ estimated quantities of scientific interest $\\hat{Q}$s, to a single pooled estimate $\\bar{Q}$."}
knitr::include_graphics("../Figures/DiagramMI.pdf") #adapted from [@buur18] par. 1.4.1.
```

A popular method to obtain imputations  $\dot y_{\rm imp, \ell}$ is to use the 'Multiple Imputation by Chained Equations' algorithm, shorthand 'MICE'[@mice]. With MICE, imputed values are drawn from the posterior predictive distribution of the missing values $y_{\rm mis}$. The algorithm is named after the iterative algorithmic procedure by which imputed values are generated: a multivariate distribution is obtained by iterating over a sequence of univariate imputations. The iterative nature of algorithms like MICE introduces a potential threat to the validity of the imputations: What if the algorithm has not converged? Are the imputations then to be trusted? And can we rely on the inference obtained on the completed data?

These remain open questions, since the convergence properties of iterative imputation algorithms have not been systematically studied [@buur18]. There is no scientific consensus on how to evaluate the convergence of imputation algorithms [@taka17]. Moreover, the behavoir of such algorithms under certain default imputation models (e.g., 'predictive mean matching') is an entirely open question [@murr18]. **Therefore, algorithmic convergence should be monitored carefully, **although not straightforward. Iterative imputation algorithms such as MICE are special cases of Markov chain Monte Carlo (MCMC) methods. In MCMC methods, convergence is not from a scalar to a point, but from one distribution to another. The values generated by the algorithm (e.g., imputed values) will vary even after convergence [@gelm13]. Since MCMC algorithms do not reach a unique point at which convergence is established, diagnostics  may only identify signs of *non*-convergence [@hoff09]. Several of such non-convergence diagnostics exist, but it is not known whether these are appropriate for iterative imputation algorithms. 
<!-- "a weak diagnostics is better than no diagnostic at all" [@cowl96]. -->

In this paper, we study different methods for assessing non-convergence in iterative imputation algorithms. We define several diagnostics and evaluate how these diagnostics could be appropriate for iterative imputation applications. We then adress the impact of inducing non-convergence in iterative imputation algorithms through model-based simulation in `R` [@R]. The aim of the simulation study is to determine whether unbiased, confidence-valid inferences may be obtained if the algorithm has not (yet) converged. And additionally, to formulate an informed advice on when approxiamte convergence may be safely concluded. We will therefore assess the performance of different methods to identify non-convergence. We translate the results of the study into guidelines for applied researchers.  **ADD that these guidelines may aid to draw valid inference from incomplete data?]** For reasons of brevity, we only focus on the iterative imputation algorithm implemented in the popular `mice` package [@mice] in `R` [@R].

<!-- The results of this study are guidelines for identifying non-convergence in iterative imputation algorithms, which will aid applied researchers in drawing valid inference from incomplete datasets. -->

**[Include sub-questions and make structure of the paper clear!]** How can non-convergence be identified diagnostically? Are common MCMC non-convergence diagnostics appropriate for MICE? And if so, which threshold should be used to diagnose non-convergence? How many iterations are sufficient/needed to be able to diagnose non-convergence? Are the default number of iterations sufficient (i.e., 5 in mice, 10 in SPSS and Stata, 30 in mi)? How severe is it when the algorithm has not converged? And what are guidelines for practice? Can the parameter of interest $Q$ be correct when the algorithm is not (yet) converged, and vice versa?

<!-- The defaults of common MI software: SPSS = 10 [(link)](https://www.ibm.com/support/knowledgecenter/SSLVMB_24.0.0/spss/mva/syn_multiple_imputation_impute.html), Mplus = 100?? [(link)](https://pdfs.semanticscholar.org/e20e/29e008592cbfbaa567931f74cdfdb5451405.pdf?_ga=2.55354671.54033656.1584698748-527613517.1584698748), Stata = 10 [(link)](https://www.stata.com/manuals13/mi.pdf, p. 139), Amelia = NA, because re-sampling, not convergence [(link)](https://cran.r-project.org/web/packages/Amelia/Amelia.pdf), en MI = 30 [(link)](https://cran.r-project.org/web/packages/mi/mi.pdf). -->

<!-- The goal of this research project is to develop novel methodology and guidelines for evaluating MI methods. These tools will subsequently be implemented in an interactive evaluation framework for multiple imputation, which will aid applied researchers in drawing valid inference from incomplete datasets. -->

## Some notation

Let $y$ denote an $n \times k$ matrix containing the data values on $k$ variables for all $n$ units in a sample. The data value of unit $i$ ($i = 1, 2, \dots, n$) on variable $j$ ($j = 1, 2, \dots, k$) may be either observed or missing. The number of units $i$ with at least one missing value, divided by the total number of units $n$, is called the missingness proportion $p_{\rm mis}$ in dataset $y$. The collection of observed data values in $y$ is denoted by $y_{\rm obs}$; the missing part of $y$ is referred to as $y_{\rm mis}$. For each datapoint in $y_{\rm mis}$, we sample $m \times T$ plausible values, where $m$ is the number of imputations ($\ell = 1, 2, \dots, m$) and $T$ is the number of iterations ($t = 1, 2, \dots, T$) in the imputation algorithm. The state space of the algorithm at a certain iteration $t$ may be summarized by a scalar summary $\theta$ (e.g., **the average of the imputed values**). The collection of $\theta$-values between the **initial value** (at $t=1$) and the final imputed value (at $t=T$) will be referred to as an 'imputation chain'. 
<!-- **[If not defined elsewhere, add: $\theta$s are scalar summaries of the algorithm at a certain iteration (e.g., chain means; the average of the imputed values in each imputation chain).] ** -->

# Algorithmic non-convergence

There are two requirements for convergence of iterative algorithms: mixing and stationarity [@gelm13]. In iterative imputation algorithms mixing implies that imputation chains intermingle nicely, and stationarity is characterized by **the absence of trending** within chains. If one of these two requirements is not met, there is non-convergence in the algorithm. Without mixing, chains may be 'stuck' at a local optimum, instead of sampling imputed values from the entire predictive posterior distribution of the missing values. The distribution of imputed values then differs across imputations. This may cause under-estimation of the variance between chains, which results in spurious, invalid inferences. Without stationarity, there is trending within imputation chains. Trending implies that further iterations would yield a systematically lower or higher set of imputations. Iterative imputation algorithms that have not (yet) reached stationarity, may thus yield biased estimates.

<!-- When we declare approximate convergence, we are actually concluding that each individual sequence appears stationary and that the observed sequences -->
<!-- have mixed well with each other. @gelm13, p. 288-->

<!-- **The goal of imputation through e.g. MICE is to obtain a multivariate distribution by iterating over a sequence of univariate imputations. Traditionally, the inspection of the multivariate convergence is done by visually inspecting the univariate chain means and variances over the iterations. The downside to these approaches is that they either focus on the univariate state space, or primarily track the change over the iterations of a multivariate outcome conform the scientific model of interest. Focusing on the convergence of these outcome parameters may influence this procedure in the sense that the model of evaluation favors the model of interest. Ideally, one would like to evaluate a model-independent parameter that summarizes the multivariate nature of the data. We propose eigenvalue as such a parameter, because it.... ** -->

To illustrate what non-mixing and non-stationarity looks like in iterative imputation algorithms, we reproduce an example from van Buuren [-@buur18, $\S$ 6.5.2]. Figure \ref{fig:non-conv} displays two scenarios from the example. The graph on the left-hand side of the figure shows typical convergence of an iterative imputation algorithm. The right-hand side displays pathological non-convergence, induced by purposefully mis-specifying the imputation model. Each line portrays one imputation chain: scalar summary $\theta$ plotted against the iteration number $t$. The $\theta$ in this example is the chain mean of a variable that we will call $j$ (i.e., the average of the imputed values for $j$ in each imputation $\dot y_{\rm imp, \ell}$). **In the typical convergence scenario, the imputation chains intermingle nicely and there is little to no trending. In the non-convergence scenario, there is a lot of trending and some chains do not intermingle.** Importantly, the chain means at the last iteration (the imputed value per imputation $\ell$) are very different between the two plots. The algorithm with the mis-specified model yields imputed values that are on average a factor two larger than those of the typically converged algorithm. Non-convergence thus influences the mean of $j$ in $\dot y_{\rm imp, \ell}$. This difference (presumably) translates to the $m$ completed data estimates $\hat{\mu}_{j,\ell}$, and consequently to the pooled estimate $\bar{\mu}_j$. Ergo, non-convergence leads to biased estimates. This shows the importance of reaching converged states in iterative imputation algorithms.

```{r non-conv, out.width='.49\\linewidth', fig.width=3.5, fig.height=3,fig.show='hold',fig.cap="Typical convergence versus pathological non-convergence. Please note that this is the same data with a different imputation model, leading to different imputations (e.g., see the units on the y-axis)."}
par(mar = c(2, 2, 0.1, 0.1))
con_plot
non_plot
```

**[MERGE with sections below]** Currently, the recommended practice for evaluating the convergence of the MICE algorithm is through visual inspection. After running the imputation algorithm for a certain number of iterations, its convergence is monitiored by plotting a scalar summary of the state space of the algorithm $\theta$ against the iteration number. Typically, the $\theta$s under evaluation are chain means and variances---the average and the variance of the imputed values per imputation. Monitoring convergence by inspecting traceplots of these $\theta$s may be undesirable for several reasons: 1) it may be challenging to the untrained eye, and 2) only severely pathological cases of non-convergence may be diagnosed [@buur18, $\S$ 6.5.2] and, 3) there is not an objective measure that quantifies convergence.

**[MERGE with section above and below]** Inspecting non-convergence through traceplots of chain means and chain variances may be insufficient because they are univariate summaries of the state space of the algorithm, while MICE is not only concerned with a single column, but the entire multivariate distribution of $y_{\rm imp}$. Ideally, a multivariate $\theta$ should be monitored. A suggestion by @buur18 for a multivariate $\theta$ to monitor is the quantity of scientific interest $Q$ (e.g., a regression coefficient), which usually is multivariate in nature. Implementing this, however, might be somewhat too technical for empirical researchers. Moreover, this scalar summary is not model-independent, i.e., it only applies to one model of scientific interest. Yet, one of the advantages of MI is that the missing data problem and scientific problem are solved independently. On these grounds, such a $\theta$ can be considered insufficient too. Van Buuren [-@buur18, $\S$ 4.5.2] also proposed multivariable evaluation of the MICE algorithm through eigenvalue decomposition, building on the work of @mack03. Eigenvalues of a variance-covariance matrix are a measure of the data's total covariance. Such a $\theta$ would be a model-independent, multivariate scalar summary to monitor, but it is not implemented in `mice`. 
<!-- Recommendation: "the different streams should be freely intermingled with one another, without showing any definite trends". -->

**[MERGE with sections above]** Convergence has historically been inspected visually, by monitoring the imputation algorithm over iterations. This is typically done through traceplots. In a traceplot, the iteration number is plotted against a certain $\theta$. $\theta$s are scalar summaries of the state space of the algorithm at a specific iteration. The default scalar summaries in `mice` are chain means and chain variances. Additionally, researchers may "monitor some statistic of scientific interest" [@buur18, $\S$ 6.5.2]. As @buur18 describes, researchers can specify their quantity of scientific interest $Q$ (e.g., a regression coefficient) as scalar summary $\theta$. Such a user-defined scalar summary, however, may be somewhat advanced for empirical researchers. And moreover, it is not universal to all complete data problems. Focusing on the convergence of outcome parameters may influence the iterative imputation procedure in the sense that the model of evaluation favors the model of interest.  The downside to these two approaches is that they either focus on the univariate state space, or primarily track the change over the iterations of a multivariate outcome conform the scientific model of interest. Ideally, one would like to evaluate a model-independent parameter that summarizes the multivariate nature of the data. Therefore, we propose a $\theta$ that summarizes the multivariate state space of the algorithm, but is independent from the model of scientific interest. We propose $\lambda_{1, \ell}$ as such a scalar summary. We define $\lambda_{1, \ell}$ as the first eigenvalue of the variance-covariance matrix of the completed data. Let $\lambda_{1, \ell} \geq \lambda_{2, \ell} \geq ... \geq \lambda_{j, \ell}$ be the eigenvalues of $\Sigma_\ell$ in each of the $m$ completed datasets {$y_{\rm obs}, \dot y_{\rm imp, \ell}$}. $\lambda_{1, \ell}$ is measure that summarizes the covariances in the completed datasets. **The first eigenvalue has the appealing property that is not dependent on the substantive model of interest. Eigenvalue decomposition is inspired by @mack03.**

<!-- **Since convergence of iterative imputation algorithms is in distribution, there is not a unique point at which the algorithm reaches a converged state. We can only evaluate scalar summaries of the multivariate state space of the algorithm, $\theta$s. The recommended scalar summaries to evaluate are chain means and chain variances. Additionally, researchers may "monitor some statistic of scientific interest" [@buur18, $\S$ 6.5.2]. Because of the issues mentioned above, we propose a novel $\theta$ to monitor: the first eigenvalue of the variance-covariance matrix of the completed data. Let $\lambda_{1, \ell} \geq \lambda_2 \geq ... \geq \lambda_j$ be the eigenvalues of $\Sigma$ in each of the $m$ completed datasets ${y_{\rm obs}, y_{\rm imp}}$. $\lambda_{1, \ell}$ is measure that summarizes the covariances in the completed datasets.**  -->

# Non-convergence diagnostics

<!-- Requirements for diagnostics to identify non-convergence in iterative imputation are: 1) work for a relatively low number of iterations (defaults in popular iterative imputation software between 5 and 30); 2) may be applied to different $\theta$s, including model-independent $\theta$s (e.g., chain mean, or eigenvalue). -->

Non-convergence diagnostics for MCMC algorithms typically identify problems in either one of the two requirements for convergence: mixing or stationarity. Non-stationarity within chains may be diagnosed with e.g., autocorrelation [$AC$; @scha97; @gelm13], numeric standard error ['MC error'; @gewe92], or Raftery and Lewis's [-@raft91] procedure to determine the effect of trending on the precision of estimates. A widely used diagnostic to monitor mixing between chains is the potential scale reduction factor $\widehat{R}$ ['Gelman-Rubin statistic'; @gelm92]. With a recently proposed adaptation, $\widehat{R}$ might also serve to diagnose non-stationarity, but this has not yet been thoroughly investigated [@veht19]. Therefore, we use $\widehat{R}$ **[WHICH? Explain that we'll use both]** and $AC$ to evaluate mixing and stationarity separately, as recommended by e.g., @cowl96.
<!-- @cowl96, p. 898. -->

**[Add: we don't use effective sample size and MC error because they assume independent samples *within* chains. In MI, we only use the final estimate. Also, computing n_eff assumes infinitely many iterations, which is fine in Bayesian analyses where T is often at least one thousand. But the default in iterative imp is often substantially lower.]**

## Potential scale reduction factor

The potential scale reduction factor $\widehat{R}$ was coined by @gelm92. An updated version of $\widehat{R}$ has been proposed by Vehtari et al. [-@veht19, p. 5]. **This version is better suited to detect non-convergence in the tails of distributions, by using three transformations on the scalar summary that it is applied to. Namely, rank-normalization, folding, and localization. The Vehtari et al. version of $\widehat{R}$ may be suitable for iterative imputation. We therefore use their definition of the diagnostic.** Let $m$ be the total number of chains, $T$ the number of iterations per chain (where $T\geq2$), and $\theta$ the scalar summary of interest. For each chain ($\ell = 1, 2, \dots, m$), we estimate the variance of $\theta$, and average these to obtain within-chain variance $W$.

\begin{align*}
W&=\frac{1}{m} \sum_{\ell=1}^{m} s_{\ell}^{2}, \text { where } s_{\ell}^{2}=\frac{1}{T-1} \sum_{t=1}^{T}\left(\theta^{(t \ell)}-\bar{\theta}^{(\cdot \ell)}\right)^{2}. 
\end{align*}
<!-- %\text{ \cite[p.~5]{veht19}} -->

\noindent We then estimate between-chain variance $B$ as the variance of the collection of average $\theta$s per chain. **[Note that this is not the usual B in MI]**

\begin{align*}
B&=\frac{T}{m-1} \sum_{\ell=1}^{m}\left(\bar{\theta}^{(\cdot \ell)}-\bar{\theta}^{(\cdot \cdot)}\right)^{2}, \text { where } \bar{\theta}^{(\cdot \ell)}=\frac{1}{T} \sum_{t=1}^{T} \theta^{(t \ell)} \text{, } \bar{\theta}^{(\cdot \cdot)}=\frac{1}{m} \sum_{\ell=1}^{m} \bar{\theta}^{(\cdot \ell)}. 
\end{align*}
<!-- %\text{ \cite[p.~5]{veht19}} -->

\noindent From the between- and within-chain variances we compute a weighted average, $\widehat{\operatorname{var}}^{+}$, which **approximates** the total variance of $\theta$. $\widehat{R}$ is then obtained as a ratio between the total variance and the within-chain variance:

\begin{equation*}
\widehat{R}=\sqrt{\frac{\widehat{\operatorname{var}}^{+}(\theta | y)}{W}},
\text{ where } \widehat{\operatorname{var}}^{+}(\theta | y)=\frac{T-1}{T} W+\frac{1}{T} B.
\end{equation*}

We can interpret $\widehat{R}$ as potential scale reduction factor since it indicates by how much the variance of $\theta$ could be shrunken down if an infinite number of iterations per chain would be run [@gelm92]. The assumption underlying this interpretation is that chains are 'over-dispersed' at $t=1$, and reach convergence as $T \to \infty$. Over-dispersion implies that the initial values of the chains are 'far away' from the target distribution and each other. When the sampled values in each chain are independent of the chain's initial value, the mixing component of convergence is satisfied. The variance between chains $B$ is then equivalent to the variance within chains $W$, and $\widehat{R}$-values will be close to one. High $\widehat{R}$-values thus indicate non-convergence.

## Autocorrelation

Autocorrelation is defined as the correlation between two subsequent $\theta$-values within the same chain [@lync07, p. 147]. In this study, we only consider $AC$ at lag 1, i.e., the correlation between the $t^{th}$ and $t+1^{th}$ iteration of the same chain. Following the same notation as for $\widehat{R}$,

\begin{equation*}
AC = \left( \frac{T}{T-1} \right) \frac{\sum_{t=1}^{T-1}(\theta_t - \bar{\theta}^{(\cdot m)})(\theta_{t+1} - \bar{\theta}^{(\cdot m)})}{\sum_{t=1}^{T}(\theta_t - \bar{\theta}^{(\cdot m)})^2}.
\end{equation*}

We can interpret $AC$-values as a measure of non-stationarity. If there is dependence between subsequent $\theta$-values in imputation chains, $AC$-values are non-zero. Positive $AC$-values occur when $\theta$-values are recurring (i.e., high $\theta$-values are followed by high $\theta$-values, and low $\theta$-values are followed by low $\theta$-values). Recurrence within imputation chains may lead to trending. Negative $AC$-values occur when $\theta$-values of subsequent iterations are less similar, or diverge from one another. Divergence within imputation poses no threat to the convergence of the algorithm---it may even speed up convergence. **Complete stationarity is reached when $AC=0$.** As non-convergence diagnostic, our interest is in positive $AC$-values.

<!-- close to zero indicate  independence between subsequent $\theta$s in imputation chains.  -->

<!-- Negative $AC$-values show no threat to the stationarity component of convergence. On the contrary even---negative $AC$-values indicate that $\theta$-values of subsequent iterations diverge from one another, which may increase the variance of $\theta$ and speed up convergence. As convergence diagnostic, the interest is therefore in positive $AC$-values.  -->

<!-- Positive $AC$-values occur when $\theta$-values of subsequent iterations are similar. When , we speak of recurrence. Recurrence within imputation chains indicates trending. Ergo, non-stationarity may be diagnosed when $AC$ values are larger than zero. -->

<!-- Negative $AC$-values indicate divergence within imputation chains. **Subsequent sampled values within each imputation chain are less alike.** Positive $AC$-values indicate recurrence. If $\theta$-values of subsequent iterations are similar, trending may occur.  -->

## Thresholds

<!-- We assume that complete convergence, defined as $\widehat{R} = 1$ and $AC = 0$, will not be observed.  -->
It is unlikely that iterative algorithms such as MICE will achieve the ideal values of $\widehat{R}=1$ and $AC=0$. Because of its convergence to a distribution, the algorithm will show some signs of non-mixing and non-stationarity even in the most converged state. The aim is therefore to reach approximate convergence [@gelm13]. Upon approximate convergence, the imputation chains intermingle such that the only difference between the chains is caused by the randomness induced by the algorithm ($\widehat{R} > 1$), and there is little dependency between subsequent iterations of imputation chains ($AC > 0$). **Approximate convergence is violated when the non-convergence diagnostics exceed a certain threshold. [or] In practice, non-convergence is therefore diagnosed when non-convergence diagnostics exeed a certain threshold.** The conventional thresholds to diagnose non-mixing are $\widehat{R} > 1.2$ [@gelm92] or $\widehat{R} > 1.1$ [@gelm13]. @veht19 proposed a much more stringent threshold of $\widehat{R} > 1.01$. The magnitude of $AC$-values may be evaluated statistically, using a Wald test with $AC=0$ as null hypothesis [@box15]. $AC$-values that are significantly higher than zero indicate non-stationarity. **[think about one-sided test]**

<!-- We assess discrepancy from convergence via practical significance (or some conventional version thereof, such as Rh > 1.1) @gelm13, p. 288 -->

<!-- We can use effective sample size ^neff to give us a sense of the precision obtained from our simulations. As we have discussed in Section 10.5, for many purposes it should suffice to have 100 or even 10 independent simulation draws. (If neff = 10, the simulation standard error is increased by p 1 + 1/10 = 1.05). As a default rule, we suggest running the simulation until ^neff is at least 5m, that is, until there are the equivalent of at least 10 independent draws per sequence (recall that m is twice the number of sequences, as we have split each sequence into two parts so that Rb can assess stationarity as well as mixing). Having an effective sample size of 10 per sequence should typically correspond to stability of all the simulated sequences.  @gelm13, p. 287 -->

## In practice

Before we evaluate the performance of the non-convergence diagnostics $\widehat{R}$ and $AC$ quantitatively through simulation, we apply them to the example of pathological non-convergence by @buur18 that we reproduced above. We want to be able to draw the same conclusions as we did from visually inspecting the two scenarios in Figure \ref{fig:non-conv}. The methods should at least indicate worse performance for the scenario with pathological non-convergence compared to the typically converged algorithm (i.e., higher $\widehat{R}$- and $AC$-values).

In Figure \ref{fig:diagnostics}A, the chain means from Figure \ref{fig:non-conv} are plotted again---now together. Panel B shows two versions of calculating $AC$s applied to the $\theta$s of panel A: the default calculation with `R` function `stats::acf()` [@R], and manual calculation as the correlation between $\theta$ in iteration $t$ and $\theta$ in iteration $t+1$. Panel C shows the traditional computation of $\widehat{R}$, and panel D shows $\widehat{R}$ as computed by implementing @veht19 's recommendations. **[Add labels for the two types of Rhat to the figure!]**

```{r diagnostics, fig.cap="Convergence diagnostics applied on the imputation algorithms of Figure 2. $\\theta$ = chain mean in $ \\dot y_{\\rm imp, \\ell} $."}
# load("./images/diagnostics_plot.Rdata") #bquote(theta ~ "= chain mean in "~ y[imp][",m"])
(theta + plot_layout(guides = "collect", ncol = 2)) + (ac_both + plot_layout(guides = "keep")) + old_rhat + new_rhat +  plot_annotation(tag_levels = "A", tag_suffix = ".") # add 'ncol=1' for plots under each other
```

**[REPHRASE as active about panel A]** From visual inspection, we know that the typically converged algorithm initially portrayed some signs of non-mixing (around $t=2$), but intermingled nicely overall. Additionally, there was very little trending. The algorithm with pathological non-convergence showed severe non-mixing, although this gradually improved (from $t=7$). In this scenario, there was a lot of trending as well initially (up-to $t=6$), after which the chains reached a somewhat more stationary state.

When we look at panel B, we conclude something weird. The $AC$-values calculated with the default function indicate equal performance for the typical convergence and the pathological non-convergence (up-to $t=5$), while there is obvious trending in the $\theta$s of the latter. Moreover, the best convergence (as indicated by the lowest $AC$-value) is observed at $t=2$, but looking at the chain means in panel A, there should be some signs of trending up-to iteration number seven. After consulting the documentation on `stats::acf()`, we conclude that this $AC$ function is not suited for iterative imputation algorithms. The function is optimized for performance when $t\geq50$ [@box15], while the default number of iterations in iterative imputation is often much lower. Therefore, we compute $AC$ manually, see the solid line. These $AC$ values indicate non-stationarity as expected. We therefore calculate $AC$ manually.

**Describe panel C here and why it is not optimal: NB. In panel C: De grotere variantie overall zorgt er kennelijk voor dat $\widehat{R}$ groeit. De $\widehat{R}$ geeft de foutieve boodschap dat convergentie steeds slechter gaat. Verklaring: de overdispersie assumptie klopt niet als er trend is.]** The $\widehat{R}$-values in panel D do meet the requirements specified above. $\widehat{R}$ as computed conform @veht19 does indeed indicate less signs of non-convergence as the number of iterations goes up.  **[explain the dip in Rhat values at t=2. Namely, because we can only use 2 of the 3 tricks by @veht19 if the number of iterations is very low (t<4). That's why the $\widehat{R}$s are more similar to the traditional GR.]**
<!-- However, the superior performance of the typical convergence over the pathological non-convergence is less prominent, even flipped for $t<4$.  -->

**In conclusion, we will use $\widehat{R}$ conform @veht19 and $AC$ computed manually as non-convergence diagnostics in the simulation study. The diagnostics will be applied to four $\theta$s of interest: chain means, chain variances, a quantity of scientific interest, and the first eigenvalue of the variance-covariance matrix. Results in eight methods to identify non-convergence**

# Simulation study

The aim of the simulation study is to determine the **impact of  non-convergence ["early stopping" en "proportion of complete cases"]** in iterative imputation algorithms. Specifically, we are interested in the bias and confidence-validity of estimates obtained from incomplete data with the MICE algortihm.

**[REMOVE?] Several quantities of scientific interest $Q$. Subsequently we will evaluate how well $\widehat{R}$ and $AC$ perform in identifying the effects of non-convergence **[and performance of different thetas]**. And finally, we will formulate an informed advice on the requirements to safely conclude sufficient convergence in practice. That is, we conclude that  convergence is sufficient when each estimate $\bar{Q}$ is an unbiased and confidence-valid estimate of the corresponding estimand $Q$.**

To induce non-convergence, we consider two sets of simulation conditions: we vary the missingness proportion $p_{\rm mis}$ in the incomplete data, and we vary the number of iterations $T$ in the imputation algorithm. The missingness conditions are chosen to reflect the difficulty of the missingness problem. The underlying assumption is that low missingness proportions lead to quick algorithmic convergence, since there is a lot of information in the observed data. Higher missingness proportions should yield slower convergence. Unless, however, the fraction of missing information is so high that the random component in the imputation algorithm outweighs the information in the observed data. Then, convergence to a stable but highly variable state may be reached instantaneously. We expect that the incremental missingness proportions in our study will result in a corresponding increase in signs of non-convergence.

The assumption inherent to the second set of simulation conditions---the number of iterations---is that terminating the imputation algorithm too early causes non-convergence. Generally, the algorithm will not reach convergence if $T=1$, because the imputed values in the first iteration (at $t=1$) depend on the initial values of the algorithm (which are sampled randomly from the set of observed datapoints). As the number of iterations increases, the imputation chains should become independent of the initial values, until the point at which adding an extra iteration does not lead to a more converged state. We assume that we can induce non-convergence at least in conditions where $T$ is smaller than the default number of iterations in `mice`, $T=5$.

**[REMOVE?] $Q$s are the targets that will be estimated in each simulation repetition, after first *amputing* the complete data with varying missingness proportions, then imputing the missingness with a varying number of iterations, then estimating the $Q$s on the $m$ completed datasets, and finally pooling the $m$ estimates $\hat{Q}_\ell$ to obtain a single $\bar{Q}$ for each $Q$. We try to predict the effect of non-convergence (i.e., the difference between $Q$ and $\bar{Q}$) with the convergence diagnostics $\widehat{R}$ and $AC$, by applying these methods to several scalar summaries of the state space of the algorithm, $\theta$s.**

## Hypotheses

1. We expect that simulation conditions with a high missingness proportion $p_{\rm mis}$ and a low number of iterations $T$ will more often result in biased, invalid estimates of the quantities of scientific interest, $Q$s.

2. We hypothesize that $\widehat{R}$ and $AC$ will correctly identify signs of non-convergence in those simulation conditions where the $\bar{Q}$s are *not* unbiased and confidence-valid estimates of the $Q$s.

3. We hypothesize that the recommended thresholds to diagnose non-convergence with $\widehat{R}$ ($\widehat{R} > 1.2$, $\widehat{R} > 1.1$, and $\widehat{R} > 1.01$) may be too stringent for iterative imputation applications. In an empirical study, where $\widehat{R}$ was used to inform the required imputation chain length, it took as many as 50 iterations to overcome the conventional non-convergence threshold $\widehat{R}>1.2$. Yet, scientific estimates were insensitive to continued iteration from $t=5$ onward [@lace07]. We therefore suspect that $\widehat{R}$ may over-estimate signs of non-convergence in iterative imputation algorithms, compared to the validity of estimates. In contrast to this, it may occur that signs of non-convergence are under-estimated by $\widehat{R}$, in exceptional cases where the initial values of the algorithm are not appropriately over-dispersed [@broo98, p. 437]. In  `mice`, initial values are chosen randomly from the observed data, hence we cannot be certain of over-dispersion in the initial values. In practice, we do not expect this to cause problems for identifying non-convergence with $\widehat{R}$. **[ADD limitation that other MI packages may have different way to obtain initial values!]**

4. We expect that high $AC$ values are implausible in iterative imputation algorithms with typical convergence. After only a few iterations, the randomness induced by the algorithm should effectively mitigate the risk of dependency within chains.

5. We further hypothesize that multivariate $\theta$s are better at detecting non-convergence than univariate $\theta$s.

<!-- **(bruggetje hier)** For multiple imputation algorithms, it holds that convergence is reached when there is no dependency between subsequent iterations of imputation chains ($AC = 0$), and chains intermingle such that the only difference between the chains is caused by the randomness induced by the algorithm ($\widehat{R} = 1$). **We expect that a completely converged state will not be reached.**  -->

<!-- This study evaluates whether $\widehat{R}$ and $AC$ could diagnose convergence of multiple imputation algorithms. We assess the performance of the two convergence diagnostics against the recommended evaluation criteria for MI methods [i.e., average bias, average confidence interval width, and empirical coverage rate across simulations; @buur18, $\S$ 2.5.2]. **That is, there is no baseline measure available to evaluate performance against.** -->

<!-- We could also look at distributional characteristics, and plausibility of imputed values, see @vinknd. For now, this is outside of the scope of this study. -->

<!-- Based on an empirical finding [@lace07], we hypothesize that $\widehat{R}$ will over-estimate non-convergence of MI algorithms **explain why?**. Lacerda concluded that rhat dropped below 1.2 after at least 50 iterations. Therefore we hypothesize that Vehtari's proposed threshold of $\widehat{R} < 1.01$ will be too stringent for diagnosing convergence in iterative imputation. This over-estimation may, however, be diminished because $\widehat{R}$ can falsely diagnose convergence if initial values of the algorithm are not appropriately over-dispersed [@broo98, p. 437]. In e.g. `mice`, initial values are chosen randomly from the observed data. Therefore, we cannot be certain that the initial values are over-dispersed. We expect this to have little effect on the hypothesized performance of $\widehat{R}$. **(Add actual hypothesis: we thought that mice would converge sooner than $\widehat{R}$ would indicate that it did). On the one hand, $\widehat{R}$ is known to under-estimate signs of non-convergence when the initial values of the algorithm are not appropriately over-dispersed [@broo98, p. 437], which may be the case for iterative imputation algorithms. In e.g. `mice`, initial values are chosen randomly from the observed data. Therefore, we cannot be certain that the initial values are sufficiently over-dispersed. 
**  -->

## Set-up

We investigate non-convergence of the MICE algorithm through model-based simulation in `R` [version 3.6.3; @R]. The number of simulation repetitions is 1000, and the simulation set-up is summarized in the pseudo-code below. The complete `R` script of the simulation study is available from [github.com/gerkovink/shinyMice](https://github.com/gerkovink/shinyMice/tree/master/3.Thesis/).

```
# pseudo-code of simulation 
1. simulate data 
for (number of simulation runs from 1 to 1000)
 for (missingness proportions 5%, 25%, 50%, 75% and 95%)
  2. create missingness
  for (number of iterations from 1 to 100)
   3. impute missingness
   4. perform analysis of scientific interest
   5. compute non-convergence diagnostics 
   6. pool results across imputations
   7. compute performance measures
 8. combine outcomes of all missingness proportions
9. aggregate outcomes of all simulation runs

```
### Data-generating mechanism.

In this study, sampling variance is not of interest. Therefore, a single complete dataset may serve as comparative truth in all simulation repetitions [@vink14]. The data-generating mechanism is a multivariate normal distribution, representing person-data on three predictor variables in a multiple linear regression problem. Let the predictor space be defined as 
$$
\begin{aligned}
\begin{pmatrix}X_1\\
X_2\\
X_3
\end{pmatrix} \sim \mathcal{N}
\begin{bmatrix}
\begin{pmatrix}
12\\
3\\
0.5
\end{pmatrix}\!\!,
\begin{pmatrix}
4 & 4 & 1.8 \\
4 & 16 & 4.8\\
1.8 & 4.8 & 9
\end{pmatrix}
\end{bmatrix}\!\!\text{.}\\[2\jot]
\end{aligned}
$$
A finite population of $N=1000$ is simulated using the `mvtnorm` package [@mvtnorm]. Subsequently, a fourth variable is constructed as outcome variable $Y$. For each unit $i = 1, 2,..., N$, let 
$$
Y_i = 1 + 2X_{1i} +.5X_{2i} - X_{3i} + \epsilon_i ,
$$
where $\epsilon \sim \mathcal{N}(0, 100)$. This results in a complete set $y$, with observed values for all units $i$ on all variables $j$ (where $j =  Y, X_1, X_2, X_3$). From the complete set we obtain the true values of the scientific estimands $Q$.

### Scientific estimands.

We consider four types of $Q$s that are often of interest in empirical research. Namely, two descriptive statistics: the mean $\mu_j$ and standard deviation $\sigma_j$ of each variable $j$. We also consider the regression coefficients of the predictors: $\beta_1$, $\beta_2$, and $\beta_3$ in regression equation
$$Y' = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \beta_3 X_3,$$
where $Y'$ is the expected value of the outcome. And finally, the proportion of variance explained by the set of predictors: coefficient of determination $r^2$ **(note that lower case $r$ is used to avoid confusion with non-convergence diagnostic $\widehat{R}$)**. These $Q$s are estimated by $\bar{Q}$ in each simulation condition.

<!-- Each of these $Q$s is estimated with a corresponding $\bar{Q}$, after the complete dataset is first *amputed* (with a varying missingness proportion $p_{\rm mis}$) and subsequently imputed (with a varying number of iterations $T$).  -->

### Simulation conditions.

In each simulation repetition, we *ampute* the complete dataset five times to obtain five different incomplete datasets, $y$s. The ratio between $y_{\rm obs}$ and $y_{\rm mis}$ in each $y$ depends on the missingness proportion: $p_{\rm mis} =.05,.25,.5,.75,.95$. We ampute the complete data with the `mice` package [function `mice::ampute()`; @mice]. We consider all possible univariate and multivariate patterns of missingness, conform a 'missing completely at random' missingness mechanism [@rubin87]. I.e., the probability to be missing is the same for all $N \times k$ cells in $y$, conditional on the missingness proportion $p_{\rm mis}$. 
<!-- This note only considers a 'missing completely at random' (MCAR) mechanism, where the probability of being missing is equal for all $n \times k$ cells in $y$ [@rubin87]. The probability to be missing is the same for all $n \times k$ cells in $y$ [a 'missing completely at random' missingness mechanism; @rubin87]. -->

Missing datapoints in each incomplete dataset $y$ are imputed with the `mice()` function [@mice]. All imputation procedures are performed with Bayesian linear regression imputation, and five imputation chains ($m=5$). Because $m=5$, each run of the `mice()` function results in five sets of imputations: $\dot y_{\rm imp, \ell}$, where $m = 1, 2, ..., 5$. The number of iterations in each run of the algorithm varies between simulation conditions ($T=1,2,...,100$). 
<!-- Each repetition of the simulation thus starts from 1 complete dataset, then splits into 5 missingness conditions ($p_{\rm mis} =.05,.25,.5,.75,.95$), each of the amputed datasets is imputed 5 times ($m = 1, 2,..., 5$) for each iteration condition ($t = 1, 2,..., 100$). From these $5\times5\times100$ imputations, we extract the $\theta$s to apply $\widehat{R}$ and $AC$ on (**i.e., chain means, chain variances, $\beta$s per imputation, and the first eigenvalue of the variance-covariance matrix per imputation**).  -->

From the 5 imputations $\dot y_{\rm imp, \ell}$ we obtain the chain means and chain variances for each variable $j$. These will serve as scalar summary $\theta$ to apply non-convergence diagnostics $\widehat{R}$ and $AC$ on. We then use `mice::complete()` to combine the imputed data $\dot y_{\rm imp, \ell}$ with the observed data $y_{\rm obs}$, resulting in five completed datasets {$y_{\rm obs}, y_{imp, m}$}.

On the 5 completed datasets {$y_{\rm obs}, y_{imp, m}$}, we:

- Estimate $\hat{Q}_\ell$s for $Q= \mu$ and $Q=\sigma$ with `base::mean()`, and `stats::sd()` [@R]. We pool the $\hat{Q}_\ell$s with `base::mean()` to get $\bar{Q}$s.

- Compute $\lambda_{1, \ell}$, to use as $\theta$ in the two convergence diagnostics. By definition, the first eigenvalue of the variance-covariance matrix, $\lambda_{1, \ell}$, is equal to the variance of the principal component solution of each {$y_{\rm obs}, y_{imp, m}$}, which we calculate with `stats::princomp()` [@R].

- Perform multiple linear regression with `stats::lm()` [@R] to get the $\hat{Q}_\ell$s for $Q=\beta$ and $Q=r^2$. We pool the $\hat{Q}_\ell$s conform @vink14 to get $\bar{Q}$s for the $\beta$s and $r^2$ **[check  finite pooling of  $r^2$, instead of `mice::pool.r.squared()`]**. Additionally, we use the $\hat{Q}_\ell$s of $Q=\beta$ as $\theta$ to apply the convergence diagnostics to.

<!-- After each run of the `mice` algorithm, the function `mice::complete()` is used to combine the imputed data $\dot y_{\rm imp, \ell}$ with the observed data $y_{\rm obs}$, resulting in five completed datasets {$y_{\rm obs}, y_{imp, m}$}.  -->

<!-- The estimator for each $Q$ is $\bar{Q}$---the pooled aggregate of the $\hat{Q}_\ell$s across imputations. To calculate the $\bar{Q}$s, we combine the observed data $y_{\rm obs}$ and the imputed data for each imputation $\dot y_{\rm imp, \ell}$ with the function `mice::complete()`. Descriptive statistics are computed as the average across imputations for each variable ($\mu_j$ and $\sigma_j$, where $j = Y, X_1, X_2, X_3$). Estimated regression coefficients are obtained with the function `stats::lm()` for each imputation, and then pooled conform @vink14. The coefficient of determination is estimated for each imputation, and pooled using `mice::pool.r.squared()`.  -->

We apply non-convergence diagnostics $\widehat{R}$ and $AC$ to each scalar summary $\theta$. We calculate $\widehat{R}$ by implementing Vehtari et al.'s [-@veht19] recommendations, and $AC$ as the correlation between the $t^{th}$ and the $(t+1)^{th}$ iteration.

### Performance measures.

We assess the performance of the iterative imputation algorithm in each simulation condition by comparing $\bar{Q}$s with $Q$s. For each $Q$, we calculate bias as $\bar{Q} - Q$. For the regression coefficients $Q=\beta_{1,2,3}$, we also compute the empirical coverage rate (CR). CR is defined as the percentage of simulation repetitions in which the 95\% confidence interval (CI) around $\bar{Q}$ covers the true estimand $Q$. Let
$$\text{CI} = \bar{Q} \pm t_{(m-1)} \times \rm SE_{\bar{Q}},$$
where $t_{(M-1)}$ is the quantile of a $t$-distribution with $m-1$ degrees of freedom, and $\rm SE_{\bar{Q}}$ is the square root of the pooled variance estimate. If we obtain nominal coverage (CR = 95%), we can conclude that $\bar{Q}$ is a confidence-valid estimate of $Q$. Finally, we inspect CI width (CIW): the difference between the lower and upper bound of the 95\% confidence interval around $\bar{Q}$. CIW is of interest because it is a measure of efficiency. Under nominal coverage, short CIs are preferred, since wider CIs indicate lower statistical power.

We evaluate the methods to identify non-convergence against these performance measures. $\widehat{R}$ and $AC$ should identify non-convergence in simulation conditions, where $\bar{Q}$ is *not* an unbiased, confidence-valid estimate of $Q$.

<!-- We measure the performance of $\widehat{R}$ and $AC$ by the four scalar summaries $\theta$: chain means (i.e., the mean in each $y_{imp, m, j}$) versus $Q=\mu$, chain variances (i.e., the variance in each $y_{imp, m, j}$) versus $Q=\sigma$, the estimated regression coefficients $\beta$ in each {$y_{\rm obs}, y_{imp, m}$} (i.e., a user-defined statistic of scientific interest $\hat{Q}_\ell$) against $Q=\beta$, and the first eigenvalue of the variance-covariance matrix in each {$y_{\rm obs}, y_{imp, m}$}, $\lambda_{1, \ell}$ versus $Q=r^2$.** -->

<!-- **[Remove this?] Primarily, we want to know what the effect is of inducing non-convergence on the estimated $Q$s. More specifically, we are interested in the bias of $\bar{Q}$ for all $Q$s and the confidence-validity of $\bar{Q}$ for $Q=\beta$. Therefore, we evaluate these performance measures against the simulation conditions (the number of iterations $T$ and the missingness proportion $p_{\rm mis}$). We look at averages across simulation repetitions. Secondly, we want to know if conditions in which the $Q$s are affected by non-convergence are also identified by the non-convergence diagnostics $\widehat{R}$ and $AC$. We do this, again, by looking at the performance measures for the $Q$s, but now we evaluate these against what $\widehat{R}$ and $AC$ indicate about signs of non-convergence in the $\theta$s.** -->

# Results

**[Add more info about figure legends and axes]**

Our results show the performance of the MICE algorithm under different conditions of missingness and iteration length. For reasons of brevity, we only discuss results for the worst-performing estimates in terms of bias. For univariate scientific estimands ($Q=\mu$ and $Q=\sigma$) we observe the largest bias in the outcome variable $Y$. And for the regression coefficients ($Q=\beta$) the bias is most pronounced in $\beta_1$ (the effect of $X_1$ on $Y$). Since there is just one estimate for $Q=r^2$ to consider, we **end up with** $Q=\mu_Y, \sigma_Y,r^2, \beta_1$. In the figures, we present all missingness conditions, but only **iteration length conditions** $1 \leq T\leq50$, since the results are more or less stable for conditions where $T \geq 30$. Full results are available from [github.com/gerkovink/shinyMice](https://github.com/gerkovink/shinyMice/tree/master/3.Thesis/1.SimulationStudy).

## Quantities of scientific interest

```{r Qs, fig.height = 6.8, fig.cap="Impact of non-convergence on statistical inferences. Depicted are the bias, coverage rate (CR) and confidence interval width (CIW) of the estimates $\\bar{Q}$. We present the subset of $Q$s with the worst performance in terms of bias: the mean and standard deviation of $Y$, and the regression coefficient of $X_1$."}
mean_bias + sd_bias + Rsq_bias + est_bias + est_cov + est_ciw + plot_layout(ncol = 2, guides = "collect")
```

Figure \ref{fig:Qs} displays the influence of **inducing non-convergence ["early stopping" en "proportion of complete cases"]** on the estimated quantities of scientific interest, $Q$s. **[Add panel labels, describe panels]** Simulation conditions that are impacted by non-convergence portray more extreme bias in the estimated $Q$s, and non-nominal coverages.

### $Q=\mu$.
The estimated univariate means seem unaffected by the number of iterations in the algorithm. This implies that approximately unbiased estimates may be obtained with as little as one iteration ($T \geq 1$). The magnitude of the bias in $\bar{Q}$ depends solely on the missingness proportion $p_{\rm mis}$, with missingness proportions of 75% and 95% leading to more extreme biases than other missingness proportions. Completely unbiased estimates are only obtained in conditions where $p_{\rm mis}\leq.50$. **This is curious because the missingness mechanism MCAR should yield unbiased univariate estimates without employing multiple imputation.** Note however, that the magnitude of the bias in these conditions is small: $\mu_Y$ is maximally under-estimated by 0.02 units, while the true value of $\mu_Y$ is 25.81 ($\sigma_Y = 11.32$). We, therefore, conclude that non-convergence does not substantially impact the validity of the inferences for $Q=\mu$.

### $Q=\sigma$.
The estimated standard deviations not affected substantially by **early stopping** either. While stable estimates are obtained in conditions with two iterations, the bias in the first iteration is actually *less* severe than in other **iteration conditions**. Therefore, no iteration at all is needed to obtain approximately unbiased estimates for $Q=\sigma$. Completely unbiased estimates are only obtained in conditions where $p_{\rm mis}=.05$. Conditions where $p_{\rm mis}\geq.25$ are impacted by non-convergence in order of increasing missingness proportion. Similar to the bias for $Q=\mu$ the magnitude of the bias in $Q=\sigma$ is negligible. We therefore conclude that neither of the univariate $Q$s is affected substantially by non-convergence.
<!-- in the first iteration. how impact of non-convergence similar to the means. The magnitude of the bias $\bar{Q}$ follows the incremental order of the missingness proportions. The number of iterations in the algorithm has little effect on the bias. Only the estimates in simulation conditions with just one iteration ($T=1$) differ from other conditions that have the same missingness proportion but a higher number of iterations. Interestingly, the bias in conditions where $T=1$ is *less* severe than after continued iteration. For the least possible bias, we may use any number of iterations $T\geq1$, but only $p_{\rm mis} = 0.05$ results in completely unbiased $\bar{Q}$s. -->

### $Q=r^2$.
The estimated coefficient of determination is clearly affected by early stopping. As expected, we observe worse performance in conditions with a lower number of iterations. The magnitude of the bias, however, depends on the missingness proportion---following the incremental order in $p_{\rm mis}$. **And so does** the number of iterations necessary to reach stable, approximately unbiased estimates. In conditions where $p_{\rm mis}=.05$, estimates are unbiased after one iteration. We observe approximately unbiased estimates when $T\geq 2,4, \text{and }5$, in conditions where $p_{\rm mis} = .25, .50, \text{and }.75$, respectively. The highest number of iterations necessary to reach approximate unbiasedness is seven. Completely unbiased estimates are only obtained in conditions where $p_{\rm mis}\leq.25$. This implies that even for biased estimates, $T\geq7$ would suffice to reach a stable solution.

<!-- portion is at most 25%. Approximately unbiased estimates are observed for any missingness proportion. For $p_{\rm mis}\leq.25$, the minimum number of iterations is two ($T\geq2$),  for $p_{\rm mis}=.50$ it is $T\geq4$, and $p_{\rm mis}=.75$ requires $T\geq5$. The highest number of iterations needed to obtain valid inferences is seven ($p_{\rm mis}=.95$; $T\geq7$).  **[interpret sign of bias? ADD what this implies: dat - zelfs bij een gebiasede methode - je maar 7 iteraties nodig hebt om het eindresultaat te bereiken.]** -->

<!-- while the best performance occurs with less missingness and more iterations. Only conditions with the lowest missingness proportion ($p_{\rm mis}=0.05$) results in unbiased estimates for every number of iterations. In other conditions there is at least some bias irrespective of $T$. The inferences do not improve in any condition where $T \geq 6$. This suggests that the algorithm is as converged as necessary to obtain valid inferences. -->

### $Q=\beta$.
For the estimated regression coefficients, we first consider bias, before discussing CR and CIW. As Figure \ref{fig:Qs} shows, the bias in $\bar{Q}$ is affected by both the missingness proportion and the number of iterations. Approximately unbiased estimates are observed after one to seven iterations, depending on $p_{\rm mis}$.Completely unbiased estimates are only obtained in conditions where $T\geq3$ and $p_{\rm mis}\leq .50$.

Despite persistant bias in conditions where $p_{\rm mis}\geq.75$, coverage rates indicate that all **missingness conditions** have nominal coverages after just three iterations. Confidence-valid estimates may be obtained in any condition where $T\geq3$. 
<!-- , whereas we need $T\geq7$ to reach approximate unbiasedness. -->

The nominal coverages are explained by the CIWs. Conform the theoretical foundation of MI, CIs are wider in conditions with higher missingness proportions (i.e., there is less information in the data, and thus more uncertainty due to missingness). The wider CIs around $\bar{Q}$ in conditions with higher missingness proportions **results in the inclusion of** the true value of $Q$ despite of bias.

We conclude that approximately unbiased estimates may be obtained in conditions where $T\geq7$, whereas confidence-valid estimates require at most three iterations.

<!-- confidence interval width (CIW), and coverage rate (CR) as performance measures. **[Split panels]** Unbiased estimates are obtained in conditions where $p_{\rm mis}\leq.50$ and $T\geq3$. Approximately confidence-valid estimates are obtained in any condition where $T\geq3$---despite of persistant bias in conditions where $p_{\rm mis}\geq.75$. These approximately nominal coverages are explained by the CIWs: as expected, conditions with higher missingness proportions result in more uncertainty in the estimates (i.e., wider CIs), which diminish the effect of the bias. confidence-valid inferences may be obtained when $T\geq3$, while we need at least seven iterations to reach valid inferences in terms of bias. Continued iteration after $T\geq7$ seems futile. **[add max T for other betas?]** -->
<!-- One exception is the CIW of the most extreme condition ($p_{\rm mis}=.95$; $T=1$). -->

### Summary

These results demonstrate that the estimates of univariate scientific estimands $Q$ are not impacted by early stopping of the MICE algorithm, irrespective of the missingness proportion in $y$. Unbiased estimates may be obtained after just one iteration. Multivariate estimates, by contrast, are affected by both the number of iterations and the missingness proportion. Completely unbiased estimates are only observed under low to moderate missingness ($p_{\rm mis}\leq.50$), after at most three iterations ($T\leq1,2,3$, depending on $p_{\rm mis}$). Approximately unbiased estimates are observed for any missingness condition, after at most seven iterations. This implies that the algorithm produces stable, unimproving estimates for any condition where $T\geq7$.

<!-- Overall, we see that conditions with a higher proportion of missingness are affected more severely by non-convergence. Multivariate $Q$s are also substantially influenced by the number of iterations. This means that the algorithm has indeed not converged at $t=1$, and converges gradually as $t$ moves to $T$. The point at which an additional iteration does not lead to an improvement in the estimates depends on quantity of scientific interest $Q$ and the missingness proportion $p_{\rm mis}$. The highest number of iterations that is necessary to reach this point across all simulation conditions is $T=7$. This implies that, under the current specifications, no more than seven iterations are necessary to obtain sufficient performance in terms of bias and confidence-validity.  -->

<!-- Conditions with a lower number of iterations yield more bias in $\bar{Q}$. This is especially pronounced in conditions with high missingness proportions: with 95% missingness we need at least seven iterations before an additional iteration does not lead to improved estimates. Conditions with a lower missingness proportion (e.g., $p_{\rm mis} = {0.5}$), produce *optimal* estimates with as many as two iterations. Approximately unbiased estimates are obtained for all conditions where $p_{\rm mis}\leq0.75$ and $T\leq6$. **[Add CIW and CR]** -->

<!-- The sign of the bias is logical: when there is less info in the data, the relations between the predictors and the outcome is under-estimated. -->

## Non-convergence diagnostics

Figure \ref{fig:rhats-acs} displays results for the two non-convergence diagnostics applied to the four $\theta$s. **[describe panels]** We evaluate the performance of these methods by establishing whether they correctly identify conditions in which $\bar{Q}$ is *not* an unbiased, confidence-valid estimate of $Q$. Biased estimates are observed in conditions where $T\leq7$. Non-nominal coverage rates are observed only when $T\leq3$. Additionally, there is persistant bias in conditions where $p_{\rm mis}\geq.75$, irrespective of the iteration length. **[appropriate methods thus identify these conditions]**

<!-- We evaluate the non-convergence diagnostics $\widehat{R}$ and $AC$ against the performance measures described above (i.e., bias and confidence-validity of the quantities of scientific interest $Q$). The diagnostics should identify those simulation conditions that are negatively impacted by non-convergence. We expect the most signs of non-convergence in conditions where valid estimates are not obtained (high missingness proportions, $T<7$), and at least some signs of non-convergence in conditions where estimates are not approximately unbiased ($p_{\rm mis}=.95$, any number of iterations).  -->

<!-- As we have observed, the affected conditions are $p_{\rm mis}=.95$ for all all $Q$s are impacted by the missingness proportions, but only the multivariate quantities of scientific interest are affected by the number of iterations.  -->

Figure \ref{fig:rhats-acs} displays the non-convergence diagnostics $\widehat{R}$ and $AC$ for each of the four $\theta$s under consideration. **[Add panel labels and explain how to read the plots]**

```{r rhats-acs, fig.height = 6.8, fig.cap="Non-convergence diagnostics. [Split up into 2 or 4 figures??]"}
mean_Rh + mean_AC + var_Rh + var_AC + est_Rh + est_AC + PCA_Rh + PCA_AC + plot_layout(guides = "collect", ncol = 2)
```

<!-- **Message: after one iteration ($T$=2), the algorithm is as well converged as the missingness proportion allows. Sooner even with lower missingness proportions---$T$=1 might be sufficient for $p_{\rm mis}$ = 5, 25, and 50.** Number of iterations has little influence on bias in $\hat{Q}_\ell$, only in first iteration there is some bias in missingness conditions of 25 and up. There is a clear different between the missingness conditions. The higher the missingness prop, the larger the magnitude of the bias. For this $Q$, t=2 would be sufficient. And only for $p_{\rm mis}$ = 5 the $\bar{Q}$s are unbiased. All other $p_{\rm mis}$ conditions have bias irrespective of the number of iterations. -->

### $\theta=$ chain mean.
As expected, $\widehat{R}$ lowers with the number of iterations. After an initial 'dip' in conditions where $T=3$, we observe a gradual decrease in $\widehat{R}$-values with every additional iteration. This implies improving convergence, until the decrease tapers off after 30 to 40 iterations. Using the threshold $\widehat{R}>1.2$, we diagnose non-convergence in conditions where $T\leq7$, whereas the threshold 1.1 is exceeded when $T\leq13$. The very strict threshold 1.01 is surpassed in all conditions $T\leq100$**, and therefore not informative in the current study**. With this method and $\theta$, we fail to identify simuation conditions with persistant bias due to high missingness proportions.

Similarly, $AC$ decreases with higher $T$s. $AC$ indicates improving stationarity in conditions where $T\leq6$. The $AC$-values do not exceed the threshold **defined by statistically significant $AC$s**. Based on this threshold, we fail to not diagnose non-convergence in any condition.

<!-- We evaluate the performance of the non-convergence diagnostics applied to chain means against the bias in $Q=\mu$. Based on the bias in $Q$, we concluded that one iteration is sufficient to obtain valid inferences, irrespective of the missingness proportion. The non-convergence diagnostics, however, identify some signs of non-convergence in conditions with as many as thirty iterations. This does not correspond to the impact of non-convergence in the univariate $Q$s. **Compared to the multivariate $Q$s [add before that we evaluate the thetas against all Qs?], $\widehat{R}$ and $AC$ applied to the chain means fail to identify the non-convergence in conditions where $p_{\rm mis}=.95$.**  -->

```{r mean, fig.height = 6.8, fig.cap="Convergence diagnostics chain mean."}
# mean_bias + mean_Rh + mean_AC + plot_layout(guides = "collect", ncol = 1) + plot_annotation(tag_levels = "A", tag_suffix = ".", title = bquote("Q = " ~ mu[Y] ~ "= 25.81; " ~ theta ~ "= chain mean of Y in" ~ y[imp][",m"])) #& theme(legend.position = "bottom")
```

### $\theta=$ chain variance.
As Figure \ref{fig:rhats-acs} shows, the results of the two diagnostics applied to the chain variances are highly similar to those on the chain means. This holds for both $\widehat{R}$ and $AC$. The only apparent difference is in the $\widehat{R}$ threshold 1.1. Instead of diagnosing non-convergence in conditions where $T\leq13$, we now diagnose this when $T\leq11$. All other observations are equivalent.

<!-- We thus conclude that there is no added value of evaluating non-convergence with chain variances as $\theta$ in the diagnostics. -->

```{r sd, fig.height = 6.8, fig.cap="Convergence diagnostics chain variance."}
# sd_bias + var_Rh + var_AC + plot_layout(guides = "collect", ncol = 1) + plot_annotation(tag_levels = "A", tag_suffix = ".", title = bquote("Q = " ~ sigma[Y] ~ "= 11.32; " ~ theta ~ "= chain variance of Y in" ~ y[imp][",m"]))
```

### $\theta=\hat{Q}_\ell$.
The $\widehat{R}$-values show a similar trend across iterations as we observed for chain means and chain variances. Only conditions with very high missingness proportions ($p_{\rm mis}\geq.75$) diverge from the earlier observations. The $\widehat{R}$-values in these conditions do not taper off after 30 to 40 iterations, but rather between 40 to 50 iterations. The highest number of iterations at which non-convergence is still diagnosed according to the thresholds are: 11 for $\widehat{R}>1.2$, 23 for $\widehat{R}>1.1$, and 100 for $\widehat{R}>1.01$.

The $AC$-values also follow earlier observations, with the exception of the same very high missingness proportions that **diverged** for $\widehat{R}$. To reach maximum stationarity, these conditions require not five but six or seven iterations. Moreover, the $AC$-values in these conditions exceed the threshold to diagnose non-convergence. Interestingly, this only occurs after ten or even thrity iterations, while we would expect more signs of trending early in the iterations. **Since the $AC$-values are not improving or worsening after seven iterations, we conclude that non-convergence is incorrectly diagnosed based on the threshold.**

<!-- The non-convergence diagnostics are applied to the estimated regression coefficients in each imputation, and evaluated against the bias in the corresponding pooled estimate.  -->

<!-- $\widehat{R}$ indicates signs of non-convergence across all missingness proportions and each number of iterations, in the expected order: simulation conditions with higher missingness proportions and less iterations are identified as the worst converged. The thresholds to diagnose non-convergence are overcome in different conditions, depending on the number of iterations and the missingness proportion. The most lenient threshold ($\widehat{R}>1.2$) is overcome at $T\geq9$ for $p_{\rm mis} = 0.05$, and $T\geq12$ for $p_{\rm mis} = 0.95$. The conventional $\widehat{R}>1.1$ is surpassed at $T\geq12$ and $T\geq24$, respectively. -->

<!-- For the autocorrelations applied to the $\hat{Q}_\ell$s, we see an even stronger effect of the missingness proportions. Especially the conditions where $p_{\rm mis} = 0.75$ and $p_{\rm mis} = 0.95$ are identified as non-stationary. The threshold used to evaluate whether $AC=0$, however, is only reached in conditions with at least ten iterations. This suggests that the default number of iterations may be too low to detect non-stationarity with $AC$. However, the critical value for diagnosing non-convergence with $AC$ decreases as a function of the number of iterations. The conclusion to identify non-stationarity in conditions where $T\geq10$ and $p_{\rm mis} = 0.95$ might therefore be spurious.  -->

<!-- **[Also, the $AC$ across missingness proportions shows the opposite of what you would expect: shouldn't there be higher autocorrelations in conditions with comparatively more observed data in the completed datasets? I don't understand this.]** -->

<!-- Compared to the bias in $\bar{Q}$, the diagnostics applied to the $\hat{Q}_\ell$s over-estimate the severity of the non-convergence.  -->

### $\theta=\lambda_{1, \ell}$.
Once again, we observe $\widehat{R}$-values with a similar trend across iterations. For this $\theta$, the results differ only in the most extreme missingness condition. We diagnose non-convergence in conditions where $T\leq9$ according to $\widehat{R}>1.2$, and $T\leq19$ for $\widehat{R}>1.1$. The 'dip' in $\widehat{R}$-values at $T=3$ is even more pronounced than with other $\theta$s. Some $\widehat{R}$-values even overcome the threshold of 1.2. If we would terminate the algorithm after three iterations, we would incorrectly conclude sufficient convergence according to the $\widehat{R}>1.2$ threshold.

$AC$-values are systematically higher for this $\theta$ than in the other results. The $AC$-values also taper off at a later point in the iterations, indicating improving stationarity up-to ten, or even thirty iterations, depending on the missingness proportion. Using the threashold, we diagnose non-convergence at some point for any missingness proportion.

<!-- And especially in conditions with higher missingness proportions, we observe signs of increasing stationarity the decrease in the $AC$-values across iterations is also  -->

<!-- Finally, we evaluate the non-convergence diagnostics applied to the first eigenvalue of the variance-covariance matrices of the completed data. Similar to the results the the regression coefficients as $\theta$s, we see effects of both the missingness proportion and number of iterations in each simulation condition. The most signs of non-convergence are identified in conditions where $p_{\rm mis}$ is large and $T$ is small. -->

<!-- Notwithstanding the similarities with $\theta=\hat{Q}_\ell$, there is a big difference in $\widehat{R}$-values for conditions where $T<4$. In these conditions, $\widehat{R}$ behaves oppositely of our expectation: conditions with the lowest missingness proportions yield the highest $\widehat{R}$-values. In conditions with a higher number of iterations, this artifact disappears. This suggests that the combination of $\widehat{R}$ and $\lambda_{1, \ell}$ is only appropriate for imputation algorithms where the number of iterations at least four. Interpreting the $\widehat{R}$ values of conditions where $T<3$ may lead to the incorrect conclusion that only conditions with *low* missingness proportions show signs of non-convergence. For applications as a model-independent multivariate $\theta$, we should therefore only interpret the $\widehat{R}$-values when $T\geq4$. Using this cut-off we conveniently ignore the initial 'dip' at $T=3$ as well. The traditional threshold $\widehat{R} > 1.2$ is overcome in conditions where $p_{\rm mis} = .95$ and $T\geq10$, and for all other missingness proportions in conditions where $T\geq8$. With the $\widehat{R} > 1.1$ threshold, the number of iterations are $T\geq20$ and $T\geq16$, respectively. -->

<!-- The $AC$-values largely depend on the missingness proportion in each condition. Conditions where $p_{\rm mis}=.95$ show statistically significant autocorrelations for any number of iterations. The other extreme, conditions where $p_{\rm mis}=.05$, does not yield $AC$ values above the threshold. The $AC$ values with this missingness proportion are equivalent for all conditions where $T\geq6$, which suggests that the algorithm is as stationary as it can be. A similar point is only observed after twenty iterations in conditions with higher missingness proportions.  -->

<!-- In comparison to the $Q$s, the diagnostics again over-estimate the signs of non-convergence. Conditions with low missingness proportions (e.g., 5% or 25% of the complete data) yield unbiased, confidence-valid estimates with as little as two iterations. The convergence diagnostics, however, indicate improved convergence up-to eight iterations according to $AC$ and thirty to forty iterations according to $\widehat{R}$. The model-independent $\theta$ $\lambda_{1, \ell}$ seems as well-equipped to diagnose signs of non-convergence as $\theta=\hat{Q}_\ell$. The univariate $\theta$s do not differentiate between bias induced by the missingness proportions.   -->

### Summary

Overall, the methods to diagnose non-convergence perform as expected: they indicate more signs of non-convergence in conditions with worse performance in terms of bias and confidence-validity. Under low to moderate missingness, it does not seems to matter on which $\theta$ the diagnostics are applied. If we look at complete unbiasedness, however, we notice that univariate $\theta$s fail to diagnose the persistant bias in conditions where $p_{\rm mis}\geq.75$. The number of iterations necessary to obtain approximately unbiased, confidence-valid estimates corresponds to the $\widehat{R}$ threshold 1.2. The thresholds 1.1 and 1.01 seem to strict compared to the validity of the inferences. The $AC$ threshold defined by the statistical significance of the $AC$-values does not appear to be appropriate in the current set-up **[or iterative imputation in general??]**. Non-stationarity may be better diagnosed by evaluating the $AC$-values across iterations. If the values do not substantially decrease, stationarity may be concluded **[similar to the elbow heuristic in PCA]**.

<!-- In general, the diagnostics identify more signs of non-convergence in simulation conditions with a lower number of iterations. The missingness proportions, however, only seem to impact multivariate convergence. When $\widehat{R}$ and $AC$ are applied to  univariate $\theta$s, they identify the same amount of non-convergence for any missingness proportion. When the diagnostics are applied to multivariate $\theta$s, they generally identify more signs of non-convergence for higher missingness proportions.  -->

<!-- As expected, none of the simulation conditions that we consider resulted in complete convergence of the algorithm (defined as $\widehat{R}=1$ and $AC=0$). However, the thresholds used in practice to diagnose non-convergence seem too conservative for imputation purposes. $\widehat{R}$ and $AC$ identify signs of non-convergence in conditions where the $\bar{Q}$s are unbiased and confidence-valid estimates of the $Q$s. The most recent recommended threshold for diagnosing non-convergence with $\widehat{R}$ is never overcome in this study (all $\widehat{R}$>1.01). Apparently, this threshold is too stringent for iterative imputation algorithms, and will not be discussed any further. -->

<!-- This is odd, since the $\theta$ and $Q$ are directly related.   -->

```{r est, fig.height = 6.8, fig.cap="Convergence diagnostics regression coefficient."}
# est_bias + est_Rh + est_AC + plot_layout(guides = "collect", ncol = 1) + plot_annotation(tag_levels = "A", tag_suffix = ".", title = bquote("Q = " ~ beta[1] ~ "= 2.06; " ~ theta ~ "= " ~ hat(Q) ~ "in {" ~ y[obs] ~ "," ~ y[imp][",m"] ~ "}"))
```

```{r pred, fig.height = 6.8, fig.cap="Convergence diagnostics regression coefficient."}
# Rsq_bias + PCA_Rh + PCA_AC + plot_layout(guides = "collect", ncol = 1) + plot_annotation(tag_levels = "A", tag_suffix = ".", title = bquote("Q = " ~ r^2 ~ "\u00D7 100 = 19.25; " ~ theta ~ "= " ~ lambda[1] ~ "in {" ~ y[obs] ~ "," ~ y[imp][",m"] ~ "}")) #first eigenvalue of" ~ Sigma ~ 
```

<!-- ## Univariate estimates and convergence diagnostics  -->

<!-- The bias in the estimates of the variable means shows little to no difference between simulation conditions. It doesn't seem to matter how many iterations you use in the mice algorithm, the estimates are unbiased. Similarly, the bias in the estimated variances is more or less stable across simulation conditions. These univariate quantities appear to be unaffected by the number of iterations. -->

<!-- When applied to the imputation chain means, $\widehat{R}$ indicates that the mice algorithm does not reach a converged state ($\widehat{R}$ = 1) in any of the simulation conditions. Neither is the most recent recommended threshold reached ($\widehat{R}$ < 1.01). **This is only for 75% missingness:** The conventional $\widehat{R}$ threshold of 1.2 is reached in simulation conditions $T = 3$ and $T > 6$. (With the default number of iterations (maxit = 5), this dip in $\widehat{R}$ values would be spotted, so it is no problem.) The point at which an extra iteration does not seem to improve the $\widehat{R}$ value is around $T=30$. -->

<!-- Autocorrelations indicate no sign of trending **(NOT correct anymore, fluke of the stats::acf() function!!)** within imputation chains. In most simulation conditions $AC$ is smaller than or about equal to zero. Across simulation conditions, however, the autocorrelation curve does not trend towards 0. Autocorrelation values plateau off at a value of around.1. This is a small positive autocorrelation, which would indicate some trending within chains. -->

<!-- The $\widehat{R}$ and $AC$ values for the imputation chain variances show equal trends to the chain means and are therefore not discussed separately. Taken together, univariate estimates seem robust across simulation conditions. There is no clear effect of the number of iterations on the bias in these estimates, while the convergence diagnostics indicate that the algorithm did not reach a completely converged state (yet). -->

<!-- ## Multivariate estimates  -->

<!-- **This is only true for 75% missingness:** There is a clear bias in the regression estimates in simulation conditions where the number of iterations is smaller than four. In simulation conditions where $T > 5$ there is little to no bias in the estimated regression coefficients. In most simulation conditions nominal coverage is obtained (i.e., coverage rates of.95) for the confidence intervals around the regression coefficients. Conditions with only one or two iterations show some under-coverage. Since confidence interval width is stable across conditions, the under-coverage may be attributed to the bias in the estimated regression coefficients. -->

<!-- **This is only true for 75% missingness:** If we look at the estimated proportion of explained variance in outcome variable Y we see that the coefficient of determination is underestimate estimated in conditions where the number of iterations is equal to two or less, and slightly overestimated in conditions where the number of iterations is equal to three or more.  -->

```{r thresh-Rh, include=FALSE}
# # chain means
# dat[-1, 6:9] %>% mutate(max_Rh = apply(., 1, max)) %>%.$max_Rh<1.2
# 
# # chain variances
# dat[-1, 26:29] %>% mutate(max_Rh = apply(., 1, max)) %>%.$max_Rh<1.2

# qnorm((1 + .95) / 2) / sqrt(t)

```

# Discussion

H1: As hypothesized, biased, invalid estimates occur more often in simulation conditions with a higher missingness proportion and lower number of iterations. However, this only holds for multivariate scientific estimands. Univariate $Q$s are not substantially impacted by early stopping.

H2: As expected, we observe higher $\widehat{R}$ and $AC$-values in conditions where $\bar{Q}$s were biased, invalid estimates of $Q$s.

H3: The hypothesis that the recommended thresholds would be too strict is only patially supported. $\widehat{R}$ indicated improving convergence until 30 to 40 iterations, while approximately unbiased, confidence-valid estimates were obtained after just seven iterations. This rougly corresponds to the threshold of 1.2. The more stringent trheshold 1.1 seems plausible compared to the decrease in $\widehat{R}$ values across iterations. The most recent threshold 1.01 does indeed seem to over-estimate the signs of non-convergence.

H4: The results of this simulation study do not support the hypothesis that ... **[insert hypothesis]**. We observed $AC$-values as high as theoretically possible $AC=$. The second part of the hypothesis is not refuted: the $AC$-values did decrease quicky as a function of $T$.

H5: We did find some support for the final hypothesis. Multivariate $\theta$s indeed have superior performance under certain conditions (e.g., multivariate $Q$s and extreme missingness proportions).

# Conlusion

We arrive back at the original questions that we posed.

- Non-convergence can be identified with non-convergence diagnostics $\widehat{R}$ and $AC$. But common thresholds are too conservative. Why? Part of the distribution is already determined by the observed data. in regular MCMC the entire distribution is unknown (**priors??**). In iterative imputation algorithms we have the same characteristics as MCMC in the limit, but because we combine the known distribution with the unknown distribution, convergence is reached much sooner.

- Inferences are unbiased much quicker than the algorithm has converged. All things univariate are converged almost intantly after initiating the iterative algorithm with a valid imputation model. But less quick with multivariate parameters.

In short, iterative imputation algorithms may yield approximately unbiased, confidence-valid estimates under different levels of induced non-convergence. Univariate estimates seem robust against terminating the algorithm early: There is no clear effect of the number of iterations on the bias in estimated means and standard deviations. The bias in these $Q$s was only impacted by the missingness proporion $p_{\rm mis}$, with higher $p_{\rm mis}$ resulting in more bias. Bias in multivariate estimates depended on both $p_{\rm mis}$ and the number of iterations. However, the only effect of the number of iterations was observed when the algorithm was terminated at seven iterations or earlier. This suggests that continued iteration beyond seven iterations does not yield better estimates. Yet, the non-convergence diagnostics $\widehat{R}$ and $AC$ indicated that the iterative imputation algorithm did not reach a stable state before twenty to thirty iterations. **[MERGE or remove] $\widehat{R}$ and autocorrelation indicate that there are signs of non-convergence in the algorithm up-to at least twenty iterations, while unbiased, confidence-valid estimates may be obtained with as little as one iteration. This observation is in agreement with **one of [still add the others!]** the simulation hypotheses: $\widehat{R}$ over-estimates the severity of non-convergence when applied to MI procedures.**

Over-estimation of the signs of non-convergence may be due to the methods (and their thresholds) or due to the $Q$s (descriptive and multivariate linear regression, not higher dimensional/more complex $Q$s). More complicated $Q$s (e.g., higher-order effects or variance components) might show bias, under- or over-coverage at higher $T$. Which might be why $\widehat{R}$ and $AC$ identify non-convergence.

```{r eval=FALSE, include=FALSE}
# **[INCLUDE this] The highest $\widehat{R}$-values are obtained for the condition $T=2$. We observe the same 'dip' in $\widehat{R}$ as we did in Figure \ref{fig:diagnostics} for conditions with three iterations. $\widehat{R}$ gradually decreases as the number of iterations increases across conditions. By the traditional threshold  $\widehat{R}>1.2$ we should conclude that there is non-convergence in conditions where $T<8$. The common threshold $\widehat{R}>1.1$ is overcome in conditions where $T>14$. **
#   
#   **Autocorrelations indicate some signs of non-stationarity in all simulation conditions, but primarily in conditions where $T<6$. The threshold for diagnosing non-convergence is not reached in any condition. We would thus conclude that the algorithm is sufficiently converged regardless of the number of iterations or missingness proportion. **
```

Some other interesting findings are **[REPHRASE]**

- Is there more bias in conditions where non-convergence was induced? (YES for multivar., NOPE for univariates).

- Is the point of *sufficient* convergence (at which the $\bar{Q}$s are unbiased and confidence-valid estimates of the $Q$s) the same point as identified by the non-convergence diagnostics? (NOPE, they are too conservative).

- Is one of the thetas better? (YES, while it doesn't really matter as long as it's multivariate, $\lambda_{1, \ell}$ is model-independent and therefore preferred).

- What should the thresholds be? (much HIGHER than recommended for $\widehat{R}$, difficult to say for $AC$).

- And the default nr of iterations? (MAYBE increase to 10 to detect significant $AC$ and be sure not to miss the dip in $\widehat{R}$?).

<!-- As expected, complete convergence ($\widehat{R}=1$ and $AC=0$) is never observed. But what about the thresholds for practice? $\widehat{R}$ <1.2 is not stringent enough, because conditions in which $\widehat{R}$ is smaller than 1.2 do not all yield unbiased estimates of the $Q$s. Moreover, there is a dip in the $\widehat{R}$ values, after which $\widehat{R}$ > 1.2 again. If the algorithms would be terminated at the iteration where $\widehat{R}$ < 1.2 occurs for the first time, the increase in $\widehat{R}$ values might be missed. The threshold $\widehat{R}$ < 1.1 is somewhat conservative in comparison with the performance measures. Overall it seems OK. $\widehat{R}$ < 1.01 is too stringent compared to the performance measures. And it is not obtained for any number of iterations (**but may be necessary in more complex complete data models???**).  -->

<!-- According to this simulation study, the recently proposed threshold of $\widehat{R}<1.01$ may be too stringent for MI algorithms. Under the relatively easy missing data problem of the current study, the threshold was not reached. The other extreme of the $\widehat{R}$-thresholds, the conventionally acceptable $\widehat{R} <1.2$, may be too lenient for MI procedures. Applying this threshold to the current data, lead to falsely diagnosing convergence at $T = 3$ (**because it goes up after, not because it is not converged enough**). It appears that the widely used threshold of $\widehat{R} < 1.1$ suits MI algorithms the best. We might, however, also formulate a new threshold, specifically for the evaluation of MI algorithms. The current study suggests that $\widehat{R} < 1.05$ may be implemented, since that is the level at which the $\widehat{R}$ stabilize (around $T = 20$) (**Not necessary for this Q, but maybe for more complicated $Q$s**).  -->

<!-- The negative $AC$-values obtained in this study show no threat of non-stationarity. However, the initial dip in $AC$-values **(which disappeared!)** may have implications for the default number of iterations in `mice` (`maxit = 5`). Terminating the algorithm at $T=5$ may not be the most appropriate, since this lead to the worst convergence (**nope, only for Rh, not $AC$**), as indicated by $\widehat{R}$ and $AC$. Under the current specifications, $T>20$ would be more appropriate. -->

## Recommendations for empirical researchers

After imputation it is always wise to take the following steps:

1. First, check traceplots for signs of pathological non-convergence. Adjust imputation model if necessary.

2. Monitor non-convergence diagnostics $\widehat{R}$ and $AC$. Do not use the traditional calculation for $\widehat{R}$, or the `R` function `stats::acf()` to compute $AC$. Instead, calculate $\widehat{R}$ conform @veht19 and compute autocorrelations manually (see e.g., [github.com/gerkovink/shinyMice](https://github.com/gerkovink/shinyMice/tree/master/3.Thesis/1.SimulationStudy)).

3. Track $\theta$s over iterations, e.g., the novel $\theta$ that is 'substantive model-independent'.

4. Track your own scalar summary of interest (see e.g., @buur18). Compute $\widehat{R}$ and autocorrelation values for this scalar summary.

5. Use the threshold $\widehat{R}>1.2$ for mixing, and assess stationarity by plotting the autocorrelation over iterations. Keep iterating until the threshold for mixing is reached, and until a reasonably decreasing asympototic $AC$ value is obtained. At that point, inferences are unlikely to improve by continued iteration.

## Recommendations for future research

Much remains unknown. about non-convergence in iterative imputation algorithms. Further research is needed to investigate the performance of iterative imputation algorithms under clear violation of convergence, e.g. dependency between predictors (predictors with very high correlations). For example:

- Induce non-convergence with mis-specified imputation model.

- Introduce more difficult missingness problems, i.e., M(N)AR instead of MCAR. That is, proper performance under a 'missing completely at random' missingness mechanism is necessary to demonstrate the appropriateness of $\widehat{R}$ and $AC$ as non-convergence diagnostics. However, results may not be extrapolated to other missingness mechanisms.

- Imputation technique with questionable convergence properties, e.g., PMM.

- Apply recommendations to empirical data and write a vignette for applied researchers.

- Implement recommendations in imputation software, e.g., evaluation environment for applied researchers.

- Develop an additional convergence diagnostic unique for iterative imputation: $AC$ across imputations, instead of iterations.

In short, we have shown that an iterative imputation algorithm can yield correct outcomes, even when a converged state has not yet formally been reached. Any further iterations would then burn computational resources without improving the statistical inferences. To conclude, in the case of drawing inference from incomplete data, convergence of the iterative imputation algorithm is often a convenience but not a necessity. **[Our study found that - in the cases considered - inferential validity was achieved after 5-10 iterations, much earlier than indicated by the $\widehat{R}$ and AC diagnostics. Of course, it never hurts to iterate longer, but such calculations hardly bring added value.]**

<!-- # Grading -->

<!-- - Scientific contribution  -->
<!-- (clear research question, convincing argumentation that the study adds to the scientific literature) -->

<!-- - Theory -->
<!-- (use of theory, clear hypotheses, relevant literature) -->

<!-- - Method -->
<!-- (correct data collection/handling, measures, choice and description of method of analysis) -->

<!-- - Results -->
<!-- (correct interpretation of findings, clear presentation) -->

<!-- - Discussion and conclusions -->
<!-- (summary, broader implications, limitations, future research) -->

<!-- - Written presentation -->
<!-- (writing, structure, tables and figures, references) -->

<!-- # References -->