### 2020-07-13 ICML/MissingThePoint.Rmd
Create presentation for ARTEMISS workshop

---
title: "Missing The Point:"
subtitle: "Non-Convergence in Iterative Imputation Algorithms"
author: "Hanne I. Oberman, Stef van Buuren, Gerko Vink"
institute: "Utrecht University"
date: "17-07-2020" #(updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    css: [default, default-fonts]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

# Algorithmic Non-Convergence

With iterative imputation, the validity of the inferences relies on algorithmic convergence.

```{r non-conv, echo=FALSE,  message=FALSE, warning=FALSE}
# set-up env
library(dplyr)
library(ggplot2)
library(plotly)
load("example_diagnostics.Rdata")

# define colorblind friendly colors
paint2 <- c('#66CCEE', '#EE6677')

# plot chains
trace <- ggplot(diagnostics) +
  geom_point(aes(x = iteration, y = `Chain 1`, color = patho),
             size = .25,
             na.rm = TRUE) +
  geom_point(aes(x = iteration, y = `Chain 2`, color = patho),
             size = .25,
             na.rm = TRUE) +
  geom_point(aes(x = iteration, y = `Chain 3`, color = patho),
             size = .25,
             na.rm = TRUE) +
  geom_point(aes(x = iteration, y = `Chain 4`, color = patho),
             size = .25,
             na.rm = TRUE) +
  geom_point(aes(x = iteration, y = `Chain 5`, color = patho),
             size = .25,
             na.rm = TRUE) +
  geom_line(aes(x = iteration, y = `Chain 1`, color = patho), size = .25, na.rm = TRUE) +
  geom_line(aes(x = iteration, y = `Chain 2`, color = patho), size = .25, na.rm = TRUE) +
  geom_line(aes(x = iteration, y = `Chain 3`, color = patho), size = .25, na.rm = TRUE) +
  geom_line(aes(x = iteration, y = `Chain 4`, color = patho), size = .25, na.rm = TRUE) +
  geom_line(aes(x = iteration, y = `Chain 5`, color = patho), size = .25, na.rm = TRUE) +
  scale_colour_manual(values=paint2) +
  scale_x_continuous(breaks = 1:10) +
  xlab("Iteration") +
  ylab("Chain mean") + #bquote(theta)) +
  labs(color = "") + 
  theme_classic() +
  theme(legend.position = "bottom")

ggplotly(trace)

# alternatively, use the actual data
# run 1.Execute.R until imputing anything, than run RunSimulationOnce.R until after amputing the data
# with the 'amps' object
load("../traceplotdata.Rdata")
set.seed(12)
d75 <- amps[[4]] %>% mice(maxit=20)
plot(d75)
gg.mids(d4, x="Y")[[1]]
```

---

# Results

```{r results, echo=FALSE, fig.width=9.8, message=FALSE, warning=FALSE}
load("complete.Rdata") 
#library(patchwork)

# pre-processing
results <- results %>% filter(t<51) %>% mutate(bias.est.X1 = bias.est.X1/.0206) %>% mutate(crit = qnorm((1 + .95) / 2) / sqrt(t), thresh1.01 = 1.01, thresh1.1 = 1.1, thresh1.2 = 1.2)
results$crit[results$crit>1] <- NA

# define colorblind friendly colors
paint5 <- c('#228833', '#66CCEE', '#CCBB44','#EE6677', '#AA3377')

# create plots
est_bias <- results %>% ggplot(aes(x = t, y = bias.est.X1, color = as.factor(p*100))) +
  geom_hline(yintercept = 0,
             color = "grey",
             lwd = 1) +
  geom_point(size = .25, na.rm = TRUE) +
  geom_line(size = .25, na.rm = TRUE) +
  scale_colour_manual(values=paint5) +
  xlab("") +
  ylab("Bias (%)") +
  labs(colour = "Proportion of missing cases (%)") +
  theme_classic() +
  theme(legend.position = "bottom")

est_Rh <- results %>% ggplot(aes(x = t, y = r.hat.max.beta, color = as.factor(p*100))) +
  geom_hline(yintercept = 1,
             color = "grey",
             lwd = 
               1) +
  geom_point(size = .25, na.rm = TRUE) +
  geom_line(size = .25, na.rm = TRUE) +
  geom_line(aes(x=t, y=thresh1.2), color = "grey", linetype = "dashed", size = .25, na.rm = TRUE) +
  geom_line(aes(x=t, y=thresh1.1), color = "grey", linetype = "dashed", size = .25, na.rm = TRUE) +
  geom_line(aes(x=t, y=thresh1.01), color = "grey", linetype = "dashed", size = .25, na.rm = TRUE) +
  scale_colour_manual(values=paint5) +
  scale_y_continuous(limits = c(1,1.53), breaks = c(1,1.1,1.2,1.3,1.4,1.5)) +
  xlab("Number of iterations") +
  ylab(bquote(widehat(R))) +
  labs(colour = "Proportion of missing cases (%)") +
  theme_classic() +
  theme(legend.position = "")

est_AC <- results %>% ggplot(aes(x = t, y = ac.max.beta, color = as.factor(p*100))) +
  geom_hline(yintercept = 0,
             color = "grey",
             lwd = 1) +
  geom_line(aes(x = t, y = crit), color = "grey", linetype = "dashed", size = .25, na.rm = TRUE) + 
  geom_point(size = .25, na.rm = TRUE) +
  geom_line(size = .25, na.rm = TRUE) +
  scale_colour_manual(values=paint5) +
  xlab("Number of iterations") +
  ylab("Autocorrelation") + 
  labs(colour = "Proportion of missing cases (%)") +
  theme_classic() +
  theme(legend.position = "")

est_cov <- results %>% ggplot(aes(x = t, y = cov.est.X1*100, color = as.factor(p*100))) +
  geom_hline(yintercept = 95,
             color = "grey",
             lwd = 1) +
  geom_point(size = .25, na.rm = TRUE) +
  geom_line(size = .25, na.rm = TRUE) +
  scale_colour_manual(values=paint5) +
  xlab("") +
  ylab("Coverage rate (%)") +
  labs(colour = "Proportion of missing cases (%)") +
  theme_classic() +
  theme(legend.position = "")

# plot estimates, save 6.75x4.5in or 5.5
# est_bias + est_cov + est_AC + est_Rh + plot_annotation(tag_levels = "A", tag_suffix = ".") + plot_layout(ncol = 2, guides = "collect")

ggplotly(est_bias)

```

---

# Thanks!

Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).