### 2020-07-20 shinymice/about.Rmd
Initial version shinymice app

---
  title: "About"
output: html_document
---

This Shiny app has the sole purpose of showing that I'm able to program an app in `R Shiny`. However, if all goes well, I might use this as the foundation of the interactive evaluation suite for evaluating multiply imputed data that I'm developing as Research Master's thesis project. 'ShinyMICE' will be integrated into the `mice()` environment in `R`.

Watch the current state of the project on [github.com/gerkovink/ShinyMICE](https://github.com/gerkovink/ShinyMICE/).