### 2020-09-02 www/about.Rmd
Use psych::describe() instead of custom functions for descriptives #47

---
  title: "About"
output: html_document
---

This Shiny app has the sole purpose of showing that I'm able to program an app in `R Shiny`. However, if all goes well, I might use this as the foundation of the interactive evaluation suite for evaluating multiply imputed data that I'm developing as Research Master's thesis project. `shinymice` will be integrated into the `mice` environment in `R`.

Watch the current state of the project on [github.com/amices/shinyMice](https://github.com/amices/shinyMice/).