### 2021-12-16 README.Rmd
Remove duplicate info from the repo

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# shinymice

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

This repository serves the development of `{shinymice}`, an interactive evaluation device for missing data by Hanne Oberman, guided by Gerko Vink and Stef van Buuren.

![shinymice hex sticker](inst/app/www/logo.png)

For a demo version of the app, see https://hanneoberman.shinyapps.io/shinymice-demo/.

-Hanne (16-12-2021)

<!-- The goal of shinymice is to ... -->

<!-- ## Installation -->

<!-- You can install the released version of shinymice from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("shinymice") -->
<!-- ``` -->

<!-- And the development version from [GitHub](https://github.com/) with: -->

<!-- ``` r -->
<!-- # install.packages("devtools") -->
<!-- devtools::install_github("amices/shinyMice") -->
<!-- ``` -->
<!-- ## Example -->

<!-- This is a basic example which shows you how to solve a common problem: -->

<!-- ```{r example} -->
<!-- library(shinymice) -->
<!-- ## basic example code -->
<!-- ``` -->

<!-- What is special about using `README.Rmd` instead of just `README.md`? You can include R chunks like so: -->

<!-- ```{r cars} -->
<!-- summary(cars) -->
<!-- ``` -->

<!-- You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. -->

<!-- You can also embed plots, for example: -->

<!-- ```{r pressure, echo = FALSE} -->
<!-- plot(pressure) -->
<!-- ``` -->

<!-- In that case, don't forget to commit and push the resulting figure files, so they display on GitHub! -->